---
output:
  word_document:
    reference_docx: ../MPFF report template v1.1.dotx
---


```{r setup, include=FALSE}

# ==========================================================================================================
# Generate FLYSHOOT trips overview
# 
# 8/3/2023 first coding
# ==========================================================================================================

knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

options(dplyr.summarise.inform = FALSE)

# Reset lists
rm(list=ls())

# Libraries
library(rmarkdown)                   # note: requires knitr 1.21

require(tidyverse, quietly=TRUE)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
require(lubridate, quietly=TRUE)     # data handling
require(reshape2, quietly=TRUE)      # reshaping data; e.g. cast
require(RColorBrewer)                # colour schemes
library(viridis)
library(pander)

library(captioner)                   # for captions
tab_nums <- captioner::captioner(prefix = "Table " , levels=1, type=c("n"), infix=".")
fig_nums <- captioner::captioner(prefix = "Figure ", levels=1, type=c("n"), infix=".")

# Source all the utils
source("../r/FLYSHOOT utils.R")

spatialdir = "C:/Users/MartinPastoors/OneDrive - Martin Pastoors/DATA/RDATA"

# spatial data files

load(file.path(spatialdir, "world_mr_sf.RData"))
load(file.path(spatialdir, "world_mr_df.RData"))
load(file.path(spatialdir, "rect_lr_sf.RData"))

load(file.path(spatialdir, "world_lr_df.RData"))

load(file.path(spatialdir, "asfis.RData"))

# ===============================================================================
#  Load and filter the trip data
# ===============================================================================

# set onedrive directory
onedrive <- get_onedrive(team="Martin Pastoors", site="FLYSHOOT - General")

# load datasets
haul   <- loadRData(file.path(onedrive, "rdata/haul.RData"))
kisten <- loadRData(file.path(onedrive, "rdata/kisten.RData")) 
elog   <- loadRData(file.path(onedrive, "rdata/elog.RData")) %>% 
  mutate(species = toupper(species)) %>% 
  mutate(species = ifelse(species=="JAX","HOM",species))
gfw   <- loadRData(file.path(onedrive, "rdata/gfw.RData")) 

yrs     <- 2021:2025
mnths   <- NA
vessels <- c("SL9","SCH99", "Z99", "SCH135","SCH65")

myyear <- max(yrs)

# making the data selections
h <- 
  haul %>%
  ungroup() %>% 
  {if(!is.na(all(vessels))) filter(., vessel %in% vessels) else . } %>% 
  {if(!is.na(all(yrs)))     filter(., year %in% yrs) else . } %>% 
  {if(!is.na(all(mnths)))   filter(., month %in% mnths) else .} %>% 
  ungroup()

# h %>%
#   dplyr::select(shoottime, haultime, lat, lon, windforce, winddirection, depth=waterdepth) %>%
#   sample_n(1100) %>%
#   drop_na(shoottime, haultime) %>% 
#   writexl::write_xlsx(path="testfile.xlsx")


k <- 
  kisten %>%
  mutate(
    year  = lubridate::year(datetime),
    month = lubridate::month(datetime),
    week = lubridate::week(datetime)
  ) %>% 
  ungroup() %>% 
  {if(!is.na(all(vessels))) filter(., vessel %in% vessels) else . } %>% 
  {if(!is.na(all(yrs)))     filter(., year %in% yrs) else . } %>% 
  {if(!is.na(all(mnths)))   filter(., month %in% mnths) else .} %>% 
  mutate(haul = ifelse(is.na(haul), haul2, haul)) %>% 
  dplyr::select(-haul2) %>% 
  left_join(dplyr::select(h, 
                          vessel, trip, haul, division, lat, lon, rect),
            by=c("vessel","trip","haul"))

e <- 
  elog %>%
  ungroup() %>% 
  {if(!is.na(all(vessels))) filter(., vessel %in% vessels) else . } %>% 
  {if(!is.na(all(yrs)))     filter(., year %in% yrs) else . } %>% 
  {if(!is.na(all(mnths)))   filter(., month %in% mnths) else .} %>% 
  ungroup() %>% 
  # mutate(species = tolower(species)) %>% 
  mutate(species = ifelse(species == "SQC", "SQR", species)) %>% 
  left_join(dplyr::select(asfis,
                          species, scientificname, englishname, dutchname),
            by = "species") %>% 
  mutate(species = toupper(species))  

g <- 
  gfw %>%
  ungroup() %>% 
  {if(!all(is.na(vessels))) filter(., vessel %in% vessels) else . } %>% 
  {if(!all(is.na(yrs)))     filter(., year %in% yrs) else . } %>% 
  {if(!all(is.na(mnths)))   filter(., month %in% mnths) else .} %>% 
  dplyr::select(vessel, year, month, week, lon, lat)

# calculate the scaling of plots
xmin <- floor(2 * (min(e$lon, na.rm=TRUE)-0.5))/2;
xmax <- ceiling(2 * (max(e$lon, na.rm=TRUE)+0.5))/2;
ymin <- floor(2 * (min(e$lat, na.rm=TRUE)-0.5))/2;
ymax <- ceiling(2* (max(e$lat, na.rm=TRUE)+0.5))/2



n <- 15

effort <-
  e %>% 
  dplyr::group_by(vessel, year) %>% 
  dplyr::summarize(ndays = n_distinct(paste(vessel, date))) 

catch <-
  e %>% 
  dplyr::group_by(vessel, year, species, 
                  scientificname, englishname, dutchname) %>% 
  dplyr::summarize(weight = sum(weight, na.rm=TRUE)) %>% 
  dplyr::group_by(vessel, year) %>% 
  dplyr::mutate(perc = weight/sum(weight, na.rm=TRUE)) %>% 
  left_join(effort, by=c("vessel", "year")) %>% 
  mutate(catch_day = weight / ndays)

top <-
  catch %>% 
  # filter(year == max(yrs)) %>% 
  dplyr::group_by(species) %>% 
  dplyr::summarize(weight     = sum(weight, na.rm=TRUE)) %>% 
  arrange(desc(weight)) %>% 
  slice_head(n=n)


```


**FLYSHOOT trips overview**

For: Jaczon BV

By: Martin Pastoors

Report generated: `r format(Sys.time(), '%d/%m/%Y')`


```{r triptable, echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.asp=0.6}

tab_nums(name    = "trips", level = 1, display = FALSE, caption = "Self-sampling overview")

t <-
  bind_cols(
    h %>% 
      filter(year %in% yrs) %>% 
      group_by(year, month) %>% 
      summarise(
        nvessels      = n_distinct(vessel),
        ntrips        = n_distinct(paste(vessel, trip)),
        ndays         = n_distinct(paste(vessel, trip, date)),
        nhauls        = n_distinct(paste(vessel, trip, haul)),
        nrects        = n_distinct(rect),
        # nlandedweight  = sum(!is.na(landingweight)),
        ntotalcatch   = sum(!is.na(catchweight))
      ),
    k %>% 
      filter(year %in% yrs) %>% 
      group_by(vessel, trip, haul, year, month) %>% 
      summarise(gewicht = sum(gewicht, na.rm=TRUE)) %>% 
      group_by(year, month) %>% 
      summarise(
        nlandedcatch  = sum(!is.na(gewicht)),
        landedcatch   = sum(gewicht)
      ) %>% 
      ungroup() %>% 
      dplyr::select(-year, -month)
    
  )  %>% 
  
  # discard ratio (for Aravis only)
  left_join(
    k %>% 
      filter(vessel %in% c("SCH99","Z99"), as.Date(datetime) >= lubridate::dmy("1/11/2023")) %>% 
      group_by(vessel, trip, haul, year, month) %>% 
      summarise(gewicht = sum(gewicht, na.rm=TRUE)) %>% 
      left_join(h, by=c("vessel","trip","haul", "year", "month")) %>% 
      filter(!is.na(catchweight)) %>% 
    
      group_by(year, month) %>% 
      summarise(
        landedcatch   = sum(gewicht, na.rm=TRUE),
        totalcatch    = sum(catchweight, na.rm=TRUE)
      ) %>% 
      mutate(discardperc = scales::percent(1-(landedcatch/totalcatch))) %>% 
      ungroup() %>% 
      dplyr::select(-landedcatch, -totalcatch),
    by = c("year","month")

  ) %>% 
  
  mutate(year=as.character(year), month=as.character(month)) 

t2 <-
  t %>% 
  ungroup() %>% 
  summarise(
    ndays = sum(ndays),
    nhauls = sum(nhauls),
    ntotalcatch = sum(ntotalcatch),
    nlandedcatch = sum(nlandedcatch),
    landedcatch = sum(landedcatch)
  ) %>% 
  mutate(year = as.character(myyear)) %>% 
  mutate(month = "all")

bind_rows(t, t2) %>% 
  pander::pandoc.table(., 
               style        = "simple",
               split.tables = 200, 
               split.cells  = c(rep(7,10)),
               justify      = "right",
               missing      =" ",
               big.mark     = ',', 
               round        = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0) )

  


```

_`r tab_nums("trips")`_

\newpage

```{r eval=FALSE, fig.asp=0.6, message=FALSE, warning=FALSE, comment=NA, include=FALSE}

# NOT USED
fig_nums(
  name    = "c", 
  level = 1, display = FALSE,
  caption = "Combined look at different data sources by vessel and trip")

hh <- 
  h %>%
  group_by(vessel,trip) %>%
  summarise(
    nhauls            = n(),
    startdate         = min(date, na.rm=TRUE),
    enddate           = max(date, na.rm=TRUE),
    weeks             = paste(unique(week), collapse=","),
    months            = paste(unique(month), collapse=","),
    divisions         = paste(unique(division), collapse=";"),
    rects             = paste(unique(rect), collapse=";"),
  ) %>% 
  mutate (
    ndays             = as.numeric(difftime(enddate, startdate, units="days")) + 1,
    haulsperday       = nhauls/ndays
  ) 

kk <- 
  k %>%
  group_by(vessel,trip) %>%
  summarise(
    nkisten           = n(),
    nsoorten          = n_distinct(soorten),
    startdate         = as.Date(min(datetime, na.rm=TRUE)),
    enddate           = as.Date(max(datetime, na.rm=TRUE)),
    weeks             = paste(unique(week(datetime)), collapse=","),
    months            = paste(unique(month(datetime)), collapse=","),
    gewicht           = sum(gewicht, na.rm=TRUE)
  ) %>% 
  mutate (
    ndays             = as.numeric(difftime(enddate, startdate, units="days")) + 1
  ) 

ee <- 
  e %>%
  filter(year %in% myyear) %>% 
  group_by(vessel,trip) %>%
  summarise(
    startdate         = min(date, na.rm=TRUE),
    enddate           = max(date, na.rm=TRUE),
    weeks             = paste(unique(week), collapse=","),
    months            = paste(unique(month), collapse=","),
    divisions         = paste(unique(faozone), collapse=";"),
    rects             = paste(unique(rect), collapse=";"),
    nsoorten          = n_distinct(species),
    gewicht           = sum(weight, na.rm=TRUE)
  ) %>% 
  mutate (
    ndays             = as.numeric(difftime(enddate, startdate, units="days")) + 1
  ) 

isEven <- function(x) {
    ## traditional modulo check
    x %% 2 == 0
}

# summarize the trip informatie
comb <-
  bind_rows(
    hh %>% mutate(source="haul"),
    kk %>% mutate(source="marelec"),
    ee %>% mutate(source="elog")
  ) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  dplyr::select(vessel, trip, startdate, enddate, source) %>% 
  mutate(source = factor(source, levels=c("haul", "marelec", "elog"))) %>% 
  mutate(trip = gsub("^2023|^2024","", trip)) %>% 
  mutate(vjust=ifelse(isEven(row_number()), 1, -1))

comb %>% 
  ggplot() +
  theme_publication() +
  ggalt::geom_dumbbell(aes(x = startdate, xend = enddate, y=source, group=1, colour=source),
                       size=3, size_x = 0, size_xend = 0, alpha=0.4) +
  # geom_text(data=comb %>% filter(source=="haul"),
  #           aes(x = startdate, y=source, label=trip, vjust=vjust), hjust=0.5, size=3) +
  facet_grid(vessel~.)

# _`r fig_nums("c")`_

```



Totale aanvoer per schip en per week

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.asp=0.6}

fig_nums(
  name    = "d", 
  level = 1, display = FALSE,
  caption = "Totale aanvoer per schip en per week. Zwarte lijn is de gemiddelde aanvoer per schip en per week")

# kk <- 
#   k %>%
#   mutate(week=lubridate::week(datetime)) %>% 
#   mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
#   group_by(vessel, year, week) %>% 
#   summarise(landingweight = sum(gewicht, na.rm=TRUE)) %>% 
#   group_by(vessel, year) %>% 
#   mutate(meanweight = mean(landingweight, na.rm=TRUE))

# k %>% mutate(week=lubridate::week(datetime)) %>% filter(week==5, vessel=="SCH135") %>% View()  
# k %>% filter(grepl("_", trip)) %>% distinct(trip) 
# k %>% filter(grepl("085", trip)) %>% distinct(trip) 

# plot total catch and landings
# kk %>% 
#   ggplot(aes(group=vessel)) +
#   theme_publication() +
#   theme(legend.position = "none") +
#   geom_bar(aes(x=week, y=landingweight, fill=vessel), stat="identity") +
#   geom_hline(aes(yintercept=meanweight, colour=vessel)) +
#   geom_text(aes(x=1, y=meanweight, label=as.integer(meanweight)), hjust=0, vjust=0, nudge_y = 500) +
#   facet_grid(year~vessel)

kk <- 
  e %>%
  mutate(week=lubridate::week(catchdate)) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  group_by(vessel, year, week) %>% 
  summarise(landingweight = sum(weight, na.rm=TRUE)) %>% 
  group_by(vessel, year) %>% 
  mutate(meanweight = mean(landingweight, na.rm=TRUE)) 

se <- function(x) sqrt(var(x)/length(x))

kkk <- 
  kk %>%
  filter(year %in% yrs[-length(yrs)]) %>% 
  group_by(vessel, week) %>% 
  summarise(
    meanweightperweek = mean(landingweight, na.rm=TRUE),
    sd                = sd(landingweight,na.rm=TRUE),
    sterr             = se(landingweight)
  )

# k %>% mutate(week=lubridate::week(datetime)) %>% filter(week==5, vessel=="SCH135") %>% View()  
# k %>% filter(grepl("_", trip)) %>% distinct(trip) 
# k %>% filter(grepl("085", trip)) %>% distinct(trip) 

# plot total catch and landings
kk %>% 
  ggplot(aes(group=vessel)) +
  theme_publication() +
  theme(legend.position = "none") +
  geom_bar(aes(x=week, y=landingweight, fill=vessel), stat="identity") +
  
  geom_step(data=kkk, aes(x=week, y=meanweightperweek), fill=NA, colour="black") +
  # geom_point(data=kkk, aes(x=week, y=meanweightperweek), fill=NA, colour="black") +
  # geom_errorbar(data=kkk, aes(x=week, ymin=meanweightperweek-sterr, ymax=meanweightperweek+sterr), fill=NA, colour="gray") +
  
  # geom_hline(aes(yintercept=meanweight, colour=vessel)) +
  # geom_text(aes(x=1, y=meanweight, label=as.integer(meanweight)), hjust=0, vjust=0, nudge_y = 500) +
  facet_grid(vessel~year)

```

\newpage

Aanvoer per soort en per week

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.asp=1.0}

# kk <- 
#   k %>%
#   filter(species %in% c("CTC", "SQR", "MUR","MAC")) %>% 
#   mutate(week=lubridate::week(datetime)) %>% 
#   mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
#   group_by(species, vessel, year, week) %>% 
#   summarise(landingweight = sum(gewicht, na.rm=TRUE)) %>% 
#   group_by(species, year) %>% 
#   mutate(meanweight = mean(landingweight, na.rm=TRUE))

# k %>% mutate(week=lubridate::week(datetime)) %>% filter(week==5, vessel=="SCH135") %>% View()  
# k %>% filter(grepl("_", trip)) %>% distinct(trip) 
# k %>% filter(grepl("085", trip)) %>% distinct(trip) 

# plot total catch and landings
# kk %>% 
#   ggplot(aes(group=vessel)) +
#   theme_publication() +
#   # theme(legend.position = "none") +
#   # geom_bar(aes(x=week, y=landingweight, fill=vessel), stat="identity") +
#   geom_line(aes(x=week, y=landingweight, colour=vessel)) +
#   geom_hline(aes(yintercept=meanweight)) +
#   geom_text(aes(x=1, y=meanweight, label=as.integer(meanweight)), hjust=0, vjust=0, nudge_y = 200) +
#   facet_grid(rows=vars(year), cols=vars(species), scales="free_y")

# ee <- 
#   e %>%
#   filter(species %in% top$species[1:6]) %>% 
#   mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
#   group_by(species, year, week) %>% 
#   summarise(landingweight = sum(weight, na.rm=TRUE)) %>% 
#   group_by(species, year) %>% 
#   mutate(meanweight = mean(landingweight, na.rm=TRUE)) %>% 
#   mutate(year = as.character(year)) %>% 
#   mutate(species = factor(species, levels=top$species[1:6]))

# plot total catch and landings
# ee %>% 
#   ggplot() +
#   theme_publication() +
#   # theme(legend.position = "none") +
#   # geom_bar(aes(x=week, y=landingweight, fill=vessel), stat="identity") +
#   geom_bar(aes(x=week, y=landingweight, fill=species), stat="identity") +
#   facet_wrap(~year, ncol=1)

ee <- 
  e %>%
  filter(species %in% top$species[1:6]) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  group_by(species, year, week) %>% 
  summarise(landingweight = sum(weight, na.rm=TRUE)) %>% 
  group_by(species, year) %>% 
  mutate(meanweight = mean(landingweight, na.rm=TRUE)) %>% 
  mutate(year = as.character(year)) %>% 
  mutate(species = factor(species, levels=top$species[1:6]))

eee <- 
  ee %>%
  filter(year %in% yrs[-length(yrs)]) %>% 
  group_by(species, week) %>% 
  summarise(
    meanweightperweek = mean(landingweight, na.rm=TRUE),
    sd                = sd(landingweight,na.rm=TRUE),
    sterr             = se(landingweight)
  )

# plot total catch and landings
ee %>% 
  ggplot() +
  theme_publication() +
  theme(legend.position = "none") +
  # geom_bar(aes(x=week, y=landingweight, fill=vessel), stat="identity") +
  geom_bar(aes(x=week, y=landingweight, fill=species), stat="identity") +
  geom_step(data=eee, aes(x=week, y=meanweightperweek), fill=NA, colour="black") +
  facet_grid(species~year, scales="free_y")

```


```{r eval=FALSE, fig.asp=1.0, message=FALSE, warning=FALSE, comment=NA, include=FALSE}

# NOT SHOWN

myspecies <- "SQR"

ee <- 
  e %>%
  filter(species %in% myspecies) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  group_by(species, vessel, year, week) %>% 
  summarise(landingweight = sum(weight, na.rm=TRUE)) %>% 
  group_by(species, vessel, year) %>% 
  mutate(meanweight = mean(landingweight, na.rm=TRUE)) %>% 
  mutate(year = as.character(year)) 

eee <- 
  ee %>%
  filter(year %in% yrs[-length(yrs)]) %>% 
  group_by(species, week) %>% 
  summarise(
    meanweightperweek = mean(landingweight, na.rm=TRUE),
    sd                = sd(landingweight,na.rm=TRUE),
    sterr             = se(landingweight)
  )

# plot total catch and landings
ee %>% 
  ggplot() +
  theme_publication() +
  theme(legend.position = "none") +
  # geom_bar(aes(x=week, y=landingweight, fill=vessel), stat="identity") +
  geom_bar(aes(x=week, y=landingweight, fill=species), stat="identity") +
  geom_step(data=eee, aes(x=week, y=meanweightperweek), fill=NA, colour="black") +
  labs(title=myspecies) +
  facet_grid(vessel~year)

```

\newpage

Aanvoer per soort, per jaar en per week. Rode lijn is het lopende jaar. 

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.asp=1.0}

ee <- 
  e %>%
  filter(species %in% top$species) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  group_by(species, year, week) %>% 
  summarise(landingweight = sum(weight, na.rm=TRUE)) %>% 
  group_by(species, year) %>% 
  mutate(meanweight = mean(landingweight, na.rm=TRUE)) %>% 
  mutate(year = as.character(year)) %>% 
  mutate(species = factor(species, levels=top$species))



# plot total catch and landings
ee %>% 
  ggplot(aes(group=year)) +
  theme_publication() +
  # theme(legend.position = "none") +
  geom_line(aes(x=week, y=landingweight, colour=year, linewidth=year, alpha=year)) +
  scale_linewidth_manual("year", values = c(.5, .5, .5, .5, 1.5), guide = "none") +
  scale_colour_manual("year", values = c("gray80", "gray60", "gray40", "gray20", "red")) +
  scale_alpha_manual("year", values = c(1, 1, 1, 1, 0.5), guide = "none") +
  # geom_hline(aes(yintercept=meanweight, colour=species)) +
  # geom_text(aes(x=1, y=meanweight, label=as.integer(meanweight)), hjust=0, vjust=0, nudge_y = 200) +
  facet_wrap(~species, scales="free_y")
  # facet_grid(rows=vars(year), cols=vars(species), scales="free_y")



```

\newpage

Cumulatieve aanvoer per jaar en per week. Rode lijn is het lopende jaar. 

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.asp=1.0}

ee <- 
  e %>%
  group_by(year, week) %>% 
  summarise(landingweight = sum(weight, na.rm=TRUE)) %>% 
  group_by(year) %>% 
  mutate(cumweight = cumsum(landingweight/1000)) %>% 
  mutate(year = as.character(year)) 

# plot total catch and landings
ee %>% 
  ggplot() +
  theme_publication() +
  geom_line(aes(x=week, y=cumweight, colour=year, linewidth=year, alpha=year)) +
  scale_linewidth_manual("year", values = c(.5, .5, .5, .5, 1.5), guide = "none") +
  scale_colour_manual("year", values = c("gray80", "gray60", "gray40", "gray20", "red")) +
  scale_alpha_manual("year", values = c(1, 1, 1, 1, 0.5), guide = "none") +
  labs(y="aanvoer (ton)")


```

\newpage

Cumulatieve aanvoer per soort, per jaar en per week. Rode lijn is het lopende jaar. 

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.asp=1.0}

ee <- 
  e %>%
  filter(species %in% top$species) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  group_by(species, year, week) %>% 
  summarise(landingweight = sum(weight, na.rm=TRUE)) %>% 
  group_by(species, year) %>% 
  mutate(cumweight = cumsum(landingweight)) %>% 
  mutate(year = as.character(year)) %>% 
  mutate(species = factor(species, levels=top$species))

# plot total catch and landings
ee %>% 
  ggplot() +
  theme_publication() +
  # theme(legend.position = "none") +
  geom_line(aes(x=week, y=cumweight, colour=year, linewidth=year, alpha=year)) +
  scale_linewidth_manual("year", values = c(.5, .5, .5, .5, 1.5), guide = "none") +
  scale_colour_manual("year", values = c("gray80", "gray60", "gray40", "gray20", "red")) +
  scale_alpha_manual("year", values = c(1, 1, 1, 1, 0.5), guide = "none") +
  # geom_hline(aes(yintercept=meanweight, colour=species)) +
  # geom_text(aes(x=1, y=meanweight, label=as.integer(meanweight)), hjust=0, vjust=0, nudge_y = 200) +
  facet_wrap(~species, scales="free_y")
  # facet_grid(rows=vars(year), cols=vars(species), scales="free_y")



```


```{r eval=FALSE, fig.asp=1.0, message=FALSE, warning=FALSE, comment=NA, include=FALSE}

# NOT USED

ee <- 
  e %>%
  filter(species %in% top$species) %>% 
  filter(year == max(e$year, na.rm=TRUE)) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  group_by(species, vessel, week) %>% 
  summarise(landingweight = sum(weight, na.rm=TRUE)) %>% 
  group_by(species, vessel) %>% 
  mutate(cumweight = cumsum(landingweight)) %>% 
  mutate(species = factor(species, levels=top$species))

# plot total catch and landings
ee %>% 
  ggplot() +
  theme_publication() +
  # theme(legend.position = "none") +
  geom_line(aes(x=week, y=cumweight, colour=vessel)) +
  facet_wrap(~species, scales="free_y")
  # facet_grid(rows=vars(year), cols=vars(species), scales="free_y")



```


```{r eval=FALSE, fig.asp=1.0, message=FALSE, warning=FALSE, comment=NA, include=FALSE}

# NOT USED 

# Vergelijking Marelec wegingen en Elog 

kk <- 
  k %>%
  filter(species %in% top$species) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  group_by(species, vessel, year, week) %>% 
  summarise(marelec_weight = sum(gewicht, na.rm=TRUE))  

ee <- 
  e %>%
  filter(year %in% kk$year) %>% 
  filter(species %in% top$species) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  group_by(species, vessel, year, week) %>% 
  summarise(elog_weight = sum(weight, na.rm=TRUE))  

t <-
  left_join(ee, kk, by=c("species","vessel", "year","week")) %>% 
  drop_na(marelec_weight, elog_weight) %>% 
  pivot_longer(names_to = "variable", values_to = "data", elog_weight:marelec_weight) %>% 
  mutate(variable = gsub("_weight","", variable)) %>% 
  arrange(variable, species, year, week) %>% 
  group_by(variable, species, year, week) %>% 
  summarise(data = sum(data)) %>% 
  group_by(variable, species, year) %>% 
  mutate(cumdata = cumsum(data))

# plot total catch and landings
  # mutate(year = as.character(year)) %>% 
  # mutate(species = factor(species, levels=top$species)) 

t %>%   
  ggplot() +
  theme_publication() +
  geom_line(aes(x=week, y=cumdata, colour=variable)) +
  # geom_hline(aes(yintercept=meanweight, colour=species)) +
  # geom_text(aes(x=1, y=meanweight, label=as.integer(meanweight)), hjust=0, vjust=0, nudge_y = 200) +
  facet_wrap(~species, scales="free_y")
  # facet_grid(rows=vars(year), cols=vars(species), scales="free_y")



```


```{r eval=FALSE, fig.asp=0.6, message=FALSE, warning=FALSE, comment=NA, include=FALSE}

# NOT SHOWN 
# SCH99 discard percentage per trip

hh <- 
  h %>%
  filter(vessel == "SCH99", !is.na(catchweight)) %>% 
  dplyr::select(vessel, trip, haul, date, catchweight) 

kk <- 
  k %>%
  filter(paste(vessel,trip,haul) %in% paste(hh$vessel, hh$trip, hh$haul)) %>% 
  mutate(date=as.Date(datetime)) %>% 
  group_by(vessel, trip, haul, date) %>% 
  summarise(landingweight = sum(gewicht, na.rm=TRUE)) 

comb <-
  left_join(hh, kk, by=c("vessel","trip","haul", "date")) %>% 
  drop_na(catchweight, landingweight) %>% 
  mutate(discardweight = catchweight - landingweight) %>% 
  mutate(discardperc = discardweight/catchweight) %>% 
  group_by(vessel, trip) %>% 
  mutate(trip2 = paste(min(date), max(date), sep="_"))

# plot total catch and landings
comb %>% 
  dplyr::select(-discardweight, -discardperc) %>% 
  pivot_longer(names_to = "variable", values_to = "data", catchweight:landingweight) %>% 
  ggplot() +
  theme_publication() +
  geom_bar(aes(x=haul, y=data, fill=variable), stat="identity", position=position_dodge2(preserve="single", padding=-0.25, width=1)) +
  facet_wrap(~trip)

# plot total catch and landings, with cleaned up hauls
comb %>% 
  filter(!(trip=="2023003" & haul %in% c(10,11,12,13))) %>% 
  filter(!(trip=="2023084" & haul %in% c(1,2,3,4))) %>% 
  filter(!(trip=="2023085" & haul %in% c(19,20,26,27,28))) %>% 
  filter(discardweight>0) %>% 
  dplyr::select(-discardweight, -discardperc) %>% 
  pivot_longer(names_to = "variable", values_to = "data", catchweight:landingweight) %>% 
  # dplyr::select(-catchweight, -discardperc) %>% 
  # pivot_longer(names_to = "variable", values_to = "data", landingweight:discardweight) %>% 
  # View()

  ggplot() +
  theme_publication() +
  geom_bar(aes(x=haul, y=data, fill=variable), stat="identity", position=position_dodge2(preserve="single")) +
  facet_wrap(~trip2)

# plot total catch and landings, with cleaned up hauls
comb %>% 
  filter(!(trip=="2023003" & haul %in% c(10,11,12,13))) %>% 
  filter(!(trip=="2023084" & haul %in% c(1,2,3,4))) %>% 
  filter(!(trip=="2023085" & haul %in% c(19,20,26,27,28))) %>% 
  filter(discardweight>0) %>% 
  dplyr::select(-discardweight, -landingweight, -catchweight) %>% 
  pivot_longer(names_to = "variable", values_to = "data", discardperc) %>% 

  ggplot() +
  theme_publication() +
  geom_bar(aes(x=haul, y=data, fill=variable), stat="identity", position=position_dodge2(preserve="single")) +
  scale_y_continuous(labels=scales::percent) +
  facet_wrap(~trip2)

# average percentage discards
comb %>% 
  filter(!(trip=="2023003" & haul %in% c(10,11,12,13))) %>% 
  filter(!(trip=="2023084" & haul %in% c(1,2,3,4))) %>% 
  filter(!(trip=="2023085" & haul %in% c(19,20,26,27,28))) %>% 
  filter(discardweight>0) %>% 
  group_by(vessel, trip, trip2) %>% 
  summarise(
    catchweight = sum(catchweight),
    discardweight = sum(discardweight)
  ) %>% 
  mutate(
    discardperc = discardweight / catchweight
  ) %>% 
  pivot_longer(names_to = "variable", values_to = "data", discardperc) %>% 

  ggplot() +
  theme_publication() +
  geom_bar(aes(x=trip2, y=data, fill=variable), stat="identity", position=position_dodge2(preserve="single")) +
  coord_flip() +
  scale_y_continuous(labels=scales::percent) 

```


```{r eval=FALSE, fig.asp=1.0, message=FALSE, warning=FALSE, comment=NA, include=FALSE}

h %>% 
  filter(year == 2023) %>% 
  mutate(source="jaczon") %>% 
  bind_rows(g %>% mutate(source="gfw")) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  mutate(quarter=paste0("Q", ceiling(month / 3))) %>% 
  
  ggplot(aes(x=lon, y=lat)) +
  theme_publication() +
  # coord_quickmap(xlim=c(-3,5) , ylim=c(49.5,53)) +
  coord_quickmap(xlim=c(-4,6) , ylim=c(49.5,55)) +
  geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +
  geom_point(aes(colour=source), size=0.5, alpha=0.2) +
  # geom_point(aes(colour=vesselname), size=0.5, alpha=0.5) +
  # facet_wrap(~month, ncol=3)
  # facet_wrap(~week, ncol=10)
  facet_grid(vessel ~ quarter)

```


```{r aanvoerendiscards, echo=FALSE, fig.align="center", fig.asp=0.9, message=FALSE, warning=FALSE}

print_option <- FALSE

# create list of catch by species by haul. From hm, or sph or else species = "oth"
if (nrow(k) > 0 & sum(!is.na(h$catchweight))>0) {

  print_option <- TRUE

  fig_nums(
    name    = "percdiscards", level   = 1, display = FALSE,
    caption = "Berekende percentage discards per dag")

  t1 <-
    haul %>%
    filter(catchweight>0, vessel=="SCH99", date >= dmy("23/10/2023")) %>% 
    filter(date <= max(h$date, na.rm=TRUE)) %>% 
    group_by(vessel, trip, haul) %>%
    summarise(data = sum(catchweight, na.rm=TRUE)) %>%
    mutate(variable = "totaal")

  t2 <-
    kisten %>%
    filter(paste(vessel, trip, haul) %in% paste(t1$vessel, t1$trip, t1$haul)) %>%
    group_by(vessel, trip, haul) %>%
    summarise(data = sum(gewicht, na.rm=TRUE)) %>%
    mutate(variable = "aanvoer")

  tt <-
    bind_rows(t1, t2) %>% 
    arrange(vessel, trip, haul) %>% 
    tidyr::pivot_wider(names_from = variable, values_from = data) %>% 
    drop_na(totaal) %>% 
    mutate(aanvoer = ifelse(is.na(aanvoer),0,aanvoer)) %>% 
    mutate(
      discards = totaal - aanvoer, 
      perc_discards = round((discards) / totaal, digits=2)
    ) %>% 
    # mutate(perc_discards = scales::percent(perc_discards, accuracy=1)) %>% 
    left_join(dplyr::select(haul,
                            vessel, trip, haul, year, week, date),
              by = c("vessel","trip","haul"))
  
  # tt %>% filter(perc_discards < 0 | perc_discards > 0.3) %>% View()
  # tt %>% filter(trip=="2024156" & (perc_discards < 0 | perc_discards > 0.3)) %>% View()
  # tt %>% filter(trip == "2024143", perc_discards > 0.3) %>% View()

  tt2 <-
    tt %>% 
    group_by(vessel, trip, year, week, date) %>% 
    summarise(
      aanvoer = sum(aanvoer, na.rm=TRUE),
      discards = sum(discards, na.rm=TRUE),
      totaal  = sum(totaal, na.rm=TRUE)
    ) %>% 
    mutate(aanvoer = ifelse(is.na(aanvoer),0,aanvoer)) %>% 
    mutate(perc_discards = round((totaal - aanvoer) / totaal, digits=2)) %>% 
    mutate(week2 = year + week/52) %>% 
    mutate(yday = lubridate::yday(date))
    
  # tt2 %>% filter(perc_discards > 0.3) %>% View()
  # tt2 %>% ggplot(aes(x=totaal, y=perc_discards)) + geom_point()
  
  # plot as one timeseries
  tt2 %>%
    ggplot(aes(x=date, y=perc_discards)) +
    theme_publication() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    geom_point(aes(size=totaal)) +
    geom_smooth() +
    tidyquant::geom_ma(n=5) + 
    scale_size(
      range=c(0.5,5),
      limits =  c(100,1000*ceiling(max(tt2$totaal)/1000)),
      breaks =  seq(100,1000*ceiling(max(tt2$totaal)/1000),length.out=6)
    ) +
    scale_x_date(date_breaks = "month") +
    scale_y_continuous(labels = scales::percent) +
    guides(size = guide_legend(nrow = 1)) +
    labs(x="", y="%discards / day", size= "Totale vangst (kg)") + 
    labs(title="Percentage discards per dag SCH99 Aravis")
  

  # plot by year
  tt2 %>%
    mutate(year = as.character(year)) %>% 
    ggplot(aes(x=week, y=perc_discards, group=year)) +
    theme_publication() +
    # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    
    geom_point(aes(size=totaal, colour=year)) +
    # geom_smooth(aes(colour=year, fill=year), span=0.9) +
    geom_smooth(aes(colour=year, fill=year), method = "lm", formula = y ~ poly(x, 3)) +
    # stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1)
    # tidyquant::geom_ma(n=5) + 
    scale_size(
      range=c(0.5,5),
      limits =  c(100,1000*ceiling(max(tt2$totaal)/1000)),
      breaks =  seq(100,1000*ceiling(max(tt2$totaal)/1000),length.out=6)
    ) +
    # scale_x_date(date_breaks = "month") +
    scale_y_continuous(labels = scales::percent) +
    guides(size = guide_legend(nrow = 1)) +
    labs(x="week", y="%discards / day", size= "Totale vangst (kg)") + 
    labs(title="Percentage discards per dag SCH99 Aravis")



}

my_caption   <- ifelse(print_option, paste0("_", fig_nums("aanvoerentotalevangst"),"_"), "")
print_option <- FALSE


```

<!-- ########################################################################################## -->
<!-- Figure: Landings per Species and per Haul (map)                                            -->
<!-- ########################################################################################## -->

```{r echo=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE}

# calculate the scaling of plots
xmin <- floor(2 * (min(k$lon, na.rm=TRUE)-0.5))/2;
xmax <- ceiling(2 * (max(k$lon, na.rm=TRUE)+0.5))/2;
ymin <- 49
ymax <- ceiling(2* (max(k$lat, na.rm=TRUE)+0.5))/2

xlim <- c(xmin, xmax); ylim <- c(ymin, ymax)

# create list of catch by species by haul. From hm, or sph or else species = "oth"
myspecies <- "COD"

for (myspecies in top$species) {
  
  t <-
    k %>%
    filter(species %in% myspecies) %>% 
    filter(year %in% 2023:2025) %>% 
    filter(!(lat==0 & lon == 0)) %>% 
    mutate(date = lubridate::as_date(datetime)) %>% 
    mutate(quarter = lubridate::quarter(date)) %>% 
    # mutate(quarter = paste0("q",quarter)) %>% 
    # group_by(species, vessel, year, month, date, rect, haul, lat, lon) %>%
    group_by(species, vessel, year, quarter, rect, date, haul, lat, lon) %>%
    summarise(gewicht = sum(gewicht, na.rm=TRUE)) %>% 
    filter(gewicht >= 20)
  

  tt <- scales::trans_breaks("sqrt", function(x) x^2)(c(0, max(t$gewicht)))
  
  p <-
    ggplot(data=t) +
  
    theme_publication() +
    theme(
      # panel.grid.major = element_blank(),
      # panel.grid.minor = element_blank(),
      text             = element_text(size=12),
      plot.margin      = unit(c(0,0,0,0), "cm"),
      # panel.background = element_rect(fill = "lightskyblue1")
    ) +
  
    coord_quickmap(xlim=xlim , ylim=ylim) +
  
    # geom_text(data=rect_df, aes(x=lon, y=lat, label=rect),
    #           hjust=0, vjust=0, nudge_x=0.05, nudge_y=0.025, colour="white") +
  
    geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +
  
    # geom_point(data=slice(tt, c(1,nrow(tt))),
    #           aes(x=lon, y=lat),
    #           colour="red", size=5) +
  
    geom_point(data=h %>% filter(year %in% unique(t$year)),
               aes(x=lon, y=lat, colour=as.character(year)),
               shape=3, size=0.5, stroke=0.1, alpha=0.4) +
    geom_point(aes(x=lon, y=lat, size = gewicht, colour = as.character(year)),
                alpha = 0.4, shape=21) +
    # scale_size(range = c(0.5, 5)) +
    scale_size_continuous(breaks=tt, range=c(0.1, 5)) +
    # scale_x_continuous(breaks = seq(floor(xmin), ceiling(xmax), 1)) +
    # scale_y_continuous(breaks = seq(floor(2*ymin)/2, ceiling(2*ymax)/2, 0.5)) +
    labs(x = NULL, y = NULL, size = "kg aanvoer/trek", colour="",
         title=paste(myspecies,"Landings by haul")) +
    guides(size = guide_legend(nrow = 1)) + 
    facet_grid(quarter~year)

  png(filename=file.path(onedrive, "report", "FLYSHOOT trips overview", 
                         paste0(myspecies," landings by haul.png")),
       width=7, height=9.8, units="in", res=300)

  print(p)
  
  dev.off()

}
```

```{r, echo=FALSE, fig.align="left", message=FALSE, warning=FALSE, cache=FALSE}
knitr::include_graphics(file.path(onedrive, "report", "FLYSHOOT trips overview", 
                         paste0("MAC"," landings by haul.png")))
```

```{r, echo=FALSE, fig.align="left", message=FALSE, warning=FALSE, cache=FALSE}
knitr::include_graphics(file.path(onedrive, "report", "FLYSHOOT trips overview", 
                         paste0("SQR"," landings by haul.png")))
```

```{r, echo=FALSE, fig.align="left", message=FALSE, warning=FALSE, cache=FALSE}
knitr::include_graphics(file.path(onedrive, "report", "FLYSHOOT trips overview", 
                         paste0("CTC"," landings by haul.png")))
```

```{r, echo=FALSE, fig.align="left", message=FALSE, warning=FALSE, cache=FALSE}
knitr::include_graphics(file.path(onedrive, "report", "FLYSHOOT trips overview", 
                         paste0("MUR"," landings by haul.png")))
```

```{r, echo=FALSE, fig.align="left", message=FALSE, warning=FALSE, cache=FALSE}
knitr::include_graphics(file.path(onedrive, "report", "FLYSHOOT trips overview", 
                         paste0("WHG"," landings by haul.png")))
```

```{r, echo=FALSE, fig.align="left", message=FALSE, warning=FALSE, cache=FALSE}
knitr::include_graphics(file.path(onedrive, "report", "FLYSHOOT trips overview", 
                         paste0("HOM"," landings by haul.png")))
```

<!-- ########################################################################################## -->
<!-- Figure: Mackerel by size category                                            -->
<!-- ########################################################################################## -->

```{r eval=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE, include=FALSE}

# NOT SHOWN

# create list of catch by species by haul. From hm, or sph or else species = "oth"
myspecies <- "MAC"

k %>%
  filter(species %in% myspecies) %>% 
  group_by(species, year, month, maat) %>%
  summarise(gewicht = sum(gewicht, na.rm=TRUE)) %>%

  ggplot() +

  theme_publication() +
  theme(
    # panel.grid.major = element_blank(),
    # panel.grid.minor = element_blank(),
    text             = element_text(size=12),
    # plot.margin      = unit(c(0,0,0,0), "cm"),
    # panel.background = element_rect(fill = "lightskyblue1")
  ) +
  geom_bar(aes(x=month, y=gewicht, fill=maat), stat="identity") +
  labs(x = NULL, y = NULL, title=paste(myspecies,"Landings by size category")) +
  guides(fill = guide_legend(nrow = 1)) + 
  scale_x_continuous(breaks=1:12) +
  facet_wrap(~year, ncol=2)

```

<!-- ########################################################################################## -->
<!-- Figure: catch per division                                         -->
<!-- ########################################################################################## -->

```{r eval=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE, include=FALSE}

# NOT SHOWN

# k %>%
#   filter(division %in% c("27.4.b","27.4.c","27.7.d", "27.7.e")) %>% 
#   group_by(species, year, week, division) %>%
#   summarise(
#     gewicht = sum(gewicht, na.rm=TRUE),
#     date    = as_date(min(datetime, na.rm=TRUE))
#   ) %>%
e %>%
  rename(division=faozone) %>% 
  filter(division %in% c("27.4.b","27.4.c","27.7.d", "27.7.e")) %>% 
  group_by(species, year, week, division) %>%
  summarise(
    gewicht = sum(weight, na.rm=TRUE),
    date    = as_date(min(catchdate, na.rm=TRUE))
  ) %>%

  ggplot() +

  theme_publication() +
  theme(
    # panel.grid.major = element_blank(),
    # panel.grid.minor = element_blank(),
    text             = element_text(size=12),
    # plot.margin      = unit(c(0,0,0,0), "cm"),
    # panel.background = element_rect(fill = "lightskyblue1")
  ) +
  geom_bar(aes(x=week, y=gewicht, fill=division), stat="identity") +
  labs(x = "week", y = "kg", title=paste("Total landings by division")) +
  guides(fill = guide_legend(nrow = 1)) + 
  # scale_x_continuous(breaks=1:53) +
  # facet_grid(year~division)
  facet_wrap(~year, ncol=1)

```


<!-- ########################################################################################## -->
<!-- Figure: Mackerel catch per division                                         -->
<!-- ########################################################################################## -->

```{r eval=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE, include=FALSE}

# NOT SHOWN

# create list of catch by species by haul. From hm, or sph or else species = "oth"
e %>%
  rename(division=faozone, gewicht=weight) %>% 
  filter(species %in% top$species[1:6]) %>% 
  filter(division %in% c("27.4.b","27.4.c","27.7.d","27.7.e")) %>% 
  group_by(species, year, month, division) %>%
  summarise(gewicht = sum(gewicht, na.rm=TRUE)) %>%

  ggplot() +

  theme_publication() +
  theme(text = element_text(size=12)) +
  geom_bar(aes(x=month, y=gewicht, fill=species), stat="identity") +
  labs(x = "week", y = "kg", title=paste("Landings by species")) +
  guides(fill = guide_legend(nrow = 1)) + 
  # scale_x_continuous(breaks=1:53) +
  facet_grid(year~division)

```


<!-- ########################################################################################## -->
<!-- Figure: total catch against mean                                                           -->
<!-- ########################################################################################## -->

```{r eval=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE, include=FALSE}

# NOT SHOWN

t1 <- 
  e %>%
  rename(division=faozone) %>% 
  filter(division %in% c("27.4.b","27.4.c","27.7.d", "27.7.e")) %>% 
  mutate(vessel = ifelse(vessel == "Z99", "SCH99", vessel)) %>% 
  mutate(catchdate = as_date(catchdate)) %>% 
  group_by(vessel, year, week) %>%
  summarise(
    gewicht = sum(weight, na.rm=TRUE),
    ndays   = n_distinct(catchdate)
  ) %>%
  group_by(vessel, year) %>% 
  mutate(
    cumweight = cumsum(gewicht),
    catch_day = gewicht / ndays)

# t1 %>% group_by(vessel, year) %>% summarise(catch = sum(gewicht, na.rm=TRUE), ndays=sum(ndays, na.rm=TRUE)) %>% View()

t2 <-
  t1 %>% 
  filter(year %notin% myyear) %>% 
  group_by(vessel, week) %>% 
  summarise(
    meancumweight = mean(cumweight, na.rm=TRUE),
    meancatch_day = mean(catch_day, na.rm=TRUE))

t3 <-
  left_join(t1, t2, by=c("vessel", "week")) %>% 
  mutate(
    dev1 = cumweight - meancumweight,
    dev2 = catch_day - meancatch_day
  )

t3 %>% 
  ggplot() +

  theme_publication() +
  geom_bar(aes(x=week, y=dev2, fill=vessel), stat="identity") +
  labs(x = "week", y = "kg", title=paste("Deviance")) +
  guides(fill = guide_legend(nrow = 1)) + 
  # scale_x_continuous(breaks=1:53) +
  # facet_grid(year~division)
  facet_grid(vessel~year)

```

