---
title: "JACZON FLYSHOOT MULTIVESSEL"
output:
  flexdashboard::flex_dashboard:
runtime: shiny
---



```{r global}

rm(list=ls())
options(dplyr.summarise.inform = FALSE)

library(rmarkdown)                   # note: requires knitr 1.21
require(tidyverse, quietly=TRUE)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
require(lubridate, quietly=TRUE)     # data handling
require(reshape2, quietly=TRUE)      # reshaping data; e.g. cast
require(RColorBrewer)                # colour schemes
library(viridis)
library(pander)

# Source all the utils
source("../R/FLYSHOOT utils.r")

spatialdir <- "C:/DATA/RDATA"
load(file.path(spatialdir, "world_lr_sf.RData"))
load(file.path(spatialdir, "world_mr_sf.RData"))
load(file.path(spatialdir, "world_lr_df.RData"))
load(file.path(spatialdir, "world_mr_df.RData"))

rect_df <-
  loadRData(file.path(spatialdir, "rect_df.RData")) %>% 
  rename(rect=ICESNAME) %>% 
  group_by(rect) %>% 
  filter(row_number() ==1) %>% 
  dplyr::select(rect, lon=long, lat)

asfis <-
  loadRData(file.path(spatialdir, "asfis.RData")) %>% 
  lowcase()


# set onedrive directory
onedrive <- get_onedrive(team="Martin Pastoors", site="FLYSHOOT - General/rdata")

load(file.path(onedrive, "haul.RData"))
trips <-
  haul %>% 
  group_by(vessel) %>% 
  distinct(vessel, date) %>% 
  mutate(diff = as.numeric(date-lag(date))) %>% 
  mutate(id   = ifelse(is.na(diff) | diff>1, 1, 0)) %>% 
  mutate(trip = cumsum(id)) %>% 
  group_by(vessel, trip) %>% 
  mutate(trip = paste0(trip," ",format(min(date), "%d/%m"), "-", format(max(date), "%d/%m"))) %>% 
  dplyr::select(-id, -diff)

haul <- haul %>% left_join(trips, by=c("vessel","date")) 

kisten <- loadRData(file.path(onedrive, "kisten.RData")) %>% 
  mutate(date=as.Date(datetime)) %>% 
  left_join(trips, by=c("vessel","date")) 
elog <- loadRData(file.path(onedrive, "elog.RData")) %>% 
  left_join(trips, by=c("vessel","date"))
elog_trek <- loadRData(file.path(onedrive, "elog_trek.RData")) %>% 
  left_join(trips, by=c("vessel","date"))
trip <- loadRData(file.path(onedrive, "trip.RData")) %>% 
  left_join(trips, by=c("vessel","date"))


maxdate <- max(haul$date, na.rm=TRUE)




```


Sidebar {.sidebar data-width=225}
=====================================

```{r}
# setvessel<-"SCH135"   ;settrip<-c("2023360","2023361") ; comparisons <- FALSE
# setvessel<-"SCH65"   ;settrip<-c("2023275") ; comparisons <- FALSE
# setvessel<-"SL9"   ;settrip<-c("2023081600041", "2023082100042") ; comparisons <- FALSE
# setvessel<-c("Z99", "SCH99")   ;settrip<- c("2023091") ; comparisons <- FALSE

vessels <- c("SCH65", "SCH99", "SCH135", "SL9")
shiny::selectInput(inputId = "setvessel", label = "Vessel:",
                   choices = vessels, selected = c("SL9", "SCH65"), 
                   multiple=TRUE)

dateRangeInput(inputId = "dateRange", 
               label   = "Date Select", 
               format  = "dd-mm-yyyy", 
               start   = maxdate-30, end=maxdate)

shiny::selectInput(inputId = "showby", label = "Show by:",
                   choices = c("by week","by vessel"), selected = "by week")


my_vars <- shiny::reactive({
    x =  c(ifelse(input$showby == "by week", c("week"), c("vessel")))
    })

my_showby <- shiny::reactive({
    x =  ifelse(input$showby == "by week", "week", "vessel") 
    })

# renderTable(input())

# h <- 
#   haul %>%
#   ungroup() %>% 
#   filter(vessel  %in% input$setvessel) %>%  
#   # filter(trip    %in% settrip) %>%  
#   ungroup()



```

Trial stuff - draft explanation will follow. 




Page 1
=====================================  

Column {data-width=600}
-------------------------------------

### Catch by day

```{r echo=FALSE}


renderPlot(
  ggplot(data=filter(elog, 
                     vessel == input$setvessel,
                     date >= input$dateRange[1],
                     date <= input$dateRange[2]) %>% 
           mutate(week = as.character(week)) %>% 
           group_by_at(vars(one_of(c(my_vars(), "date")))) %>%
           summarise(weight = sum(weight, na.rm=TRUE)), 
    aes(x=date, y=weight)) + 
    theme_publication() +
    {if(input$showby == "by week") {
      geom_bar(aes(fill=week), stat="identity")       
    } else {
      geom_bar(aes(fill=vessel), stat="identity") 
    }}+
    labs(x="") +
    guides(fill = guide_legend(nrow = 1))
  )


```



Column {data-width=400}
-------------------------------------

### Species composition

```{r echo=FALSE}

# select species
s = shiny::reactive({
    x =  filter(elog, 
                vessel == input$setvessel,
                date >= input$dateRange[1],
                date <= input$dateRange[2]) %>% 
         group_by_at(vars(one_of(c("vessel", "species")))) %>% 
         summarise(weight = sum(weight, na.rm=TRUE)) %>% 
         
         arrange(desc(weight)) %>% 
         filter(row_number()<=10) %>%
         mutate(species = forcats::fct_reorder(species, weight)) 
    })

         
e = shiny::reactive({
    x =  filter(elog,
                vessel == input$setvessel,
                date >= input$dateRange[1],
                date <= input$dateRange[2]) %>%

         mutate(species = ifelse(species %in% s()$species, species, "ZZZ")) %>%
         mutate(species = factor(species, levels=c(as.character(s()$species),"ZZZ"))) %>%
         group_by_at(vars(one_of(c(my_vars(), "species")))) %>%
         summarise(weight = sum(weight, na.rm=TRUE)) %>%

         group_by_at(vars(one_of(c(my_vars())))) %>%
         mutate(perc = weight / sum(weight)) %>%

         mutate(label_pos = cumsum(weight) - weight / 2,
                perc_text = paste0(round(perc * 100), "%"))

    })
   
# shiny::reactive(writexl::write_xlsx(e(), path="e.xlsx"))
 

# renderTable(head(e()))
# renderTable(head(e(), n=20))

renderPlot(
  ggplot(data=e(),
         aes_string(x=my_showby(), y="weight", fill="species")) +
    theme_publication() +
    geom_bar(stat="identity", width=1) +
    # geom_text(aes(y = label_pos, label = ifelse(perc > 0.1, weight,"")), size = 4) +
    # geom_text(aes(y = label_pos, label = species), size = 4) +
    geom_text(aes(label=ifelse(perc>=0.1,paste(as.character(species)),"")), 
              size=4, position=position_stack(vjust=0.5)) +
    # coord_polar("y", start=0) +
    guides(fill = guide_legend(nrow = 1)) +

    labs(title=input$showby) 

  )

```


### Map

```{r echo=FALSE}

t = shiny::reactive({
    x =  filter(haul, vessel == input$setvessel,
                 date >= input$dateRange[1],
                 date <= input$dateRange[2])} %>% 
           mutate(week = as.character(week), trip = as.character(trip)) 
    ) 

xlim <- shiny::reactive({x=c(floor(min(t()$lon, na.rm=TRUE)), 
                             ceiling(max(t()$lon, na.rm=TRUE)))})
ylim <- shiny::reactive({x=c(floor(min(t()$lat, na.rm=TRUE)), 
                             ceiling(max(t()$lat, na.rm=TRUE)))})

renderPlot(
  ggplot(data=t(), 
         aes(x=lon, y=lat)) + 
    theme_publication() +
    coord_quickmap(xlim=xlim() , ylim=ylim()) +
    geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +

    {if(input$showby == "by week") {
      geom_point(aes(colour=week), size=1)       
    } else {
      geom_point(aes(colour=trip), size=1) 
    }}+
    guides(colour = guide_legend(nrow = 1))
  )



```


Page 2
=====================================

Column {.tabset}
-------------------------------------

### Catch per tow

```{r echo=FALSE}

t = shiny::reactive({
    x =  filter(haul, vessel == input$setvessel,
                 date >= input$dateRange[1],
                 date <= input$dateRange[2])} %>% 
           mutate(week = as.character(week), trip = as.character(trip)) 
    ) 

xlim <- shiny::reactive({x=c(floor(min(t()$lon, na.rm=TRUE)), 
                             ceiling(max(t()$lon, na.rm=TRUE)))})
ylim <- shiny::reactive({x=c(floor(min(t()$lat, na.rm=TRUE)), 
                             ceiling(max(t()$lat, na.rm=TRUE)))})

renderPlot(
  ggplot(data=t(), 
         aes(x=lon, y=lat)) + 
    theme_publication() +
    theme(legend.position="none") +
    coord_quickmap(xlim=xlim() , ylim=ylim()) +
    geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +

    {if(input$showby == "by week") {
      geom_point(aes(colour=week), size=1)       
    } else {
      geom_point(aes(colour=trip), size=1) 
    }}+
    
    guides(colour = guide_legend(nrow = 1)) +
    
    labs(title=input$showby) +
    
    {if(input$showby == "by week") {
      facet_wrap(~week)       
    } else {
      facet_wrap(~trip)       
    }}
    
  )



```

### Catch compositions

```{r echo=FALSE}

my_vars <- shiny::reactive({
    x =  c("vessel", ifelse(input$showby == "by week", c("week"), c("trip")))
    })

# renderText(my_vars())

my_showby <- shiny::reactive({
    x =  ifelse(input$showby == "by week", "week", "trip") 
    })

# select species
s = shiny::reactive({
    x =  filter(elog, 
                vessel == input$setvessel,
                date >= input$dateRange[1],
                date <= input$dateRange[2]) %>% 
         group_by_at(vars(one_of(c("vessel", "species")))) %>% 
         summarise(weight = sum(weight, na.rm=TRUE)) %>% 
         
         arrange(desc(weight)) %>% 
         filter(row_number()<=10) %>%
         mutate(species = forcats::fct_reorder(species, weight)) 
    })

         
e = shiny::reactive({
    x =  filter(elog,
                vessel == input$setvessel,
                date >= input$dateRange[1],
                date <= input$dateRange[2]) %>%

         mutate(species = ifelse(species %in% s()$species, species, "ZZZ")) %>%
         mutate(species = factor(species, levels=c(as.character(s()$species),"ZZZ"))) %>%
         group_by_at(vars(one_of(c(my_vars(), "species")))) %>%
         summarise(weight = sum(weight, na.rm=TRUE)) %>%

         group_by_at(vars(one_of(c(my_vars())))) %>%
         mutate(perc = weight / sum(weight)) %>%

         mutate(label_pos = cumsum(weight) - weight / 2,
                perc_text = paste0(round(perc * 100), "%"))

    })
   
# shiny::reactive(writexl::write_xlsx(e(), path="e.xlsx"))
 

# renderTable(head(e()))
# renderTable(head(e(), n=20))

renderPlot(
  ggplot(data=e(),
         aes_string(x=my_showby(), y="weight", fill="species")) +
    theme_publication() +
    geom_bar(stat="identity", width=1) +
    # geom_text(aes(y = label_pos, label = ifelse(perc > 0.1, weight,"")), size = 4) +
    # geom_text(aes(y = label_pos, label = species), size = 4) +
    geom_text(aes(label=ifelse(perc>0.05,paste(as.character(species), as.integer(weight)),"")), 
              size=4, position=position_stack(vjust=0.5)) +
    # coord_polar("y", start=0) +
    guides(fill = guide_legend(nrow = 1)) +

    labs(title=input$showby) 

  )


```



### Catch proportions

```{r echo=FALSE}

my_vars <- shiny::reactive({
    x =  c("vessel", ifelse(input$showby == "by week", c("week"), c("trip")))
    })

# renderText(my_vars())

my_showby <- shiny::reactive({
    x =  ifelse(input$showby == "by week", "week", "trip") 
    })

# select species
s = shiny::reactive({
    x =  filter(elog, 
                vessel == input$setvessel,
                date >= input$dateRange[1],
                date <= input$dateRange[2]) %>% 
         group_by_at(vars(one_of(c("vessel", "species")))) %>% 
         summarise(weight = sum(weight, na.rm=TRUE)) %>% 
         
         arrange(desc(weight)) %>% 
         filter(row_number()<=10) %>%
         mutate(species = forcats::fct_reorder(species, weight)) 
    })

         
e = shiny::reactive({
    x =  filter(elog,
                vessel == input$setvessel,
                date >= input$dateRange[1],
                date <= input$dateRange[2]) %>%

         mutate(species = ifelse(species %in% s()$species, species, "ZZZ")) %>%
         mutate(species = factor(species, levels=c(as.character(s()$species),"ZZZ"))) %>%
         group_by_at(vars(one_of(c(my_vars(), "species")))) %>%
         summarise(weight = sum(weight, na.rm=TRUE)) %>%

         group_by_at(vars(one_of(c(my_vars())))) %>%
         mutate(perc = weight / sum(weight)) %>%

         mutate(label_pos = cumsum(weight) - weight / 2,
                perc_text = paste0(round(perc * 100), "%"))

    })
   
# shiny::reactive(writexl::write_xlsx(e(), path="e.xlsx"))
 

# renderTable(head(e()))
# renderTable(head(e(), n=20))

renderPlot(
  ggplot(data=e(),
         aes_string(x=my_showby(), y="weight", fill="species")) +
    theme_publication() +
    geom_bar(stat="identity", width=1, position = position_fill()) +
    # geom_text(aes(y = label_pos, label = ifelse(perc > 0.1, weight,"")), size = 4) +
    # geom_text(aes(y = label_pos, label = species), size = 4) +
    geom_text(aes(label=ifelse(perc>0.02,paste(as.character(species), perc_text),"")), 
              size=4, position=position_fill(vjust=0.5)) +
    # coord_polar("y", start=0) +
    guides(fill = guide_legend(nrow = 1)) +

    labs(title=input$showby) 

  )


```


Page 3
===================================== 

Column {.tabset}
-------------------------------------

### The data source

***

Something nice about the program

- ...

- ...

- ...

- ...

### The time panel

***

Here possibly a snapshot of an illustrative species, explaining how to read the graph

### The space panel


***

Here possibly a snapshot of an illustrative species, explaining how to read the graph

### Method


***

Very brief points on methods with then link to more detailed documentation