---
output:
  word_document:
    reference_docx: ../report_template.dotx
---


```{r setup, include=FALSE}

# ==========================================================================================================
# Generate standard FLYSHOOT tripreport
# 
# 11/01/2023 first coding
# 20/01/2023 v1 with inclusion of elog data
# 23/02/2023 added multiple trips possibility
# 19/08/2023 PEFA only
# ==========================================================================================================

# Reset lists
rm(list=ls())


knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

options(dplyr.summarise.inform = FALSE)


# Libraries
library(rmarkdown)                   # note: requires knitr 1.21

require(tidyverse, quietly=TRUE)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
require(lubridate, quietly=TRUE)     # data handling
require(reshape2, quietly=TRUE)      # reshaping data; e.g. cast
require(RColorBrewer)                # colour schemes
library(viridis)
library(pander)

library(captioner)                   # for captions
tab_nums <- captioner::captioner(prefix = "Table " , levels=1, type=c("n"), infix=".")
fig_nums <- captioner::captioner(prefix = "Figure ", levels=1, type=c("n"), infix=".")

# Source all the utils
source("../r/FLYSHOOT utils.R")

spatialdir = "C:/Users/MartinPastoors/OneDrive - Martin Pastoors/DATA/RDATA"

# spatial data files

load(file.path(spatialdir, "world_lr_sf.RData"))
load(file.path(spatialdir, "world_mr_sf.RData"))
load(file.path(spatialdir, "eez.df.RData"))

# load(file.path(onedrive, "fao.df.RData"))
# load(file.path(onedrive, "depth200.df.RData"))
# load(file.path(onedrive, "icesrectangles.df.RData"))

load(file.path(spatialdir, "world_lr_df.RData"))
load(file.path(spatialdir, "world_mr_df.RData"))

rect_df <-
  loadRData(file.path(spatialdir, "rect_df.RData")) %>% 
  rename(rect=ICESNAME) %>% 
  group_by(rect) %>% 
  filter(row_number() ==1) %>% 
  dplyr::select(rect, lon=long, lat)

asfis <-
  loadRData(file.path(spatialdir, "asfis.RData")) %>% 
  lowcase()


# ===============================================================================
#  Load and filter the trip data
# ===============================================================================

# set onedrive directory
onedrive <- get_onedrive(team="Martin Pastoors", site="FLYSHOOT - General/rdata")

# load datasets
load(file.path(onedrive, "elog.RData"))
load(file.path(onedrive, "elog_trek.RData"))

# elog %>% filter(vessel=="SL9", year==2023) %>% distinct(trip) %>% View()


# ===============================================================================
# ===============================================================================
# Select vessel and trip
# ===============================================================================
# ===============================================================================

# setvessel<-"SCH135"
# setvessel<-"SCH65"
setvessel<-"SL9"

startdate<-dmy("4/11/2025"); enddate <- dmy("11/11/2025") ;

# ===============================================================================
# ===============================================================================

settrip  <- 
  elog %>% 
  filter(vessel %in% setvessel, catchdate >= startdate, catchdate <= enddate) %>% 
  distinct(trip) %>% 
  .$trip


e <-
  elog %>% 
  filter(vessel  %in% setvessel,
         trip    %in% settrip) 

et <-
  elog_trek %>% 
  filter(vessel  %in% setvessel,
         trip    %in% settrip) %>% 
  
  # fix both lon and lat are zero
  mutate(lon = ifelse(lon==0 & lat==0, NA, lon)) %>% 
  mutate(lat = ifelse(is.na(lon) & lat==0, NA, lat)) 


if(nrow(et) > 0) {
  xmin <- floor(min(et$lon, na.rm=TRUE))
  xmax <- ceiling(max(et$lon, na.rm=TRUE))
  ymin <- floor(2*min(et$lat, na.rm=TRUE))/2
  ymax <- ceiling(2*max(et$lat, na.rm=TRUE))/2
} else {
  xmin <- floor(min(e$lon, na.rm=TRUE))
  xmax <- ceiling(max(e$lon, na.rm=TRUE))+1
  ymin <- floor(2*min(e$lat, na.rm=TRUE))/2
  ymax <- ceiling(2*max(e$lat, na.rm=TRUE))/2 + 0.5
}

# xmin <- floor(2 * (min(h$lon, na.rm=TRUE)-0.5))/2;
# xmax <- ceiling(2 * (max(h$lon, na.rm=TRUE)+0.5))/2;
# ymin <- floor(2 * (min(h$lat, na.rm=TRUE)-0.5))/2;
# ymax <- ceiling(2* (max(h$lat, na.rm=TRUE)+0.5))/2

# xdistance <- geosphere::distGeo(matrix(c(xmin, ymin), ncol=2, byrow=FALSE),
#                                 matrix(c(xmax, ymin) , ncol=2, byrow=FALSE) )/1000
# ydistance <- geosphere::distGeo(matrix(c(xmin, ymin), ncol=2, byrow=FALSE),
#                                 matrix(c(xmin, ymax) , ncol=2, byrow=FALSE) )/1000
 
xdistance <- icosa::arcdist(p1=c(xmin,ymin), p2=c(xmax,ymin))
ydistance <- icosa::arcdist(p1=c(xmin,ymin), p2=c(xmin,ymax))
# xdistance <- geo::arcdist(lat=ymin, lon=xmin, lat1 = ymin, lon1 = xmax, scale = "km")
# ydistance <- geo::arcdist(lat=ymin, lon=xmin, lat1 = ymax, lon1 = xmin, scale = "km")
aspect    <- ydistance/xdistance

if(aspect > 3) {
  xmin = xmin -2; xmax = xmax +2
} else if (aspect >2) {
  xmin = xmin -1; xmax = xmax +1
} else if (aspect < 0.5) {
  ymin = ymin -0.5; ymax = ymax + 0.5
} else if (aspect >0.3) {
  ymin = ymin -0.5; ymax = ymax + 0.5
}

# elog %>% 
#   filter(vessel == "SL9", week == 22, year==2025, date==lubridate::dmy("1/6/2025")) %>% 
#   # group_by(date, species) %>% 
#   # mutate(ntrips = n_distinct(trip)) %>% 
#   # group_by(trip, date) %>%
#   # summarise(catch = sum(weight, na.rm=TRUE)) %>%
#   View()

```


# Vessel tripreport `r unique(e$vessel)`, trip `r as.character(unique(e$trip))`

### Trip start: `r format(min(e$date), '%d/%m/%Y')`, end: `r format(max(e$date), '%d/%m/%Y')`

Report generated: `r format(Sys.time(), '%d/%m/%Y')`

&nbsp;

<!-- ########################################################################################## -->
<!-- Table: samenvatting reis                                       --------------------------- -->
<!-- ########################################################################################## -->

```{r samenvatting, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}

tab_nums(
  name    = "samenvatting", 
  level = 1, display = FALSE,
  caption = "Samenvatting van de reis")

# summarize the haul information
if (sum(!is.na(et$weight)) >0 ) {
  
  c1 <- 
    et %>%
    group_by(vessel,trip) %>%
    summarise(
      nhauls            = n_distinct(haul),
      startdate         = min(date, na.rm=TRUE),
      enddate           = max(date, na.rm=TRUE),
      weeks             = paste(unique(week), collapse=", "),
      months            = paste(unique(month), collapse=", "),
      divisions         = paste(unique(na.omit(faozone)), collapse="; "),
      rects             = paste(unique(na.omit(rect)), collapse="; "),
    ) %>% 
    mutate (
      ndays             = as.numeric(difftime(enddate, startdate, units="days")) + 1,
      haulsperday       = nhauls/ndays
    ) 
  
} else if (sum(!is.na(e$weight)) >0 ) {
  
  cat("Note: No scale/haul information available")
  
  c1 <- 
    e %>%
    group_by(vessel,trip) %>%
    summarise(
      startdate         = min(date, na.rm=TRUE),
      enddate           = max(date, na.rm=TRUE),
      weeks             = paste(unique(week), collapse=", "),
      months            = paste(unique(month), collapse=", "),
      divisions         = paste(unique(na.omit(faozone)), collapse="; "),
      rects             = paste(unique(na.omit(rect)), collapse="; "),
    ) %>% 
    mutate (
      ndays             = as.numeric(difftime(enddate, startdate, units="days")) + 1,
      nhauls            = as.character(NA)
    ) 
  
} else {

    cat("No elog information available")
  
}


# summarize the catch data from elog and elog by trek data
c2 <- 
  bind_rows(
    e %>%
      group_by(vessel,trip) %>%
      summarise(aanvoer = sum(weight, na.rm=TRUE)) %>% 
      mutate(source="elog"),  
    et %>%
      group_by(vessel,trip) %>%
      summarise(aanvoer = sum(weight, na.rm=TRUE)) %>% 
      mutate(source="elog per trek")
  )

# species
s <-
  e %>% 
    group_by(species) %>% 
    summarise(gewicht = sum(weight, na.rm=TRUE)) %>% 
    arrange (desc(gewicht)) %>% 
    slice_head(n=8) %>% 
    distinct(species)

  
# soorten en presentatievorm
c3 <-
  bind_rows(
    # by species
    e %>% 
      filter(species %in% s$species) %>% 
      mutate(gewicht = as.numeric(ifelse(is.na(weight),0,weight)) + as.numeric(ifelse(is.na(weightundersized), 0, weightundersized))) %>% 
      group_by(trip, species, presentation) %>% 
      summarise(gewicht = sum(gewicht, na.rm=TRUE)),  
      # arrange (desc(gewicht)) %>% 
      # slice_head(n=8),
    #other
    e %>% 
      filter(species %notin% s$species) %>% 
      # group_by(trip, species) %>% 
      # summarise(gewicht = sum(weight, na.rm=TRUE)) %>% 
      # arrange (desc(gewicht)) %>% 
      # filter(row_number() > 8) %>% 
      mutate(gewicht = as.numeric(ifelse(is.na(weight),0,weight)) + as.numeric(ifelse(is.na(weightundersized), 0, weightundersized))) %>% 
      group_by(trip, presentation) %>% 
      summarise(gewicht = sum(gewicht)) %>% 
      mutate(species="OTHER"),
    # total
    # e %>% 
    #   group_by(trip) %>% 
    #   summarise(gewicht = sum(weight, na.rm=TRUE)) %>% 
    #   mutate(species="TOTAL")
  ) %>% 
  mutate(species = factor(species, levels=c(s$species,"OTHER","TOTAL")))

# discard reasons
c4 <-
  bind_rows(
    e %>% 
      filter(presentation == "DIS") %>% 
      filter(species %in% s$species) %>% 
      mutate(gewicht = as.numeric(ifelse(is.na(weight),0,weight)) + as.numeric(ifelse(is.na(weightundersized), 0, weightundersized))) %>% 
      group_by(trip, species, discardreason) %>% 
      summarise(gewicht = sum(gewicht, na.rm=TRUE)), 
      # arrange (desc(gewicht)) %>% 
      # slice_head(n=8),
    e %>% 
      filter(presentation == "DIS") %>% 
      filter(species %notin% s$species) %>% 
      mutate(gewicht = as.numeric(ifelse(is.na(weight),0,weight)) + as.numeric(ifelse(is.na(weightundersized), 0, weightundersized))) %>% 
      # group_by(trip, species) %>% 
      # summarise(gewicht = sum(weight, na.rm=TRUE)) %>% 
      # arrange (desc(gewicht)) %>% 
      # filter(row_number() > 8) %>% 
      group_by(trip, discardreason) %>% 
      summarise(gewicht = sum(gewicht)) %>% 
      mutate(species="OTHER"),
    # et %>% 
    #   group_by(trip) %>% 
    #   summarise(gewicht = sum(weight, na.rm=TRUE)) %>% 
    #   mutate(species="TOTAL")
  ) %>% 
  mutate(species = factor(species, levels=c(s$species,"OTHER","TOTAL")))

if(nrow(c1) > 0) {
  c1 %>% 
    t() %>% 
    no.emphasis.table() %>% 
    pandoc.table(.,
               style = "simple",
               split.tables=400, 
               justify = "left",
               missing=".",
               round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
}

if(nrow(c2) > 0) {
  c2 %>% 
    ungroup() %>% 
    dplyr::select(-vessel) %>% 
    reshape2::dcast(trip ~ source, value.var = "aanvoer", sum, margins="trip") %>% 
    # mutate(vangst = "aanvoer") %>% 
    # pivot_wider(names_from = source, values_from = aanvoer) %>% 
    pandoc.table(.,
               style = "simple",
               split.tables=400, 
               justify = c("left",rep("right",length(unique(c2$source)))),
               missing=".",
               round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
}

if(nrow(c3) > 0) {
  cat("Catch and presentation (elog)")
  
  c3 %>% 
    ungroup() %>% 
    reshape2::dcast(trip+presentation ~ species, value.var = "gewicht", sum, margins=c("trip", "presentation", "species")) %>% 
    group_by(trip) %>% 
    do(add_row(., .after=0)) %>%
    ungroup() %>% 
    filter(row_number() > 1) %>%
    pandoc.table(.,
               style = "simple",
               split.tables=400, 
               justify = "left",
               missing=" ",
               round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
  
}

if (nrow(c4) == 0 ) {
    
    cat(" ")  
    cat("(No discards reported)") 

  } else {
  
    cat("Discard reasons")
    
    c4 %>% 
      reshape2::dcast(trip+discardreason ~ species, value.var = "gewicht", sum, margins=c("trip", "discardreason", "species")) %>% 
      group_by(trip) %>% 
      do(add_row(., .after=0)) %>%
      ungroup() %>% 
      filter(row_number() > 1) %>% 
      pandoc.table(.,
                 style = "simple",
                 split.tables=400, 
                 justify = "left",
                 missing=" ",
                 round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
  
}  

```

_`r tab_nums("samenvatting")`_


<!-- ########################################################################################## -->
<!-- Table: vangst --------------------------- -->
<!-- ########################################################################################## -->

\newpage

**Aanvoer per soort (elog, kg)**

```{r vangstpersoort, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}

tab_nums(
  name    = "vangstpersoort", 
  level = 1, display = FALSE,
  caption = "Vangst per soort (kg)")

t1 <-
  e %>% 
  dplyr::group_by(trip, species) %>% 
  dplyr::summarize(gewicht     = sum(weight, na.rm=TRUE))  

t1 %>% 
  reshape2::dcast(species ~ trip, value.var="gewicht", sum, margins = c("species","trip")) %>% 
  # mutate(soorten = tolower(species)) %>% 

  pander::pandoc.table(.,
               style = "simple",
               split.tables=400, 
               justify = "right",
               missing=".",
               round=c(0))
  

```

_`r tab_nums("vangstpersoort")`_

**Aanvoer per soort (elog, percentage)**

```{r vangstpersoort2, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}

tab_nums(
  name    = "vangstpersoort2", 
  level = 1, display = FALSE,
  caption = "Vangst per soort (percentage)")

t1 <-
  e %>% 
  dplyr::group_by(trip, species) %>% 
  dplyr::summarize(gewicht     = sum(weight, na.rm=TRUE)) %>% 
  dplyr::group_by(trip) %>% 
  dplyr::mutate(perc = gewicht/sum(gewicht, na.rm=TRUE))  %>% 
  dplyr::select(-gewicht)

t1 %>% 
  reshape2::dcast(species ~ trip, value.var="perc", sum, margins = c("species")) %>% 
  mutate(across(!grep("species", names(.)),~scales::percent(., accuracy=1))) %>% 

  pander::pandoc.table(.,
               style = "simple",
               split.tables=400, 
               justify = "right",
               missing=".",
               round=c(0))
  

```

\newpage

**Aanvoer per soort en dag (elog)**

```{r vangstpersoortendag, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}

tab_nums(
  name    = "vangstpersoortendag", 
  level = 1, display = FALSE,
  caption = "Vangst per soort (kg) per dag")

e %>% 
  # mutate(date = as.Date(datetime)) %>% 
  mutate(species = tolower(species)) %>% 
  dplyr::group_by(date, species) %>% 
  dplyr::summarize(gewicht     = sum(weight, na.rm=TRUE)) %>% 
  dplyr::mutate(date = format(date, "%d/%m")) %>% 
  reshape2::dcast(species ~ date, value.var = "gewicht", sum, margins = c("date", "species")) %>% 
  pander::pandoc.table(.,
               style = "simple",
               split.tables=400, 
               justify = "right",
               missing=".",
               round=c(0))
  

```

_`r tab_nums("vangstpersoortendag")`_

\newpage

<!-- ########################################################################################## -->
<!-- Figure: Total Landings per Haul or rectangle(map)                                                         -->
<!-- ########################################################################################## -->

```{r catchperhaul, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "catchperhaul", level   = 1, display = FALSE,
  caption = "Vangst (aanvoer) per trek (kg)")

xlim <- c(xmin,xmax); ylim <- c(ymin,ymax)


if (sum(!is.na(et$weight)) >0 ) {

  et %>%
    
    group_by(vessel, trip, haul, lat, lon, date) %>%
    summarise(gewicht = sum(weight, na.rm=TRUE)) %>%

    ggplot() +

    theme_publication() +
    theme(
      # panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      text             = element_text(size=12),
      plot.margin      = unit(c(0,0,0,0), "cm"),
      panel.background = element_rect(fill = "lightskyblue1")
    ) +

    coord_quickmap(xlim=xlim , ylim=ylim) +

    geom_text(data=rect_df, aes(x=lon, y=lat, label=rect),
              hjust=0, vjust=0, nudge_x=0.05, nudge_y=0.025, colour="white") +

    geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +
    geom_polygon(data=eez.df, aes(x=long, y=lat, group=group), fill = NA, colour="gray30") +

    geom_path(aes(x=lon, y=lat, colour=trip)) +
    # geom_point(data=slice(tt, c(1,nrow(tt))),
    #           aes(x=lon, y=lat),
    #           colour="red", size=5) +

    geom_point(aes(x=lon, y=lat, size = gewicht, colour = as.character(trip)),
                alpha = 0.5) +
    scale_size(range = c(1,10)) +

    scale_x_continuous(breaks = seq(floor(xmin), ceiling(xmax), 1)) +
    scale_y_continuous(breaks = seq(floor(2*ymin)/2, ceiling(2*ymax)/2, 0.5)) +
    labs(x = NULL, y = NULL, size = "kg aanvoer/trek", colour="") +
    guides(size = guide_legend(nrow = 1))
    # + facet_wrap(~source)

} else if (sum(!is.na(e$weight)) >0 ) {
  
  e %>%
    group_by(vessel, trip, haul, lat, lon, date) %>%
    summarise(gewicht = sum(weight, na.rm=TRUE)) %>%
    mutate(
      lon = lon + 0.5,
      lat = lat + 0.25
    ) %>% 

    ggplot() +

    theme_publication() +
    theme(
      # panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      text             = element_text(size=12),
      plot.margin      = unit(c(0,0,0,0), "cm"),
      panel.background = element_rect(fill = "lightskyblue1")
    ) +

    coord_quickmap(xlim=xlim , ylim=ylim) +
    # coord_quickmap(xlim=c(0,5) , ylim=c(52,55)) +

    geom_text(data=rect_df, aes(x=lon, y=lat, label=rect),
              hjust=0, vjust=0, nudge_x=0.05, nudge_y=0.025, colour="white") +

    geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +
    geom_polygon(data=eez.df, aes(x=long, y=lat, group=group), fill = NA, colour="gray30") +

    geom_jitter(aes(x=lon, y=lat, size = gewicht, colour = as.character(trip)),
                width=0.1, height=0.05, alpha = 0.5) +
    scale_size(range = c(1,10)) +

    scale_x_continuous(breaks = seq(floor(xmin), ceiling(xmax), 1)) +
    scale_y_continuous(breaks = seq(floor(2*ymin)/2, ceiling(2*ymax)/2, 0.5)) +
    labs(x = NULL, y = NULL, size = "kg aanvoer/dag/ICES vak", colour="") +
    guides(size = guide_legend(nrow = 1))
    # + facet_wrap(~source)
  
} else {
  
  cat("No lat - lon information available")
  
}

  

```


<!-- ########################################################################################## -->
<!-- Figure: Landings per Species and per Haul (map)                                                         -->
<!-- ########################################################################################## -->

```{r catchperhaul2, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "catchperhaul2", level   = 1, display = FALSE,
  caption = "Vangst (aanvoer) per soort en trek (kg)")

xlim <- c(xmin,xmax); ylim <- c(ymin,ymax)

nspecies=12


if (sum(!is.na(et$weight)) >0 ) {
  
  # create list of catch by species by haul. From hm, or sph or else species = "oth"
  myspecies <-
    et %>% 
    group_by(species) %>% 
    summarise(gewicht = sum(weight, na.rm=TRUE)) %>% 
    arrange (desc(gewicht)) %>% 
    slice_head(n=nspecies)

  et %>%
    filter(species %in% myspecies$species) %>% 
    group_by(vessel, trip, haul, lat, lon, species) %>%
    summarise(gewicht = sum(weight, na.rm=TRUE)) %>%
  
    ggplot() +
  
    theme_publication() +
    theme(
      # panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      text             = element_text(size=12),
      plot.margin      = unit(c(0,0,0,0), "cm"),
      panel.background = element_rect(fill = "lightskyblue1")
    ) +
  
    coord_quickmap(xlim=xlim , ylim=ylim) +
  
    # geom_text(data=rect_df, aes(x=lon, y=lat, label=rect),
    #           hjust=0, vjust=0, nudge_x=0.05, nudge_y=0.025, colour="white") +
  
    geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +
  
    geom_path(aes(x=lon, y=lat, colour=trip)) +
    # geom_point(data=slice(tt, c(1,nrow(tt))),
    #           aes(x=lon, y=lat),
    #           colour="red", size=5) +
  
    geom_point(aes(x=lon, y=lat, size = gewicht, colour = as.character(trip)),
                alpha = 0.5) +
    scale_size(range = c(1,10)) +
  
    scale_x_continuous(breaks = seq(floor(xmin), ceiling(xmax), 1)) +
    scale_y_continuous(breaks = seq(floor(2*ymin)/2, ceiling(2*ymax)/2, 0.5)) +
    labs(x = NULL, y = NULL, size = "kg aanvoer/trek", colour="") +
    guides(size = guide_legend(nrow = 1)) + 
    facet_wrap(~species, ncol=3)
  
} else if (sum(!is.na(e$weight)) >0 ) {
  
  # create list of catch by species by haul. From hm, or sph or else species = "oth"
  myspecies <-
    e %>% 
    group_by(species) %>% 
    summarise(gewicht = sum(weight, na.rm=TRUE)) %>% 
    arrange (desc(gewicht)) %>% 
    slice_head(n=nspecies)
  
  e %>%
    filter(species %in% myspecies$species) %>% 
    group_by(vessel, trip, haul, lat, lon, species) %>%
    summarise(gewicht = sum(weight, na.rm=TRUE)) %>%
    mutate(
      lon = lon + 0.5,
      lat = lat + 0.25
    ) %>% 
  
    ggplot() +
  
    theme_publication() +
    theme(
      # panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      text             = element_text(size=12),
      plot.margin      = unit(c(0,0,0,0), "cm"),
      panel.background = element_rect(fill = "lightskyblue1")
    ) +
  
    coord_quickmap(xlim=xlim , ylim=ylim) +
  
    geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +
  
    geom_jitter(aes(x=lon, y=lat, size = gewicht, colour = as.character(trip)),
                width=0.1, height=0.05, alpha = 0.5) +
    scale_size(range = c(1,10)) +
  
    scale_x_continuous(breaks = seq(floor(xmin), ceiling(xmax), 1)) +
    scale_y_continuous(breaks = seq(floor(2*ymin)/2, ceiling(2*ymax)/2, 0.5)) +
    labs(x = NULL, y = NULL, size = "kg aanvoer/trek", colour="") +
    guides(size = guide_legend(nrow = 1)) + 
    facet_wrap(~species, ncol=3)
  
} else {

    cat("No lat - lon information available")
  
}



  

```


<!-- ########################################################################################## -->
<!-- Figure: vangst per soort --------------------------- -->
<!-- ########################################################################################## -->

```{r vangstpersoortenklasse, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "totalcatchperspecies", level   = 1, display = FALSE,
  caption = paste0("Totale vangst per trek en soort (",nspecies," soorten met hoogste vangst)"))

if (sum(!is.na(et$weight)) >0 ) {
  
  t <-
    et %>%
    filter(species %in% myspecies$species) %>% 
    group_by(vessel, trip, haul, species) %>% 
    summarise(gewicht=sum(weight))
  
  s <-
    t %>% 
    group_by(vessel, trip, species) %>% 
    summarise(totaal=sum(gewicht))  %>% 
    ungroup() %>% 
    arrange(desc(totaal)) %>% 
    dplyr::select(species) %>% 
    unlist()
  
  t %>% 
    ggplot(aes(x=haul, y=gewicht)) +
    theme_publication() +
    geom_bar(aes(fill=species), stat="identity") +
    coord_flip() +
    scale_x_reverse(breaks=seq(1, max(t$haul), 1)) +
    scale_y_continuous(breaks=scales::pretty_breaks(n=3)) +
    labs(y="aanvoer (kg)", title="aanvoer per soort en trek") +
    facet_wrap(.~trip, scales="free_x")
  

} else if (sum(!is.na(e$weight)) >0 ) {
  
  # function to reverse date axis: https://github.com/tidyverse/ggplot2/issues/4014 
  reverse2_trans <- function() {
    scales::trans_new(
      "reverse2",
      function(x) -1 * as.numeric(x), # Force values to be numeric for Date objects
      function(x) -1 * as.numeric(x)
    )
  }
  
  t <-
    e %>%
    filter(species %in% myspecies$species) %>% 
    group_by(vessel, date, species) %>% 
    summarise(gewicht=sum(weight))
  
  s <-
    t %>% 
    group_by(vessel, species) %>% 
    summarise(totaal=sum(gewicht))  %>% 
    ungroup() %>% 
    arrange(desc(totaal)) %>% 
    dplyr::select(species) %>% 
    unlist()
  
  t %>% 
    ggplot(aes(x=date, y=gewicht)) +
    theme_publication() +
    geom_bar(aes(fill=species), stat="identity") +
    coord_flip() +
    # scale_x_reverse() +
    scale_y_continuous(breaks=scales::pretty_breaks(n=3)) +
    scale_x_continuous(trans=c("date", "reverse2"), breaks=scales::pretty_breaks(n=length(unique(t$date)))) +
    # scale_x_date(date_breaks = "1 day", trans="reverse") +
    labs(y="aanvoer (kg)", title="aanvoer per soort en dag") 

} else {

    cat("No catch per species information available")
  
}



```

_`r fig_nums("totalcatchperspecies")`_

\newpage

<!-- ########################################################################################## -->
<!-- Figure: aanvoer per week                                                                   -->
<!-- ########################################################################################## -->

```{r aanvoerperweek, echo=FALSE, fig.align="center", fig.asp=1.2, message=FALSE, warning=FALSE}


# create list of catch by species by week
if (nrow(elog) > 0) {
  
  fig_nums(
    name    = "aanvoerperweek", level   = 1, display = FALSE,
    caption = "Vergelijking van aanvoer per soort en week")
  
  # sequence
  s <-
    elog %>% 
    filter(vessel %in% setvessel, year >= 2023, date <= max(e$date)) %>% 
    
    group_by(species) %>% 
    summarise(weight = sum(weight, na.rm=TRUE)) %>% 
    
    arrange (desc(weight)) %>% 
    group_by() %>% 
    slice_head(n=10) %>% 
    left_join(dplyr::select(asfis,
                            species, englishname), 
              by=c("species")) %>% 
    mutate(english_species = paste(englishname, species))
  
  # date
  t <-
    elog %>% 
    filter(vessel %in% setvessel, year >= 2023, date <= max(e$date)) %>% 
    
    mutate(species = ifelse(species == "SQC", "SQR", species)) %>%  
    mutate(species = ifelse(species %in% s$species, species, "ZZZ")) %>%

    group_by(year, week, species) %>% 
    summarise(weight = sum(weight, na.rm=TRUE)) %>% 
    
    left_join(dplyr::select(asfis,
                            species, dutchname, englishname, scientificname), 
              by=c("species")) %>% 
    mutate(
      dutchname   = ifelse(species == "ZZZ", "Overig", dutchname),
      englishname = ifelse(species == "ZZZ", "Other", englishname),
      scientificname = ifelse(species == "ZZZ", "", dutchname),
      english_species = paste(englishname, species)
    ) %>% 
    mutate(english_species = factor(english_species, levels=c(s$english_species, "Other ZZZ")))
  
  t %>% 
    ggplot(aes(x=week, y=weight)) +
    theme_publication() +
    theme(legend.position = "right", legend.direction = "vertical") +
    geom_bar(aes(fill=english_species), 
             stat="identity", position = position_stack(reverse = TRUE), colour="lightgray", linewidth=0.3) +
    labs(y="aanvoer (kg)", title=paste(paste(setvessel,collapse="/"), "aanvoer per soort en week"), fill="") +
    scale_fill_brewer(palette="Paired") +
    facet_wrap(~year, ncol=1)
  
}


```

_`r fig_nums("aanvoerperweek")`_

\newpage

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.asp=0.4}

library(patchwork)

fig_nums(
  name    = "aanvoerperweekenjaartotaal", level   = 1, display = FALSE,
  caption = "Cumulatieve aanvoer per week en jaar")

ee <- 
  elog %>%
  filter(vessel %in% setvessel, year >= year(max(e$date))-3, date <= max(e$date)) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  group_by(vessel, year, week) %>% 
  summarise(landingweight = sum(weight, na.rm=TRUE)) %>% 
  group_by(vessel, year) %>% 
  mutate(cumweight = cumsum(landingweight)) %>% 
  mutate(year = as.character(year)) 

# plot catch and landings by week
p1 <-
  ee %>% 
  ggplot() +
  theme_publication() +
  # theme(legend.position = "none") +
  geom_line(aes(x=week, y=landingweight, colour=year, linewidth=year, alpha=year)) +
  scale_linewidth_manual("year", values = c("2020"=.5,       "2021"=.5,       "2022"=.5,       "2023"=0.5,      "2024"=0.5,    "2025"=1.5), guide = "none") +
  scale_colour_manual   ("year", values = c("2020"="gray80", "2021"="gray60", "2022"="gray40", "2023"="gray40", "2024"="blue", "2025"="red")) +
  scale_alpha_manual("year", values = c(1, 1, 1, 1, 1, 0.5), guide = "none") +
  # geom_hline(aes(yintercept=meanweight, colour=species)) +
  # geom_text(aes(x=1, y=meanweight, label=as.integer(meanweight)), hjust=0, vjust=0, nudge_y = 200) +
  labs(title=paste(unique(ee$vessel), "Aanvoer week en jaar"), y="aanvoer (kg)")  
  # facet_grid(rows=vars(year), cols=vars(species), scales="free_y")

# plot cumulative catch and landings
p2 <-
  ee %>% 
  ggplot() +
  theme_publication() +
  # theme(legend.position = "none") +
  geom_line(aes(x=week, y=cumweight, colour=year, linewidth=year, alpha=year)) +
  scale_linewidth_manual("year", values = c("2020"=.5,       "2021"=.5,       "2022"=.5,       "2023"=0.5,      "2024"=0.5,    "2025"=1.5), guide = "none") +
  scale_colour_manual   ("year", values = c("2020"="gray80", "2021"="gray60", "2022"="gray40", "2023"="gray40", "2024"="blue", "2025"="red")) +
  scale_alpha_manual("year", values = c(1, 1, 1, 1, 1, 0.5), guide = "none") +
  # geom_hline(aes(yintercept=meanweight, colour=species)) +
  # geom_text(aes(x=1, y=meanweight, label=as.integer(meanweight)), hjust=0, vjust=0, nudge_y = 200) +
  labs(title=paste(unique(ee$vessel), "cumulatieve aanvoer week en jaar"), y="aanvoer (kg, cumulatief)")  
  # facet_grid(rows=vars(year), cols=vars(species), scales="free_y")

print(p1 + p2   + plot_layout(guides="collect", nrow = 1, heights = unit(c(14, 8, 4), c('null'))) & theme(legend.position = 'bottom',
        legend.direction = 'horizontal'))  
```

_`r fig_nums("aanvoerperweekenjaartotaal")`_

\newpage

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.asp=1.0}

fig_nums(
  name    = "aanvoerperweekenjaar", level   = 1, display = FALSE,
  caption = "Cumulatieve aanvoer per soort, week en jaar")

s <-
  elog %>% 
  filter(vessel %in% setvessel, year >= 2023, date <= max(e$date)) %>% 
  mutate(species = ifelse(species == "SQC", "SQR", species)) %>%  
  
  group_by(species) %>% 
  summarise(weight = sum(weight, na.rm=TRUE)) %>% 
  
  arrange (desc(weight)) %>% 
  group_by() %>% 
  slice_head(n=10) %>% 
  left_join(dplyr::select(asfis,
                          species, englishname), 
            by=c("species")) %>% 
  mutate(english_species = paste(englishname, species))

ee <- 
  elog %>%
  mutate(species = ifelse(species == "SQC", "SQR", species)) %>%  
  filter(vessel %in% setvessel, year >= year(max(e$date))-3, date <= max(e$date), species %in% s$species) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  group_by(vessel, species, year, week) %>% 
  summarise(landingweight = sum(weight, na.rm=TRUE)) %>% 
  group_by(vessel, species, year) %>% 
  mutate(cumweight = cumsum(landingweight)) %>% 
  mutate(year = as.character(year)) %>% 
  mutate(species = factor(species, levels=s$species))

# plot total catch and landings
ee %>% 
  ggplot() +
  theme_publication() +
  # theme(legend.position = "none") +
  geom_line(aes(x=week, y=cumweight, colour=year, linewidth=year, alpha=year)) +
  scale_linewidth_manual("year", values = c("2020"=.5,       "2021"=.5,       "2022"=.5,       "2023"=0.5,      "2024"=0.5,    "2025"=1.5), guide = "none") +
  scale_colour_manual   ("year", values = c("2020"="gray80", "2021"="gray60", "2022"="gray40", "2023"="gray40", "2024"="blue", "2025"="red")) +
  scale_alpha_manual("year", values = c(1, 1, 1, 1, 1, 0.5), guide = "none") +
  # geom_hline(aes(yintercept=meanweight, colour=species)) +
  # geom_text(aes(x=1, y=meanweight, label=as.integer(meanweight)), hjust=0, vjust=0, nudge_y = 200) +
  labs(title=paste(unique(ee$vessel), "cumulatieve aanvoer per soort, week en jaar"), y="aanvoer (kg, cumulatief)") +
  facet_wrap(~species, scales="free_y") 
  # facet_grid(rows=vars(year), cols=vars(species), scales="free_y")



```


_`r fig_nums("aanvoerperweekenjaar")`_

\newpage


```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.asp=1.0}

fig_nums(
  name    = "vangstsuccespersoort", level   = 1, display = FALSE,
  caption = "gemiddelde vangst (aanvoer, kg) per dag en per soort ")

ee <- 
  elog %>%
  filter(vessel %in% setvessel) %>% 
  filter(species %in% myspecies$species) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  mutate(species = factor(species, levels=c(myspecies$species))) %>% 
  group_by(vessel, year, date, species) %>% 
  summarise(catchperday = sum(weight, na.rm=TRUE)) %>% 
  group_by(vessel, year, species) %>% 
  summarise(catchperday = mean(catchperday, na.rm=TRUE)) 

# plot catchrate by species
ee %>% 
  ggplot(aes(x=year, y=catchperday)) +
  theme_publication() +
  # theme(legend.position = "none") +
  geom_point(data=ee %>% filter(year < max(ee$year, na.rm=TRUE))) +
  geom_point(data=ee %>% filter(year == max(ee$year, na.rm=TRUE)), colour="red", size=2) +
  geom_line() +
  labs(title=paste(unique(ee$vessel), "Gemiddelde aanvoer per dag"), y="aanvoer per dag (kg)")  +
  expand_limits(y=0) +
  facet_wrap(~species, scales="free_y")

```

_`r fig_nums("vangstsuccespersoort")`_
