---
title: "JACZON FLYSHOOT v2"
output:
  flexdashboard::flex_dashboard:
runtime: shiny
---



```{r global}

rm(list=ls())
options(dplyr.summarise.inform = FALSE)

library(rmarkdown)                   # note: requires knitr 1.21
require(tidyverse, quietly=TRUE)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
require(lubridate, quietly=TRUE)     # data handling
require(reshape2, quietly=TRUE)      # reshaping data; e.g. cast
require(RColorBrewer)                # colour schemes
library(viridis)
library(pander)

# Source all the utils
source("../R/FLYSHOOT utils.r")

spatialdir <- "C:/DATA/RDATA"
load(file.path(spatialdir, "world_lr_sf.RData"))
load(file.path(spatialdir, "world_mr_sf.RData"))
load(file.path(spatialdir, "world_lr_df.RData"))
load(file.path(spatialdir, "world_mr_df.RData"))

rect_df <-
  loadRData(file.path(spatialdir, "rect_df.RData")) %>% 
  rename(rect=ICESNAME) %>% 
  group_by(rect) %>% 
  filter(row_number() ==1) %>% 
  dplyr::select(rect, lon=long, lat)

asfis <-
  loadRData(file.path(spatialdir, "asfis.RData")) %>% 
  lowcase()


# set onedrive directory
onedrive <- get_onedrive(team="Martin Pastoors", site="FLYSHOOT - General/rdata")

load(file.path(onedrive, "haul.RData"))
trips <-
  haul %>% 
  group_by(vessel) %>% 
  distinct(vessel, date) %>% 
  mutate(diff = as.numeric(date-lag(date))) %>% 
  mutate(id   = ifelse(is.na(diff) | diff>1, 1, 0)) %>% 
  mutate(trip2 = cumsum(id)) %>% 
  group_by(vessel, trip2) %>% 
  mutate(trip2 = paste0(trip2," ",format(min(date), "%d/%m"), "-", format(max(date), "%d/%m"))) %>% 
  dplyr::select(-id, -diff)

haul <- haul %>% left_join(trips, by=c("vessel","date")) 

kisten <- loadRData(file.path(onedrive, "kisten.RData")) %>% 
  mutate(date=as.Date(datetime)) %>% 
  left_join(trips, by=c("vessel","date")) 
elog <- loadRData(file.path(onedrive, "elog.RData")) %>% 
  left_join(trips, by=c("vessel","date"))
elog_trek <- loadRData(file.path(onedrive, "elog_trek.RData")) %>% 
  left_join(trips, by=c("vessel","date"))
trip <- loadRData(file.path(onedrive, "trip.RData")) %>% 
  left_join(trips, by=c("vessel","date"))


maxdate <- max(haul$date, na.rm=TRUE)




```


Sidebar {.sidebar data-width=225}
=====================================

```{r}
# setvessel<-"SCH135"   ;settrip<-c("2023360","2023361") ; comparisons <- FALSE
# setvessel<-"SCH65"   ;settrip<-c("2023275") ; comparisons <- FALSE
# setvessel<-"SL9"   ;settrip<-c("2023081600041", "2023082100042") ; comparisons <- FALSE
# setvessel<-c("Z99", "SCH99")   ;settrip<- c("2023091") ; comparisons <- FALSE

vessels   <- c("SCH65", "SCH99", "SCH135", "SL9", "all")
years     <- c("2013":"2023","all")
months    <- c("1":"12","all")
divisions <- c("27.7.e","27.7.d","27.4.c","27.4.b","27.4.a","all")
species   <- c("CTC","SQR","MAC","MUR","HOM","WHG", "all")

shiny::selectInput(inputId  = "vessels", 
                   label    = "Vessels:",
                   choices  = vessels, 
                   selected = c("SL9","SCH135"),
                   multiple = TRUE)

shiny::selectInput(inputId = "years", 
                   label = "years:",
                   choices = years, 
                   selected = c("2021":"2023"),
                   multiple=TRUE)

shiny::selectInput(inputId = "months", 
                   label   = "months:",
                   choices = months, 
                   selected = "all",
                   multiple=TRUE)

shiny::selectInput(inputId = "divisions", 
                   label   = "divisions:",
                   choices = divisions, 
                   selected = "27.7.d",
                   multiple=TRUE)

shiny::selectInput(inputId = "species", 
                   label   = "species:",
                   choices = species, 
                   selected = c("MAC","CTC"),
                   multiple=TRUE)


shiny::selectInput(inputId = "showby", 
                   label   = "show by:",
                   choices = c("year","month","week"), 
                   selected = c("month"),
                   multiple=FALSE)

# variables to group by
my_vars <- 
  shiny::reactive({ x =  c("vessel", "year", "month") })

# species
s <-
  shiny::reactive({
    x =  
      elog %>% 
      mutate(division = tolower(faozone)) %>% 
      
      {if(all(!grep("all", input$vessels)))    filter(., vessel %in% input$vessels) else {.}} %>%  
      {if(all(!grep("all", input$years)))      filter(., year %in% input$years) else {.}} %>%  
      {if(all(!grep("all", input$months)))     filter(., month  %in% input$months) else {.}} %>%  
      {if(all(!grep("all", input$divisions))) filter(., division %in% input$divisions) else {.}} %>% 
      {if(all(!grep("all", input$species)))   filter(., species %in% input$species) else {.}} %>%  
      
      group_by(species) %>%
      summarise(weight = sum(weight, na.rm=TRUE)) %>%
      arrange(desc(weight)) %>%
      filter(row_number()<=10) %>%
      mutate(species = forcats::fct_reorder(species, desc(weight)))
    })

# elog
e = shiny::reactive({
    x =  elog %>%

      mutate(division = tolower(faozone)) %>%

      {if(all(!grep("all", input$vessels)))    filter(., vessel %in% input$vessels) else {.}} %>%  
      {if(all(!grep("all", input$years)))      filter(., year %in% input$years) else {.}} %>%  
      {if(all(!grep("all", input$months)))     filter(., month  %in% input$months) else {.}} %>%  
      {if(all(!grep("all", input$divisions))) filter(., division %in% input$divisions) else {.}} %>% 
      {if(all(!grep("all", input$species)))   filter(., species %in% input$species) else {.}} %>%  

      mutate(species = ifelse(species %in% s()$species, species, "ZZZ")) %>%
      mutate(species = factor(species, levels=c(as.character(s()$species),"ZZZ"))) %>%
      group_by_at(vars(one_of(c(my_vars(), "species")))) %>%
      summarise(weight = sum(weight, na.rm=TRUE)) %>%

      group_by_at(vars(one_of(c(my_vars())))) %>%
      mutate(perc = weight / sum(weight)) %>%

      mutate(label_pos = cumsum(weight) - weight / 2,
             perc_text = paste0(round(perc * 100), "%"))

    })



```

Trial stuff - draft explanation will follow. 

Page 1
=====================================  

Column {.tabset}
-------------------------------------

### Catch by vessel, year, month and species

```{r echo=FALSE}


# renderTable(e())

renderPlot(
  ggplot(
    data=e(), aes(x=month, y=weight)) +
    theme_publication() +
    geom_bar(aes(fill=species), stat="identity") +
    labs(x="month") +
    scale_x_continuous(breaks=as.integer(unique(e()$month))) +
    guides(fill = guide_legend(nrow = 1)) +
    facet_grid(year~vessel)
  )


```






Page 2
=====================================

Column {.tabset}
-------------------------------------

### Catch per tow

```{r eval=FALSE, include=FALSE}

t = shiny::reactive({
    x =  filter(haul, vessel == input$setvessel,
                 date >= input$dateRange[1],
                 date <= input$dateRange[2])} %>% 
           mutate(week = as.character(week), trip2 = as.character(trip2)) 
    ) 

xlim <- shiny::reactive({x=c(floor(min(t()$lon, na.rm=TRUE)), 
                             ceiling(max(t()$lon, na.rm=TRUE)))})
ylim <- shiny::reactive({x=c(floor(min(t()$lat, na.rm=TRUE)), 
                             ceiling(max(t()$lat, na.rm=TRUE)))})

renderPlot(
  ggplot(data=t(), 
         aes(x=lon, y=lat)) + 
    theme_publication() +
    theme(legend.position="none") +
    coord_quickmap(xlim=xlim() , ylim=ylim()) +
    geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +

    {if(input$showby == "by week") {
      geom_point(aes(colour=week), size=1)       
    } else {
      geom_point(aes(colour=trip2), size=1) 
    }}+
    
    guides(colour = guide_legend(nrow = 1)) +
    
    labs(title=input$showby) +
    
    {if(input$showby == "by week") {
      facet_wrap(~week)       
    } else {
      facet_wrap(~trip2)       
    }}
    
  )



```

### Catch compositions



### Catch proportions



Page 3
===================================== 

Column {.tabset}
-------------------------------------

### The data source

***

Something nice about the program

- ...

- ...

- ...

- ...

### The time panel

***

Here possibly a snapshot of an illustrative species, explaining how to read the graph

### The space panel


***

Here possibly a snapshot of an illustrative species, explaining how to read the graph

### Method


***

Very brief points on methods with then link to more detailed documentation