---
output:
  word_document:
    reference_docx: ../report_template.dotx
---


```{r setup, include=FALSE}

# ==========================================================================================================
# FLYSHOOT observer trips
# 
# 09/12/2024 first coding
# ==========================================================================================================

# Reset lists
rm(list=ls())

# ------------------------------------------------------------
# Select vessel and trip
# ------------------------------------------------------------

setvessel<-c("Z99", "SCH99")   ;settrip<- c( "2024190","2024191","2024192") ; comparisons <- FALSE

knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

options(dplyr.summarise.inform = FALSE)

# Libraries
library(rmarkdown)                   # note: requires knitr 1.21

require(tidyverse, quietly=TRUE)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
require(lubridate, quietly=TRUE)     # data handling
require(reshape2, quietly=TRUE)      # reshaping data; e.g. cast
require(RColorBrewer)                # colour schemes
library(viridis)
library(pander)

library(captioner)                   # for captions
tab_nums <- captioner::captioner(prefix = "Table " , levels=1, type=c("n"), infix=".")
fig_nums <- captioner::captioner(prefix = "Figure ", levels=1, type=c("n"), infix=".")

# Source all the utils
source("../../prf/R/my utils.r")

# source("../R/gfw.r")

get_onedrive <- function (team="Martin Pastoors", site="FLYSHOOT - General") {
  
  if (Sys.info()['sysname'] == 'Windows') {
    
    # set onedrive directory
    if(dir.exists(file.path(Sys.getenv('USERPROFILE'), team, site))) {
      onedrive <- file.path(Sys.getenv('USERPROFILE'), team, site)   
    } else if(dir.exists(file.path('C:/DATA/PFA', team, site))) {
      onedrive <- file.path('C:/DATA/PFA', team, site)
    } else if(dir.exists(file.path('D:/DATA/PFA', team, site))) {
      onedrive <- file.path('D:/DATA/PFA', team, site)
    } else {
      stop("Onedrive directory not found")
    }
  }
  
  return(onedrive)
}

spatialdir <- "C:/DATA/RDATA"

# spatial data files

load(file.path(spatialdir, "world_lr_sf.RData"))
load(file.path(spatialdir, "world_mr_sf.RData"))
# load(file.path(onedrive, "eez.df.RData"))
# load(file.path(onedrive, "fao.df.RData"))
# load(file.path(onedrive, "depth200.df.RData"))
# load(file.path(onedrive, "icesrectangles.df.RData"))

load(file.path(spatialdir, "world_lr_df.RData"))
load(file.path(spatialdir, "world_mr_df.RData"))

rect_df <-
  loadRData(file.path(spatialdir, "rect_df.RData")) %>% 
  rename(rect=ICESNAME) %>% 
  group_by(rect) %>% 
  filter(row_number() ==1) %>% 
  dplyr::select(rect, lon=long, lat)

asfis <-
  loadRData(file.path(spatialdir, "asfis.RData")) %>% 
  lowcase()


# ===============================================================================
#  Load and filter the trip data
# ===============================================================================

# set onedrive directory
onedrive <- get_onedrive(team="Martin Pastoors", site="FLYSHOOT - General/rdata")
datadir  <- get_onedrive(team="Martin Pastoors", site="FLYSHOOT - General/tripdata/SCH99")

# load datasets
load(file.path(onedrive, "haul.RData"))

load(file.path(onedrive, "kisten.RData"))
load(file.path(onedrive, "elog.RData"))
load(file.path(onedrive, "elog_trek.RData"))
load(file.path(onedrive, "trip.RData"))

load(file.path(onedrive, "gfw.RData"))


# making the data selections
h <- 
  haul %>%
  ungroup() %>% 
  filter(vessel  %in% setvessel,
         trip    %in% settrip) %>%  
  
  # TEMP FIX
  # mutate(lon = ifelse(lon>2 & lat < 51, -1*lon, lon)) %>% 

  ungroup()

 
m <-
  kisten %>% 
  filter(vessel  %in% setvessel,
         trip    %in% settrip) %>% 
  mutate(soorten = tolower(soorten)) %>% 
  distinct()

if (nrow(h) == 0 & nrow(m) == 0) {stop(paste("No hauls in the selected trip(s):", setvessel, settrip))}

e <-
  elog %>% 
  filter(vessel  %in% setvessel,
         trip    %in% settrip) 

et <-
  elog_trek %>% 
  filter(vessel  %in% setvessel,
         trip    %in% settrip) 

tt <-
  trip %>% 
  filter(vessel  %in% setvessel,
         trip    %in% settrip) %>% 
  
  # TEMP FIX
  # mutate(lon = ifelse(lon>2 & lat <51, -1*lon, lon)) %>% 

  left_join(dplyr::select(h,
                          vessel, trip, haul, catchweight, landingweight),
            by=c("vessel","trip","haul"))

if(nrow(tt) == 0) stop("error in trip data")

trip2 <-
  bind_rows(
    distinct(h, 
             vessel, trip, haul, haultime) %>% mutate(source="trek"),
    distinct(m,
             vessel, trip, haul, haultime=datetime) %>% mutate(source="marelec"),
    distinct(e, 
             vessel, trip, haul, haultime=catchdate) %>% mutate(source="elog"),
  ) %>% 
  group_by(vessel, trip, source) %>% 
  summarise(
    nhaul = n_distinct(haul),
    startdate = min(haultime, na.rm=TRUE),
    enddate   = max(haultime, na.rm=TRUE)
  )

g <-
  gfw %>% 
  filter(vessel  %in% setvessel) %>% 
  filter(date >= min(h$date), date <= max(h$date))

# calculate the scaling of plots
# h %>% filter(lat==max(lat, na.rm=TRUE)) %>% View()

xmin <- floor(min(h$lon, tt$lon, na.rm=TRUE))
xmax <- ceiling(max(h$lon, tt$lon, na.rm=TRUE))
ymin <- floor(2*min(h$lat, tt$lat, na.rm=TRUE))/2
ymax <- ceiling(2*max(h$lat, tt$lat, na.rm=TRUE))/2

# xmin <- floor(min(tt$lon, e$lon, na.rm=TRUE))
# xmax <- ceiling(max(tt$lon, e$lon, na.rm=TRUE))
# ymin <- floor(2*min(tt$lat, e$lat, na.rm=TRUE))/2
# ymax <- ceiling(2*max(tt$lat, e$lat, na.rm=TRUE))/2

# xmin <- floor(2 * (min(h$lon, na.rm=TRUE)-0.5))/2;
# xmax <- ceiling(2 * (max(h$lon, na.rm=TRUE)+0.5))/2;
# ymin <- floor(2 * (min(h$lat, na.rm=TRUE)-0.5))/2;
# ymax <- ceiling(2* (max(h$lat, na.rm=TRUE)+0.5))/2

# xdistance <- geosphere::distGeo(matrix(c(xmin, ymin), ncol=2, byrow=FALSE),
#                                 matrix(c(xmax, ymin) , ncol=2, byrow=FALSE) )/1000
# ydistance <- geosphere::distGeo(matrix(c(xmin, ymin), ncol=2, byrow=FALSE),
#                                 matrix(c(xmin, ymax) , ncol=2, byrow=FALSE) )/1000
 
xdistance <- geo::arcdist(lat=ymin, lon=xmin, lat1 = ymin, lon1 = xmax, scale = "km")
ydistance <- geo::arcdist(lat=ymin, lon=xmin, lat1 = ymax, lon1 = xmin, scale = "km")
aspect    <- ydistance/xdistance


# Read observer data
odfiles  <-
  list.files(
    path=file.path(datadir),
    pattern="Dutch_Flyshoot",
    recursive=TRUE,
    full.names = TRUE)

od <- disc <- lf <- data.frame(stringsAsFactors = FALSE)

for (fn in odfiles) {
  od <-
    bind_rows(
      od,
      readxl::read_excel(
        fn,
        sheet = "1_Operation_details",
        col_names = TRUE,
        col_types = "text"
      ) %>% 
        lowcase() %>% 
        mutate(across (c("vesseltotalcatchkg", "observertotalcatchkg", 
                         "vesseltotalcatch-landedkg", "observertotaldiscardedkg",
                         "totallandedcatchkg"),
                       as.numeric)) %>% 
        rename(vesseldiscardedkg = "vesseltotalcatch-landedkg") %>% 
        mutate(
          triphaul = paste(trip, stringr::str_pad(haul, pad="0",width=2), sep="_"),
          observercalctotalcatchkg = ifelse(is.na(observertotaldiscardedkg), NA, totallandedcatchkg + observertotaldiscardedkg)
        )
    )
} 

for (fn in odfiles) {
  disc <-
    bind_rows(
      disc,
      readxl::read_excel(
        fn,
        sheet = "2_Discard_Sample_Weights",
        col_names = TRUE,
        col_types = "text",
        range="A2:E200"
      ) %>% 
        lowcase() %>% 
        mutate(across (c("totalweightkg"), as.numeric)) %>% 
        mutate(trip = paste0("2024",trip))
    )
} 

disc2 <-
  disc %>% 
  drop_na(vessel) %>% 
  group_by(vessel, trip, haul) %>% 
  mutate(prop = totalweightkg/sum(totalweightkg)) %>% 
  # mutate(trip = paste0("2024",trip)) %>%
  mutate(haul = as.integer(haul)) %>% 
  left_join(dplyr::select(h,
                          vessel, trip, haul, catchweight), 
            by=c("vessel","trip","haul")) %>% 
  left_join(m %>% group_by(vessel, trip, haul) %>% summarise(landingweight=sum(gewicht,na.rm=TRUE)),
            by=c("vessel","trip","haul")) %>% 
  mutate(discardweight = catchweight - landingweight) %>% 
  mutate(discardweightperspecies = prop * discardweight) %>% 
  mutate(triphaul = paste(trip, stringr::str_pad(haul, pad="0",width=2), sep="_"))


fn <- odfiles[1]

lf <- data.frame(stringsAsFactors = FALSE)

for (fn in odfiles) {
  print(fn)
  mysheets   <- readxl::excel_sheets(fn)
  mylfsheets <- mysheets[grepl("LF_", mysheets)]
  mytrip     <- paste0("2024", 189 + as.integer(substr(basename(fn),25,26)))
  
  ms         <- mylfsheets[1]
  for(ms in mylfsheets) {
    print(ms)
    t1 <- readxl::read_excel(fn, sheet=ms, range="B2:M2", col_names = FALSE) %>% tidyr::pivot_longer(names_to = "test", values_to = "haul", c("...1":"...12")) %>% 
      drop_na(haul) %>% 
      dplyr::select(-test)
  
    myvar     <- as.character(readxl::read_excel(fn, sheet=ms, range="A5", col_names = FALSE))     
    myspecies <- substr(ms,4,6)
    mylengths <- readxl::read_excel(fn,sheet=ms,range="A6:A200", col_names = FALSE) %>% setNames("length") %>% filter(!is.na(length) & length != "Average") %>% 
      mutate(length = as.integer(length))
  
    for(counter in 1:nrow(t1)) {
      myh <- as.integer(t1[counter,1])
      print(myh) 
      t2 <- readxl::read_excel(fn,sheet=ms,range=cellranger::cell_limits(c(6, 2), c(6+nrow(mylengths)-1, 2)), col_names = FALSE) %>% 
        setNames("count") %>% 
        mutate(count = as.integer(count)) %>% 
        bind_cols(trip = mytrip) %>% 
        bind_cols(haul = myh) %>% 
        bind_cols(species = myspecies) %>% 
        bind_cols(variable = myvar) %>% 
        bind_cols(mylengths) %>% 
        drop_na(count) %>% 
        relocate(count, .after=length)
      
      lf <-
        bind_rows(lf, t2)
    }
  }  
} 
  

```


# Results of observer trips on `r unique(trip2$vessel)`, trips: `r paste(as.character(unique(trip2$trip)), collapse=", ")`

### Trip start: `r format(min(trip2$startdate), '%d/%m/%Y')`, end: `r format(max(trip2$enddate), '%d/%m/%Y')`

Report generated: `r format(Sys.time(), '%d/%m/%Y')`

&nbsp;

**Testing the total catch estimates**

Here comparing the total catch estimate from the skipper (based on Marelec) with the sum of the observer-weighed discards and the landed catch of that haul. Colours indicate the three different trips. Haul 7 of 3rd trip (i.e. 192) has the highest deviation in total catch. 

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

# od %>% filter(!is.na(observertotaldiscardedkg)) %>% View()

od %>% 
  ggplot(aes(x=vesseltotalcatchkg, y=observercalctotalcatchkg)) +
  theme_publication() +
  geom_point(aes(colour=trip, size=vesseltotalcatchkg), shape=21) +
  geom_text(aes(colour=trip, label=triphaul), hjust=0, nudge_x = 5) +
  geom_abline(slope=1) +
  expand_limits(x=0, y=0) +
  scale_size_continuous(range=c(0.5,5)) +
  scale_x_continuous(expand = c(0, 100))

  

```

Discard rates from observer data

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

# od %>% filter(!is.na(observertotaldiscardedkg)) %>% View()

od %>% 
  drop_na(observertotaldiscardedkg) %>% 
  ungroup() %>% 
  dplyr::select(vessel, trip, haul, observertotaldiscardedkg, totallandedcatchkg, observercalctotalcatchkg) %>% 
  # summarise(
  #   observertotaldiscardedkg = sum(observertotaldiscardedkg),
  #   totallandedcatchkg       = sum(totallandedcatchkg),
  #   observercalctotalcatchkg = sum(observercalctotalcatchkg)
  # ) %>% 
  mutate(
    discardrate             = observertotaldiscardedkg/observercalctotalcatchkg
  ) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", observertotaldiscardedkg:discardrate) %>% 
  reshape2::dcast(vessel+trip+haul~variable, value.var = "data", mean, margins="trip") %>% 
  mutate(discardrate = scales::percent(discardrate, accuracy=1)) %>% 
  relocate(observercalctotalcatchkg, .after=totallandedcatchkg) %>% 
  relocate(discardrate,              .after=observercalctotalcatchkg) %>% 
  pandoc.table(.,
           style = "simple",
           split.tables=400, 
           justify = c("right"),
           missing=".",
           round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))


```

Discard rates from vessel records

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

# od %>% filter(!is.na(observertotaldiscardedkg)) %>% View()


od %>% 
  drop_na(observertotaldiscardedkg) %>% 
  ungroup() %>% 
  dplyr::select(vessel, trip, haul, 
                totallandedcatchkg, 
                vesseltotalcatchkg) %>% 
  mutate(
    vesselcalculateddiscardskg = vesseltotalcatchkg - totallandedcatchkg,
    discardrate = vesselcalculateddiscardskg/vesseltotalcatchkg
  ) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", 
                      totallandedcatchkg:discardrate) %>% 
  reshape2::dcast(vessel+trip+haul~variable, value.var = "data", mean, margins="trip") %>% 
  mutate(discardrate = scales::percent(discardrate, accuracy=1)) %>% 
  relocate(vesseltotalcatchkg, .after=totallandedcatchkg) %>% 
  relocate(discardrate,        .after=vesseltotalcatchkg) %>% 
  pandoc.table(.,
           style = "simple",
           split.tables=400, 
           justify = c("right"),
           missing=".",
           round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))

```


Here comparing the discards as calculated from the total catch estimate of the skipper minus the landed catch of that haul ("vesseldiscardedkg") with observer-weighed discards. Colours indicate the three different trips. Again, haul 7 of 3rd trip (i.e. 192) has the highest deviation in total discards (overestimated based on the Marelec readings)

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

# od %>% filter(!is.na(observertotaldiscardedkg)) %>% View()

od %>% 
  ggplot(aes(x=vesseldiscardedkg, y=observertotaldiscardedkg)) +
  theme_publication() +
  geom_point(aes(colour=trip, size=vesseltotalcatchkg), shape=21) +
  geom_text(aes(colour=trip, label=triphaul), hjust=0, nudge_x = 5) +
  geom_abline(slope=1) +
  expand_limits(x=0, y=0) +
  scale_size_continuous(range=c(0.5,5)) +
  scale_x_continuous(expand = c(0, 100))


```

\newpage

**Species composition of discards**

```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

tmp <-
  disc2 %>% 
  group_by(trip, species) %>% 
  summarise(discardweightperspecies = sum(discardweightperspecies, na.rm=TRUE)) %>% 
  group_by(species) %>% 
  mutate(totaldiscardweightperspecies = sum(discardweightperspecies, na.rm=TRUE)) %>% 
  mutate(species  = ifelse(totaldiscardweightperspecies<20, "ZZZ",species)) %>% 
  group_by(trip, species) %>% 
  summarise(discardweightperspecies = sum(discardweightperspecies, na.rm=TRUE)) %>% 
  left_join(dplyr::select(asfis, species, englishname), 
            by=c("species")) %>% 
  mutate(
    englishname = ifelse(species=="ZZZ", "Other species",englishname),
    longname    = paste(species, englishname))

tmpher <-
  tmp %>% 
  mutate(species = ifelse(species=="HER",species, "ZZZ")) %>% 
  group_by(species) %>% 
  summarise(discardweightperspecies = sum(discardweightperspecies, na.rm=TRUE)) %>% 
  mutate(prop = discardweightperspecies/sum(discardweightperspecies))

tt <-
  tmp %>% 
  group_by(species, englishname, longname) %>% 
  summarise(discardweightperspecies = sum(discardweightperspecies, na.rm=TRUE)) %>% 
  arrange((discardweightperspecies)) 

tmp %>%   
  ggplot(aes(x=factor(longname, levels=tt$longname), y=discardweightperspecies)) +
  theme_publication() +
  theme(legend.position="none") +
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  # theme(axis.text.x=element_blank()) +
  geom_bar(aes(fill=trip), stat="identity") +
  coord_flip() +
  labs(x="") +
  facet_wrap(~trip, nrow=1)

  

```


An overview of the discards weight per species and per haul. Raising based on discard proportions in the observer samples applied to the estimated discards from the vessel records (total catch weight - landings weight).   

```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

disc2 %>% 
  ggplot(aes(x=triphaul, y=discardweightperspecies)) +
  theme_publication() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(axis.text.x=element_blank()) +
  geom_bar(aes(fill=trip), stat="identity") +
  facet_wrap(~species)

  

```

\newpage

**Length compositions**

An overview of the average length compositions by species sample

```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

lf %>% 
  mutate(triphaul = paste(trip, haul)) %>% 
  group_by(triphaul, species, length) %>% 
  summarise(count = sum(count, na.rm=TRUE)) %>% 
  group_by(species, length) %>% 
  summarise(
    count = sum(count, na.rm=TRUE),
    nobs = n_distinct(triphaul)) %>% 
  group_by(species) %>% 
  mutate(prop = count/sum(count)) %>% 
  # View()

  ggplot(aes(x=length, y=prop)) +
  theme_publication() +
  geom_bar(stat="identity") +
  expand_limits(x=0) +
  facet_wrap(~species)

  

```


```{r catchperhaul, echo=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE}

fig_nums(
  name    = "totalcatchperhaul", level   = 1, display = FALSE,
  caption = "Totale aanvoer per trek (berekend vanuit Marelec).")

xlim <- c(-1,2); ylim <- c(49,51)

t1 <- 
  disc2 %>% 
  group_by(vessel, trip, haul) %>% 
  summarise(discardsweight = sum( discardweightperspecies, na.rm=TRUE)) %>% 
  left_join(dplyr::select(h,
                          vessel, trip, haul, lat, lon), 
            by=c("vessel","trip","haul")) %>% 
  left_join(m %>% group_by(vessel, trip, haul) %>% summarise(landingweight=sum(gewicht,na.rm=TRUE)),
            by=c("vessel","trip","haul")) 

t2 <- 
  disc2 %>% 
  filter(species == "HER") %>% 
  group_by(vessel, trip, haul, species) %>% 
  summarise(discardsweight = sum( discardweightperspecies, na.rm=TRUE)) %>% 
  left_join(dplyr::select(h,
                          vessel, trip, haul, lat, lon), 
            by=c("vessel","trip","haul")) 

t1 %>%
  ggplot(aes(x=lon, y=lat)) +

  theme_publication() +

  coord_quickmap(xlim=xlim , ylim=ylim) +

  geom_text(data=rect_df, aes(x=lon, y=lat, label=rect),
            hjust=0, vjust=0, nudge_x=0.05, nudge_y=0.025, colour="white") +

  geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +

  # geom_point(aes(size=landingweight), colour="black", alpha=0.4) +
  geom_point(aes(size=discardsweight), colour="red", alpha=0.4) +
  geom_point(data=t2,
             aes(size=discardsweight), colour="green", alpha=0.8) +

  scale_size(range = c(1,10)) +

  scale_x_continuous(breaks = seq(floor(xmin), ceiling(xmax), 1)) +
  scale_y_continuous(breaks = seq(floor(2*ymin)/2, ceiling(2*ymax)/2, 0.5)) +
  guides(size = guide_legend(nrow = 1))
  # + facet_wrap(~source)



```


```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

disc %>% 
  filter(paste(trip, haul, species) %in% paste(lf$trip, lf$haul, lf$species)) %>% 
  pandoc.table(.,
           style = "simple",
           split.tables=400, 
           justify = c("right"),
           missing=".",
           round=c(0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))


```

```{r, echo=FALSE, fig.asp=0.8, fig.align="center", message=FALSE, warning=FALSE}

lf %>% 
  pandoc.table(.,
           style = "simple",
           split.tables=400, 
           justify = c("right"),
           missing=".",
           round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))


```

