---
output:
  word_document:
    reference_docx: ../report_template.dotx
---


```{r setup, include=FALSE}

# ==========================================================================================================
# Generate standard FLYSHOOT tripreport
# 
# 11/01/2023 first coding
# 20/01/2023 v1 with inclusion of elog data
# 23/02/2023 added multiple trips possibility
# ==========================================================================================================

# Reset lists
rm(list=ls())

# ------------------------------------------------------------
# Select vessel and trip
# ------------------------------------------------------------

setvessel<-"SCH135"   ;settrip<-"2023330" ;
# setvessel<-"SCH65"   ;settrip<-"2023247" ;
# setvessel<-"SL9"   ;settrip<-c("2023418") ;
# setvessel<-"Z99"   ;settrip<-"2023073" ;

# ------------------------------------------------------------

knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

options(dplyr.summarise.inform = FALSE)


# Libraries
library(rmarkdown)                   # note: requires knitr 1.21

require(tidyverse, quietly=TRUE)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
require(lubridate, quietly=TRUE)     # data handling
require(reshape2, quietly=TRUE)      # reshaping data; e.g. cast
require(RColorBrewer)                # colour schemes
library(viridis)
library(pander)

library(captioner)                   # for captions
tab_nums <- captioner::captioner(prefix = "Table " , levels=1, type=c("n"), infix=".")
fig_nums <- captioner::captioner(prefix = "Figure ", levels=1, type=c("n"), infix=".")

# Source all the utils
source("../../prf/R/my utils.r")

get_onedrive <- function (team="Martin Pastoors", site="FLYSHOOT - General") {
  
  if (Sys.info()['sysname'] == 'Windows') {
    
    # set onedrive directory
    if(dir.exists(file.path(Sys.getenv('USERPROFILE'), team, site))) {
      onedrive <- file.path(Sys.getenv('USERPROFILE'), team, site)   
    } else if(dir.exists(file.path('C:/DATA/PFA', team, site))) {
      onedrive <- file.path('C:/DATA/PFA', team, site)
    } else if(dir.exists(file.path('D:/DATA/PFA', team, site))) {
      onedrive <- file.path('D:/DATA/PFA', team, site)
    } else {
      stop("Onedrive directory not found")
    }
  }
  
  return(onedrive)
}

spatialdir <- "C:/DATA/RDATA"

# spatial data files

load(file.path(spatialdir, "world_lr_sf.RData"))
load(file.path(spatialdir, "world_mr_sf.RData"))
# load(file.path(onedrive, "eez.df.RData"))
# load(file.path(onedrive, "fao.df.RData"))
# load(file.path(onedrive, "depth200.df.RData"))
# load(file.path(onedrive, "icesrectangles.df.RData"))

load(file.path(spatialdir, "world_lr_df.RData"))
load(file.path(spatialdir, "world_mr_df.RData"))

rect_df <-
  loadRData(file.path(spatialdir, "rect_df.RData")) %>% 
  rename(rect=ICESNAME) %>% 
  group_by(rect) %>% 
  filter(row_number() ==1) %>% 
  dplyr::select(rect, lon=long, lat)

asfis <-
  loadRData(file.path(spatialdir, "asfis.RData")) %>% 
  lowcase()


# ===============================================================================
#  Load and filter the trip data
# ===============================================================================

# set onedrive directory
onedrive <- get_onedrive(team="Martin Pastoors", site="FLYSHOOT - General/rdata")

# load datasets
load(file.path(onedrive, "haul.RData"))

load(file.path(onedrive, "kisten.RData"))
load(file.path(onedrive, "elog.RData"))
load(file.path(onedrive, "trip.RData"))


# making the data selections
h <- 
  haul %>%
  ungroup() %>% 
  filter(vessel  %in% setvessel,
         trip    %in% settrip) %>%  
  ungroup()

 
m <-
  kisten %>% 
  filter(vessel  %in% setvessel,
         trip    %in% settrip) %>% 
  mutate(soorten = tolower(soorten)) %>% 
  distinct()

if (nrow(h) == 0 & nrow(m) == 0) {stop(paste("No hauls in the selected trip(s):", setvessel, settrip))}

e <-
  elog %>% 
  filter(vessel  %in% setvessel,
         trip    %in% settrip) 

tt <-
  trip %>% 
  filter(vessel  %in% setvessel,
         trip    %in% settrip) %>% 
  left_join(dplyr::select(h,
                          vessel, trip, haul, totalcatch),
            by=c("vessel","trip","haul"))
  
trip2 <-
  bind_rows(
    distinct(h, 
             vessel, trip, haul, haultime) %>% mutate(source="trek"),
    distinct(m,
             vessel, trip, haul, haultime=datetime) %>% mutate(source="marelec"),
    distinct(e, 
             vessel, trip, haul, haultime=catchdate) %>% mutate(source="elog"),
  ) %>% 
  group_by(vessel, trip, source) %>% 
  summarise(
    nhaul = n_distinct(haul),
    startdate = min(haultime, na.rm=TRUE),
    enddate   = max(haultime, na.rm=TRUE)
  )

# calculate the scaling of plots
# h %>% filter(lat==max(lat, na.rm=TRUE)) %>% View()

# xmin <- min(h$lon, na.rm=TRUE)
# xmax <- max(h$lon, na.rm=TRUE)
# ymin <- min(h$lat, na.rm=TRUE)
# ymax <- max(h$lat, na.rm=TRUE)

xmin <- floor(min(tt$lon, e$lon, na.rm=TRUE))
xmax <- ceiling(max(tt$lon, e$lon, na.rm=TRUE))
ymin <- floor(2*min(tt$lat, e$lat, na.rm=TRUE))/2
ymax <- ceiling(2*max(tt$lat, e$lat, na.rm=TRUE))/2

# xmin <- floor(2 * (min(h$lon, na.rm=TRUE)-0.5))/2;
# xmax <- ceiling(2 * (max(h$lon, na.rm=TRUE)+0.5))/2;
# ymin <- floor(2 * (min(h$lat, na.rm=TRUE)-0.5))/2;
# ymax <- ceiling(2* (max(h$lat, na.rm=TRUE)+0.5))/2

# xdistance <- geosphere::distGeo(matrix(c(xmin, ymin), ncol=2, byrow=FALSE),
#                                 matrix(c(xmax, ymin) , ncol=2, byrow=FALSE) )/1000
# ydistance <- geosphere::distGeo(matrix(c(xmin, ymin), ncol=2, byrow=FALSE),
#                                 matrix(c(xmin, ymax) , ncol=2, byrow=FALSE) )/1000
 
xdistance <- geo::arcdist(lat=ymin, lon=xmin, lat1 = ymin, lon1 = xmax, scale = "km")
ydistance <- geo::arcdist(lat=ymin, lon=xmin, lat1 = ymax, lon1 = xmin, scale = "km")
aspect    <- ydistance/xdistance

# if (aspect < 1.4) {
# 
#   projecty  <- ((1.4/aspect * ydistance) - ydistance)/2
#   ymax      <- as.numeric(geosphere::destPoint(p=c(xmin, ymax), b=0, d=projecty*1000)[,2])
#   ymin      <- as.numeric(geosphere::destPoint(p=c(xmin, ymin), b=180, d=projecty*1000)[,2])
#   # ymin      <- floor  (2 * (ymin-0.5))/2
#   # ymax      <- ceiling(2* (ymax+0.5))/2
# 
# } else {
# 
#   projectx  <- ((aspect/1.4 * xdistance) - xdistance)/2
#   xmin      <- as.numeric(geosphere::destPoint(p=c(xmin, ymin), b=270, d=projectx*1000)[,1])
#   xmax      <- as.numeric(geosphere::destPoint(p=c(xmax, ymin), b=90,  d=projectx*1000)[,1])
#   # xmin      <- floor(2 * (xmin-0.5))/2;
#   # xmax      <- ceiling(2 * (xmax+0.5))/2;
# }

# tmp <- data.frame(port="Boulogne sur Mer")
# tidygeocoder::geocode(tmp, city=port)

# tmp <-
#   h %>% 
#   mutate(nexthaultime = lead(haultime)) %>% 
#   mutate(nexthaultime = ifelse(is.na(nexthaultime), lubridate::dmy_hm("31/12/2023 23:59"), nexthaultime)) %>% 
#   mutate(nexthaultime = as_datetime(nexthaultime))

# r2 <-
#   sqldf::sqldf("select tmp.vessel, tmp.trip, tmp.haul, m.datetime, m.gewicht  from tmp
#                 join m on m.datetime >= tmp.haultime and 
#                           m.datetime <  tmp.nexthaultime") %>% 
#   as_tibble() %>% 
#   group_by(vessel, trip, haul) %>% 
#   summarise(aanvoer=sum(gewicht, na.rm=TRUE))


```


# Vessel tripreport `r unique(trip2$vessel)`, trip `r as.character(unique(trip2$trip))`

### Trip start: `r format(min(trip2$startdate), '%d/%m/%Y')`, end: `r format(max(trip2$enddate), '%d/%m/%Y')`

Report generated: `r format(Sys.time(), '%d/%m/%Y')`

&nbsp;

<!-- ########################################################################################## -->
<!-- Table: samenvatting reis                                       --------------------------- -->
<!-- ########################################################################################## -->

```{r samenvatting, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}

tab_nums(
  name    = "samenvatting", 
  level = 1, display = FALSE,
  caption = "Samenvatting van de reis")

# summarize the haul information
c1 <- 
  h %>%
  group_by(vessel,trip) %>%
  summarise(
    nhauls            = n(),
    startdate         = min(date, na.rm=TRUE),
    enddate           = max(date, na.rm=TRUE),
    weeks             = paste(unique(week), collapse=", "),
    months            = paste(unique(month), collapse=", "),
    divisions         = paste(unique(division), collapse="; "),
    rects             = paste(unique(rect), collapse="; "),
  ) %>% 
  mutate (
    ndays             = as.numeric(difftime(enddate, startdate, units="days")) + 1,
    haulsperday       = nhauls/ndays
  ) %>% 
  mutate(
    approxmileage = as.integer(sum(tt$distance, na.rm=TRUE))
  )

# summarize the catch data from elog and marelec data
c2 <- 
  bind_rows(
    m %>%
      group_by(vessel,trip) %>%
      summarise(aanvoer = sum(gewicht, na.rm=TRUE)) %>% 
      mutate(source="marelec"),  
    e %>%
      group_by(vessel,trip) %>%
      summarise(aanvoer = sum(weight, na.rm=TRUE)) %>% 
      mutate(source="elog")
  )
  
# belangrijkste soorten
c3 <-
  bind_rows(
    e %>% 
      group_by(trip, species) %>% 
      summarise(gewicht = sum(weight, na.rm=TRUE)) %>% 
      arrange (desc(gewicht)) %>% 
      slice_head(n=8),
    e %>% 
      group_by(trip, species) %>% 
      summarise(gewicht = sum(weight, na.rm=TRUE)) %>% 
      arrange (desc(gewicht)) %>% 
      filter(row_number() > 8) %>% 
      group_by(trip) %>% 
      summarise(gewicht = sum(gewicht)) %>% 
      mutate(species="OTHER"),
    e %>% 
      group_by(trip) %>% 
      summarise(gewicht = sum(weight, na.rm=TRUE)) %>% 
      mutate(species="TOTAL")
  )


c1 %>% 
  t() %>% 
  no.emphasis.table() %>% 
  pandoc.table(.,
             style = "simple",
             split.tables=400, 
             justify = "left",
             missing=".",
             round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))

c2 %>% 
  ungroup() %>% 
  dplyr::select(-vessel) %>% 
  # mutate(vangst = "aanvoer") %>% 
  pivot_wider(names_from = source, values_from = aanvoer) %>% 
  pandoc.table(.,
             style = "simple",
             split.tables=400, 
             justify = "left",
             missing=".",
             round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))

c3 %>% 
  ungroup() %>% 
  mutate(vangst = "aanvoer (elog)") %>% 
  pivot_wider(names_from = species, values_from = gewicht) %>% 
  pandoc.table(.,
             style = "simple",
             split.tables=400, 
             justify = "left",
             missing=".",
             round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
  

```

_`r tab_nums("samenvatting")`_

\newpage

<!-- ########################################################################################## -->
<!-- Table: vangst --------------------------- -->
<!-- ########################################################################################## -->

**Aanvoer per soort (Marelec)**

```{r vangstpersoort, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}

tab_nums(
  name    = "vangstpersoort", 
  level = 1, display = FALSE,
  caption = "Vangst per soort (kg)")

t1 <-
  m %>% 
  dplyr::group_by(trip, soorten) %>% 
  dplyr::summarize(gewicht     = sum(gewicht, na.rm=TRUE)) %>% 
  dplyr::group_by(trip) %>% 
  dplyr::mutate(perc = gewicht/sum(gewicht, na.rm=TRUE)) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", gewicht:perc) 

t1 %>% 
  reshape2::dcast(soorten ~ trip+variable, value.var="data", sum, margins = c("soorten")) %>% 
  mutate(across(grep("perc", names(.)),~scales::percent(., accuracy=1))) %>% 
  mutate(soorten = tolower(soorten)) %>% 

  pander::pandoc.table(.,
               style = "simple",
               split.tables=400, 
               justify = "right",
               missing=".",
               round=c(0))
  

```

_`r tab_nums("vangstpersoort")`_

**Aanvoer per soort en dag (Marelec)**

```{r vangstpersoortendag, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}

tab_nums(
  name    = "vangstpersoortendag", 
  level = 1, display = FALSE,
  caption = "Vangst per soort (kg) per dag")

t1 <-
  m %>% 
  mutate(date = as.Date(datetime)) %>% 
  mutate(soorten = tolower(soorten)) %>% 
  dplyr::group_by(date, soorten) %>% 
  dplyr::summarize(gewicht     = sum(gewicht, na.rm=TRUE)) %>% 
  dplyr::mutate(date = format(date, "%d/%m")) %>% 
  reshape2::dcast(soorten ~ date, value.var = "gewicht", sum, margins = c("date", "soorten")) %>% 
  pander::pandoc.table(.,
               style = "simple",
               split.tables=400, 
               justify = "right",
               missing=".",
               round=c(0))
  

```

_`r tab_nums("vangstpersoortendag")`_

\newpage

<!-- ########################################################################################## -->
<!-- Figure: Total Landings per Haul (map)                                                         -->
<!-- ########################################################################################## -->

```{r catchperhaul, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

fig_nums(
  name    = "catchperhaul", level   = 1, display = FALSE,
  caption = "Vangst (aanvoer) per trek (kg)")

xlim <- c(xmin,xmax); ylim <- c(ymin,ymax)

# if (sum(!is.na(h$totalcatch)) >0 ) {
# 
# 
#   tt %>%
#     arrange(haul) %>%
# 
#     # ggplot(aes(x=shootlon, y=shootlat)) +
#     ggplot() +
# 
#     theme_publication() +
#     theme(
#       # panel.grid.major = element_blank(),
#       panel.grid.minor = element_blank(),
#       text             = element_text(size=12),
#       plot.margin      = unit(c(0,0,0,0), "cm"),
#       panel.background = element_rect(fill = "lightskyblue1")
#     ) +
# 
# 
#     coord_quickmap(xlim=xlim , ylim=ylim) +
# 
#     geom_text(data=rect_df, aes(x=lon, y=lat, label=rect),
#               hjust=0, vjust=0, nudge_x=0.05, nudge_y=0.025, colour="white") +
# 
#     geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +
# 
#     geom_path (aes(x=lon, y=lat, colour=trip)) +
#     geom_point(aes(x=lon, y=lat, size = totalcatch, colour=trip), 
#                 alpha = 0.5) +
#     geom_point(data=filter(tt, 
#                            action=="embarked"), 
#                aes(x=lon, y=lat), colour="green", size=5, alpha=0.5) +
#     geom_point(data=filter(tt, 
#                            action=="disembarked"), 
#                aes(x=lon, y=lat), colour="red", size=5, alpha=0.5) +
#     
#     # geom_text(aes(x=lon, y=lat, label=haul), 
#     #             hjust=0, vjust=0.5) +
#   
#     scale_size(range = c(1,10)) +
# 
#     scale_x_continuous(breaks = seq(floor(xmin), ceiling(xmax), 1)) +
#     scale_y_continuous(breaks = seq(floor(2*ymin)/2, ceiling(2*ymax)/2, 0.5)) +
#     labs(x = NULL, y = NULL, size = "totale vangst/trek", colour="") +
#     guides(size = guide_legend(nrow = 1))
#     # + facet_wrap(~source)
# } 

if (sum(!is.na(m$gewicht)) >0 ) {

  fig_nums(
    name    = "totalcatchperhaul", level   = 1, display = FALSE,
    caption = "Totale aanvoer per trek (vanuit Marelec).")

  m %>%
    group_by(vessel, trip, haul=haul2) %>%
    summarise(gewicht = sum(gewicht, na.rm=TRUE)) %>%

    left_join(dplyr::select(h,
                            vessel, trip, haul, lat, lon, date),
              by = c("vessel","trip","haul")) %>%

    # ggplot(aes(x=shootlon, y=shootlat)) +
    ggplot() +

    theme_publication() +
    theme(
      # panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      text             = element_text(size=12),
      plot.margin      = unit(c(0,0,0,0), "cm"),
      panel.background = element_rect(fill = "lightskyblue1")
    ) +


    coord_quickmap(xlim=xlim , ylim=ylim) +

    geom_text(data=rect_df, aes(x=lon, y=lat, label=rect),
              hjust=0, vjust=0, nudge_x=0.05, nudge_y=0.025, colour="white") +

    geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +

    geom_path(data=tt, aes(x=lon, y=lat)) +
    geom_point(data=slice(tt, c(1,nrow(tt))),
              aes(x=lon, y=lat),
              colour="red", size=5) +

    geom_jitter(aes(x=lon, y=lat, size = gewicht, colour = as.character(date)),
                alpha = 0.5, width=0.015, height=0.03) +
    scale_size(range = c(1,10)) +

    scale_x_continuous(breaks = seq(floor(xmin), ceiling(xmax), 1)) +
    scale_y_continuous(breaks = seq(floor(2*ymin)/2, ceiling(2*ymax)/2, 0.5)) +
    labs(x = NULL, y = NULL, size = "kg aanvoer/trek", colour="") +
    guides(size = guide_legend(nrow = 1))
    # + facet_wrap(~source)

}

  

```

```{r catchperhaul2, eval=FALSE, fig.align="center", fig.asp=1.0, message=FALSE, warning=FALSE, include=FALSE}

# NOT SHOWN; problem with elog

fig_nums(
  name    = "catchperhaul", level   = 1, display = FALSE,
  caption = "Vangst per trek (kg)")

xlim <- c(xmin,xmax); ylim <- c(ymin,ymax)

t1 <-
  bind_rows(
    tt %>% mutate(source="treklijst", catchtype="totalcatch") %>% rename(data=totalcatch),
    e %>% group_by(vessel, trip, haul, lat, lon) %>% summarise(data=sum(weight, na.rm=TRUE)) %>% mutate(source="elog", catchtype="aanvoer")
  )

t1 %>% 
  ggplot() + 
  theme_publication() +
  theme(
    panel.grid.minor = element_blank(),
    text             = element_text(size=12),
    plot.margin      = unit(c(0,0,0,0), "cm"),
    panel.background = element_rect(fill = "lightskyblue1")
  ) +  

  coord_quickmap(xlim=xlim , ylim=ylim) +

  geom_text(data=rect_df, aes(x=lon, y=lat, label=rect), 
            hjust=0, vjust=0, nudge_x=0.05, nudge_y=0.025, colour="white") +

  geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +

  geom_path (aes(x=lon, y=lat, colour=source)) +
  geom_point(aes(x=lon, y=lat, size = data, shape=catchtype, colour = source), 
              alpha = 0.5) +
  geom_point(data=filter(t1, action=="embarked"), aes(x=lon, y=lat), colour="green", size=5, alpha=0.5) +
  geom_point(data=filter(t1, action=="disembarked"), aes(x=lon, y=lat), colour="red", size=5, alpha=0.5) +

  scale_size(range = c(1,10)) +
  
  scale_x_continuous(breaks = seq(floor(xmin), ceiling(xmax), 1)) +
  scale_y_continuous(breaks = seq(floor(2*ymin)/2, ceiling(2*ymax)/2, 0.5)) +
  labs(x = NULL, y = NULL, size = "totale vangst/trek", colour="") +
  guides(size = guide_legend(nrow = 1)) 
  # + facet_wrap(~source)

```

_`r fig_nums("catchperhaul")`_


<!-- ########################################################################################## -->
<!-- Figure: vangst per soort en klasse --------------------------- -->
<!-- ########################################################################################## -->

```{r vangstpersoortenklasse, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE}

nspecies <- 10

fig_nums(
  name    = "totalcatchperclass", level   = 1, display = FALSE,
  caption = paste0("Totale vangst per trek, soort en klasse (",nspecies," soorten met hoogste vangst)"))


# create list of catch by species by haul. From hm, or sph or else species = "oth"
if (nrow(m) > 0) {
  myspecies <-
    m %>% 
    group_by(soorten) %>% 
    summarise(gewicht = sum(gewicht, na.rm=TRUE)) %>% 
    arrange (desc(gewicht)) %>% 
    slice_head(n=nspecies)
  
  t <-
    m %>% 
    filter(soorten %in% myspecies$soorten) %>% 
    group_by(vessel, trip, haul=haul2, soorten, maat) %>% 
    summarise(gewicht=sum(gewicht))
  
  s <-
    t %>% 
    group_by(vessel, trip, soorten) %>% 
    summarise(totaal=sum(gewicht))  %>% 
    ungroup() %>% 
    arrange(desc(totaal)) %>% 
    dplyr::select(soorten) %>% 
    unlist()

  t %>% 
    ggplot(aes(x=haul, y=gewicht)) +
    theme_publication() +
    geom_bar(aes(fill=soorten), stat="identity") +
    coord_flip() +
    scale_x_reverse(breaks=seq(1, max(t$haul), 1)) +
    scale_y_continuous(breaks=scales::pretty_breaks(n=3)) +
    labs(y="aanvoer (kg)", title="aanvoer per soort en trek") +
    facet_wrap(.~trip, scales="free_x")
  
}

```

_`r fig_nums("totalcatchperclass")`_


<!-- ########################################################################################## -->
<!-- Figure: aanvoer en discards                                                                -->
<!-- ########################################################################################## -->

```{r aanvoerendiscards, eval=FALSE, fig.align="center", fig.asp=1.2, message=FALSE, warning=FALSE, include=FALSE}

# print_option <- FALSE
# 
# # create list of catch by species by haul. From hm, or sph or else species = "oth"
# if (nrow(m) > 0 & sum(!is.na(h$totalcatch))>0) {
# 
#   print_option <- TRUE
#   
#   fig_nums(
#     name    = "aanvoerentotalevangst", level   = 1, display = FALSE,
#     caption = "Vergelijking van aanvoer (Marelec) en totale vangst (schatting)")
#   
#   t1 <-
#     h %>% 
#     group_by(vessel, trip, haul) %>% 
#     summarise(data = sum(totalcatch, na.rm=TRUE)) %>% 
#     mutate(variable = "totale vangst")
# 
#   t2 <-
#     m %>% 
#     group_by(vessel, trip, haul) %>% 
#     summarise(data = sum(gewicht, na.rm=TRUE)) %>% 
#     mutate(variable = "aanvoer")
#     
# 
#   p <-
#     t1 %>% 
#     ggplot(aes(x=haul, y=data)) +
#     theme_publication() +
#     theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
#     geom_bar(aes(fill="totale vangst"),
#              stat="identity", alpha=0.5, width=0.5, just=0.8) +
#     geom_bar(data=t2,
#              aes(fill="aanvoer"),
#              stat="identity", alpha=0.5, width=0.5, just=0.2) +
#     scale_fill_manual(name='',
#                      breaks=c('aanvoer', 'totale vangst'),
#                      values=c('aanvoer'='blue', 'totale vangst'='red')) +
#     scale_x_continuous(breaks=seq(1,max(t1$haul, m$haul, na.rm=TRUE), 1)) +
#     labs(x="trek", y="kg", title="totale vangst en aanvoer per trek") +
#     facet_wrap(~trip, ncol=1)
#     
#   p
#   
#   
# }
# 
# my_caption   <- ifelse(print_option, paste0("_", fig_nums("aanvoerentotalevangst"),"_"), "")

print_option <- FALSE

# create list of catch by species by haul. From hm, or sph or else species = "oth"
if (nrow(m) > 0 & sum(!is.na(h$totalcatch))>0) {

  print_option <- TRUE
  
  fig_nums(
    name    = "aanvoerentotalevangst", level   = 1, display = FALSE,
    caption = "Vergelijking van aanvoer (Marelec) en totale vangst (schatting)")
  
  t1 <-
    h %>% 
    group_by(vessel, trip, date) %>% 
    summarise(data = sum(totalcatch, na.rm=TRUE)) %>% 
    mutate(variable = "totale vangst")

  t2 <-
    m %>% 
    mutate(date = as.Date(datetime)) %>% 
    group_by(vessel, trip, date) %>% 
    summarise(data = sum(gewicht, na.rm=TRUE)) %>% 
    mutate(variable = "aanvoer")
    

  p <-
    t1 %>% 
    ggplot(aes(x=date, y=data)) +
    theme_publication() +
    # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    geom_bar(aes(fill="totale vangst"),
             stat="identity", alpha=0.5, width=0.5, just=0.8) +
    geom_bar(data=t2,
             aes(fill="aanvoer"),
             stat="identity", alpha=0.5, width=0.5, just=0.2) +
    scale_fill_manual(name='',
                     breaks=c('aanvoer', 'totale vangst'),
                     values=c('aanvoer'='blue', 'totale vangst'='red')) +
    # scale_x_continuous(breaks=seq(1,max(t1$haul, m$haul, na.rm=TRUE), 1)) +
    labs(x="trek", y="kg", title="totale vangst en aanvoer per trek") +
    facet_wrap(~trip, ncol=1, scales = "free")
    
  p
  
  
}

my_caption   <- ifelse(print_option, paste0("_", fig_nums("aanvoerentotalevangst"),"_"), "")


```


<!-- ########################################################################################## -->
<!-- Figure: aanvoer per week                                                                   -->
<!-- ########################################################################################## -->

```{r aanvoerperweek, echo=FALSE, fig.align="center", fig.asp=0.8, message=FALSE, warning=FALSE}


# create list of catch by species by week
if (nrow(m) > 0) {
  
  s <-
    elog %>% 
    filter(vessel == setvessel, year == 2023) %>% 
    
    group_by(species) %>% 
    summarise(weight = sum(weight, na.rm=TRUE)) %>% 
    
    arrange (desc(weight)) %>% 
    group_by() %>% 
    slice_head(n=10) %>% 
    left_join(dplyr::select(asfis,
                            species, englishname), 
              by=c("species")) %>% 
    mutate(english_species = paste(englishname, species))
  

  t <-
    elog %>% 
    filter(vessel == setvessel, year == 2023) %>% 
    
    mutate(species = ifelse(species %in% s$species, species, "ZZZ")) %>%

    group_by(week, species) %>% 
    summarise(weight = sum(weight, na.rm=TRUE)) %>% 
    
    left_join(dplyr::select(asfis,
                            species, dutchname, englishname, scientificname), 
              by=c("species")) %>% 
    mutate(
      dutchname   = ifelse(species == "ZZZ", "Overig", dutchname),
      englishname = ifelse(species == "ZZZ", "Other", englishname),
      scientificname = ifelse(species == "ZZZ", "", dutchname),
      english_species = paste(englishname, species)
    ) %>% 
    mutate(english_species = factor(english_species, levels=c(s$english_species, "Other ZZZ")))
  
  t %>% 
    ggplot(aes(x=week, y=weight)) +
    theme_publication() +
    theme(legend.position = "right", legend.direction = "vertical") +
    geom_bar(aes(fill=english_species), 
             stat="identity", position = position_stack(reverse = TRUE), colour="lightgray", linewidth=0.3) +
    labs(y="aanvoer (kg)", title=paste(setvessel, "aanvoer per soort en week"), fill="") +
    scale_fill_brewer(palette="Paired")
  
}


```
