---
output:
  word_document:
    reference_docx: ../report_template.dotx
---


```{r setup, include=FALSE}

# =====================================================================================================
# Global Fishing Watch effort analyse
#
# 03/10/2023
# =====================================================================================================

knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

options(dplyr.summarise.inform = FALSE)

# Reset lists
rm(list=ls())

download_data <- TRUE

# Libraries
library(rmarkdown)                   # note: requires knitr 1.21

require(tidyverse, quietly=TRUE)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
require(lubridate, quietly=TRUE)     # data handling
require(reshape2, quietly=TRUE)      # reshaping data; e.g. cast
require(RColorBrewer)                # colour schemes
library(viridis)
library(pander)
library(patchwork)
library(gridExtra)
library(sf)

library(captioner)                   # for captions
tab_nums <- captioner::captioner(prefix = "Table " , levels=1, type=c("n"), infix=".")
fig_nums <- captioner::captioner(prefix = "Figure ", levels=1, type=c("n"), infix=".")

library(gfwr)
key <- gfw_auth()

source("../../prf/R/my utils.r")
source("../../mptools/R/get_onedrive.r")
source("../R/FLYSHOOT utils.r")

spatialdir <- "C:/DATA/RDATA"
load(file.path(spatialdir, "world_mr_df.RData"))
load(file.path(spatialdir, "world_mr_sf.RData"))

fao_sf          <- loadRData(file.path(spatialdir, "fao_sf.RData"))
fao_sf_division <- fao_sf %>% filter(F_LEVEL=="DIVISION") %>% dplyr::select(F_DIVISION) %>% rename(division = F_DIVISION) %>% mutate(division=toupper(division))

df <- data.frame(stringsAsFactors = FALSE)
for (y in 2012:2022) {
  for (a in c("27.4.A","27.4.B","27.4.C","27.7.D","27.7.E")) {
    
    print(paste(y,a))
    
    mydate    <- paste0(y,'-01-01,',y,'-12-31')
    mysf      <- fao_sf_division %>% filter(division == a) %>% st_make_valid()
    mygeojson <- paste0('{"geojson":', geojsonsf::sf_geojson(mysf) ,'}') 

    df <- bind_rows(
      df,
      get_raster(
         spatial_resolution  = 'high',
         temporal_resolution = 'yearly',
         group_by            = 'flagAndGearType',
         date_range          = mydate,
         region              = mygeojson,
         region_source       = 'user_json',
         key                 = key) %>%
      lowcase() %>% 
      rename(year=timerange) %>% 
      mutate(division = a)
    )
  }
}

save(df, file=file.path(spatialdir, "gfw_effort_flagandgeartype.RData"))

# # use EEZ function to get EEZ code of Cote d'Ivoire
# code_eez <- get_region_id(region_name = 'NLD', region_source = 'eez', key = key) %>% filter(label=="Netherlands")
# 
# t <- get_raster(spatial_resolution = 'high',
#            temporal_resolution = 'yearly',
#            group_by = 'flagAndGearType',
#            date_range = '2021-01-01,2021-10-01',
#            region = code_eez$id,
#            region_source = 'eez',
#            key = key) %>%
#   lowcase()




```


# Global Fishing Watch data of the flyshoot and twinrig fishery in the North Sea and Channel 

M.A. Pastoors

Date: `r format(Sys.time(), '%d/%m/%Y')`


Introduction

GFW data
GFWR package
Method: 
get_event(
  event_type='fishing',
  vessel = trawler_ids, 
  start_date = "2012-01-01", 
  end_date = "2023-12-31",
  key = key) 

Vessels included in the analysis

```{r echo=FALSE, fig.asp=1.2, message=FALSE, warning=FALSE, comment=NA}

r %>% 
  relocate(vessel, vesselname) %>% 
  arrange(flag, vessel, vesselname) %>% 
  group_by(flag) %>% 
  do(add_row(., .after=0)) %>%

  pandoc.table(.,
             style = "simple",
             split.tables=400, 
             justify = "left",
             missing=".",
             round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))



```
\newpage


\newpage

Effort (hauls) by country

```{r echo=FALSE, fig.asp=1.2, message=FALSE, warning=FALSE, comment=NA}

df %>% 
  # filter(grepl("METANOIA", vesselname)|grepl("MDV", name)) %>% 
  # filter(vessel=="MDV2") %>% 
  # filter(year==2023) %>% View()
  filter(year < 2023) %>% 
  
  ggplot(aes(x=lon, y=lat)) +
  theme_publication() +
  # coord_quickmap(xlim=c(-3,5) , ylim=c(49.5,53)) +
  coord_quickmap(xlim=c(-3,10) , ylim=c(49.5,58)) +
  geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +
  geom_point(aes(colour=flag), size=0.5, alpha=0.2) +
  # geom_point(aes(colour=vesselname), size=0.5, alpha=0.5) +
  facet_wrap(~year, ncol=3)
  # facet_wrap(~week, ncol=10)


```

\newpage

Number of hauls per small rectangle (0.05 deg lon x 0.025 deg lat) for selected years and flag. (Note: Germany not shown)

```{r echo=FALSE, fig.asp=1.0, message=FALSE, warning=FALSE, comment=NA}

dx = 0.05; dy = 0.025

bb <- sf::st_bbox(t)

t %>% 

  ggplot() +
  theme_publication() +
  theme(plot.margin = margin(1,1,1,1, "mm")) +
  theme(plot.title = element_text(hjust = 0.0)) +
  # theme(legend.position = "none") +
  
  geom_sf(data=world_mr_sf, 
          font = "Arial") +
  # geom_sf(aes(fill=landing_interval), 
  #         font = "Arial") +
  
  geom_rect(aes(xmin=lon, ymin=lat, xmax=lon+dx, ymax=lat+dy, fill=effort_interval), linewidth=0.2) +
  scale_fill_viridis(option = "plasma", direction = -1, discrete=TRUE) +
  
  # coord_sf(xlim=xlim, ylim=ylim) + 
  coord_sf(xlim=c(bb$xmin, bb$xmax), ylim=c(bb$ymin, bb$ymax)) + 
  
  guides(fill=guide_legend(nrow = 1)) + 
  labs(x="",y="", title="Nhauls by rectangle") +
  facet_grid(flag~year)
  

```

\newpage

Apparent fishing effort (all gears)

```{r echo=FALSE, fig.asp=0.8, message=FALSE, warning=FALSE, comment=NA}

dx = 0.01; dy = 0.01

t <-
  df %>% 
  group_by(year, lat, lon) %>% 
  summarise(apparentfishinghours = sum(apparentfishinghours, na.rm=TRUE)) %>% 
  ungroup() %>% 
  mutate(., effort_interval = cut(apparentfishinghours, 
                                  scales::trans_breaks("sqrt", function(x) x ^ 2)(c(0, max(apparentfishinghours, na.rm=TRUE))), 
                               dig.lab=10 ) ) %>% 
  
  sf::st_as_sf(coords = c("lon", "lat"), crs = 4326, remove = FALSE)  


bb <- sf::st_bbox(t)

t %>% 

  ggplot() +
  theme_publication() +
  theme(plot.margin = margin(1,1,1,1, "mm")) +
  theme(plot.title = element_text(hjust = 0.0)) +
  # theme(legend.position = "none") +
  
  geom_sf(data=world_mr_sf) +

  geom_rect(aes(xmin=lon, ymin=lat, xmax=lon+dx, ymax=lat+dy, fill=effort_interval), linewidth=0.2) +
  scale_fill_viridis(option = "plasma", direction = -1, discrete=TRUE) +
  
  # coord_sf(xlim=xlim, ylim=ylim) + 
  coord_sf(xlim=c(bb$xmin, bb$xmax), ylim=c(bb$ymin, bb$ymax)) + 
  
  guides(fill=guide_legend(nrow = 1)) + 
  labs(x="",y="", title="Apparent fishing effort (all gears)") +
  facet_wrap(~year)
  

```

\newpage

Apparent fishing effort (other seines, 4c and 7d)

```{r echo=FALSE, fig.asp=0.8, message=FALSE, warning=FALSE, comment=NA}

dx = 0.01; dy = 0.01

t <-
  df %>% 
  filter(geartype == "other_seines") %>% 
  filter(division %in% c("27.4.C","27.7.D")) %>% 
  group_by(year, lat, lon) %>% 
  summarise(apparentfishinghours = sum(apparentfishinghours, na.rm=TRUE)) %>% 
  ungroup() %>% 
  mutate(., effort_interval = cut(apparentfishinghours, 
                                  scales::trans_breaks("sqrt", function(x) x ^ 2)(c(0, max(apparentfishinghours, na.rm=TRUE))), 
                               dig.lab=10 ) ) %>% 
  sf::st_as_sf(coords = c("lon", "lat"), crs = 4326, remove = FALSE)  

t2 <-
  t %>% 
  group_by(year) %>% 
  summarise(apparentfishinghours = sum(apparentfishinghours, na.rm=TRUE)) 

bb <- sf::st_bbox(t)

# maps
t %>% 
  ggplot() +
  theme_publication() +
  theme(plot.margin = margin(1,1,1,1, "mm")) +
  theme(plot.title = element_text(hjust = 0.0)) +
  # theme(legend.position = "none") +
  
  geom_sf(data=world_mr_sf) +

  geom_rect(aes(xmin=lon, ymin=lat, xmax=lon+dx, ymax=lat+dy, fill=effort_interval), linewidth=0.2) +
  scale_fill_viridis(option = "plasma", direction = -1, discrete=TRUE) +
  
  # coord_sf(xlim=xlim, ylim=ylim) + 
  coord_sf(xlim=c(bb$xmin, bb$xmax), ylim=c(bb$ymin, bb$ymax)) + 
  
  guides(fill=guide_legend(nrow = 1)) + 
  labs(x="",y="", title="Apparent fishing effort (other seines)") +
  facet_wrap(~year)
  
# time trends
t2 %>% 
  ggplot(aes(x=year, y=apparentfishinghours)) +
  theme_publication() +
  theme(plot.margin = margin(1,1,1,1, "mm")) +
  theme(plot.title = element_text(hjust = 0.0)) +
  geom_point() +
  geom_line() +
  labs(x="",y="", title="Apparent fishing effort (other seines)") +
  scale_x_continuous(breaks=unique(t$year)) +
  expand_limits(y=0)


```

\newpage

Apparent fishing effort (other seines, 4c and 7d, by country)

```{r echo=FALSE, fig.asp=0.8, message=FALSE, warning=FALSE, comment=NA}

dx = 0.01; dy = 0.01

t <-
  df %>% 
  filter(geartype == "other_seines") %>% 
  filter(division %in% c("27.4.C","27.7.D")) %>% 
  group_by(year, flag, lat, lon) %>% 
  summarise(apparentfishinghours = sum(apparentfishinghours, na.rm=TRUE)) %>% 
  ungroup() %>% 
  mutate(., effort_interval = cut(apparentfishinghours, 
                                  scales::trans_breaks("sqrt", function(x) x ^ 2)(c(0, max(apparentfishinghours, na.rm=TRUE))), 
                               dig.lab=10 ) ) %>% 
  sf::st_as_sf(coords = c("lon", "lat"), crs = 4326, remove = FALSE)  

t2 <-
  t %>% 
  group_by(year, flag) %>% 
  summarise(apparentfishinghours = sum(apparentfishinghours, na.rm=TRUE)) 

bb <- sf::st_bbox(t)

# maps
t %>% 
  ggplot() +
  theme_publication() +
  theme(plot.margin = margin(1,1,1,1, "mm")) +
  theme(plot.title = element_text(hjust = 0.0)) +
  # theme(legend.position = "none") +
  
  geom_sf(data=world_mr_sf) +

  geom_rect(aes(xmin=lon, ymin=lat, xmax=lon+dx, ymax=lat+dy, fill=effort_interval), linewidth=0.2) +
  scale_fill_viridis(option = "plasma", direction = -1, discrete=TRUE) +
  
  # coord_sf(xlim=xlim, ylim=ylim) + 
  coord_sf(xlim=c(bb$xmin, bb$xmax), ylim=c(bb$ymin, bb$ymax)) + 
  
  guides(fill=guide_legend(nrow = 1)) + 
  labs(x="",y="", title="Apparent fishing effort (other seines)") +
  facet_grid(year~flag)
  
# time trends
t2 %>% 
  ggplot(aes(x=year, y=apparentfishinghours)) +
  theme_publication() +
  theme(plot.margin = margin(1,1,1,1, "mm")) +
  theme(plot.title = element_text(hjust = 0.0)) +
  geom_point(aes(colour=flag)) +
  geom_line(aes(colour=flag)) +
  labs(x="",y="", title="Apparent fishing effort (other seines)") +
  scale_x_continuous(breaks=unique(t$year)) +
  expand_limits(y=0)


```