---
output:
  word_document:
    reference_docx: ../report_template.dotx
---


```{r setup, include=FALSE}

# =====================================================================================================
# Global Fishing Watch data analyse
#
# 22/08/2023
# =====================================================================================================

knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

options(dplyr.summarise.inform = FALSE)

# Reset lists
rm(list=ls())

download_data <- TRUE

# Libraries
library(rmarkdown)                   # note: requires knitr 1.21

require(tidyverse, quietly=TRUE)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
require(lubridate, quietly=TRUE)     # data handling
require(reshape2, quietly=TRUE)      # reshaping data; e.g. cast
require(RColorBrewer)                # colour schemes
library(viridis)
library(pander)
library(patchwork)
library(gridExtra)
library(sf)

library(captioner)                   # for captions
tab_nums <- captioner::captioner(prefix = "Table " , levels=1, type=c("n"), infix=".")
fig_nums <- captioner::captioner(prefix = "Figure ", levels=1, type=c("n"), infix=".")

library(gfwr)
key <- gfw_auth()

source("../../prf/R/my utils.r")
source("../../mptools/R/get_onedrive.r")
source("../R/FLYSHOOT utils.r")

spatialdir <- "C:/DATA/RDATA"
load(file.path(spatialdir, "world_mr_df.RData"))
load(file.path(spatialdir, "world_mr_sf.RData"))

fao_sf          <- loadRData(file.path(spatialdir, "fao_sf.RData"))
fao_sf_division <- fao_sf %>% filter(F_LEVEL=="DIVISION") %>% dplyr::select(F_DIVISION) %>% rename(division = F_DIVISION)

# flyshoot info
onedrive <- get_onedrive(team="Martin Pastoors", site="FLYSHOOT - General")
haul <-
  loadRData(file.path(onedrive, "rdata", "haul.RData")) %>% 
  arrange(vessel, week, haultime) %>% 
  group_by(vessel, week) %>% 
  mutate(haul = row_number())

# ==========================================
# read vessels from excel file
# ==========================================

# r <- 
#   readxl::read_excel(path=file.path(onedrive, "data","flyshoot vessels update EvL.xlsx")) %>% 
#   mutate(mmsi = as.character(mmsi))
r <- data.frame(vessel="kw145", vesselname="Catharina", mmsi = "246464000", stringsAsFactors = FALSE)


if (download_data) {
  
  trawlers <- 
    gfwr::get_vessel_info(
      query = paste0("mmsi=",paste(r$mmsi, collapse=" OR mmsi=")),
      search_type = "advanced",
      dataset = "fishing_vessel",
      key = key
  )
  
  # Collapse vessel ids into a commas separated list to pass to Events API
  trawler_ids <- paste0(trawlers$id, collapse = ',')
    
  df <-
    get_event(event_type='fishing',
              vessel = trawler_ids,
              # start_date = "2012-01-01", end_date = "2023-12-31",
              start_date = "2022-03-01", 
              end_date   = "2023-02-28",
              key = key) %>%
    mutate(year = lubridate::year(start),
           month = lubridate::month(start),
           week  = lubridate::week(start),
           date = as.Date(start)) %>%
    unnest_wider(regions, names_repair = "unique") %>%
    unnest_wider(fao, names_sep = "_") %>%
    unnest_wider(eez, names_sep = "_") %>%
    unnest_wider(vessel, names_repair = "unique") %>%
    unnest_wider(event_info, names_repair = "unique") %>%
    unnest_wider(distances, names_repair = "unique") %>%
    dplyr::select(-mpa, -authorizations, -rfmo, -majorFao) %>%
    left_join(r, by=c("ssvid"="mmsi")) %>%
    arrange(vesselname, year, week, end) %>%
    group_by(vesselname, year, week) %>%
    mutate(haul = row_number()) %>%
  
    sf::st_as_sf(coords = c("lon", "lat"), crs = 4326, stringsAsFactors = FALSE, remove = FALSE) %>%
    sf::st_join(., fao_sf_division, join = st_within) %>%
    sf::st_drop_geometry()
  
  gfw <- df
  save(gfw, file = file.path(onedrive, "gfw kw145.RData"))
  
}

# df <- 
#   # df %>% 
#   loadRData(file = file.path(onedrive, "gfw kw145.RData")) %>% 
#   filter(division %in% c("27.4.a","27.4.b","27.4.c","27.7.d","27.7.e")) %>% 
#   filter(active=="YES") %>% 
#   rename(flag=flag.x) %>% 
#   dplyr::select(-gt) %>% 
#   left_join(dplyr::select(r,
#                           vessel, loa, breadth, gt), 
#             by="vessel")

# df %>% filter(grepl("METANOIA", name)) %>% View() 
# trawlers %>% filter(flag=="NLD") %>% View()


xmin <- floor(2 * (min(df$lon, na.rm=TRUE)-0.5))/2;
xmax <- ceiling(2 * (max(df$lon, na.rm=TRUE)+0.5))/2;
ymin <- floor(2 * (min(df$lat, na.rm=TRUE)-0.5))/2;
ymax <- ceiling(2* (max(df$lat, na.rm=TRUE)+0.5))/2

xlim=c(xmin, xmax); ylim=c(ymin, ymax)
# xlim=c(-3,10); ylim=c(49.5,58)



```


# Global Fishing Watch data of the KW145 

M.A. Pastoors

Date: `r format(Sys.time(), '%d/%m/%Y')`


Introduction

GFW data
GFWR package
Method: 
get_event(
  event_type='fishing',
  vessel = trawler_ids, 
  start_date = "2012-01-01", 
  end_date = "2023-12-31",
  key = key) 

Vessels included in the analysis

```{r echo=FALSE, fig.asp=1.2, message=FALSE, warning=FALSE, comment=NA}

df %>% 
  group_by(vessel, year, week) %>% 
  summarise(nhauls = n()) %>% 

  pandoc.table(.,
             style = "simple",
             split.tables=400, 
             justify = "left",
             missing=".",
             round=c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))

df %>% 
  group_by(vessel, year, week) %>% 
  summarise(nhauls = n()) %>% 
  ggplot(aes(x=paste(year, week), y=nhauls)) +
  theme_publication() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_bar(stat="identity") +
  labs(x="year/week")





```
\newpage


```{r echo=FALSE, fig.asp=0.4, message=FALSE, warning=FALSE, comment=NA}
df %>% 
  ggplot(aes(x=lon, y=lat)) +
  theme_publication() +
  coord_quickmap(xlim=xlim, ylim=ylim) +
  geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +
  geom_point(size=1, alpha=0.6, shape=1) +
  facet_wrap(~paste(year,week), ncol=12)

```

\newpage

Checking on number of hauls per week identified from Jaczon selfsampling and derived from GFW (AI-model). 

Conclusion: GFW identifies fewer hauls than actually carried out in the flyshoot fishery (no information on twinrig included in this example). 

```{r echo=FALSE, fig.asp=0.4, message=FALSE, warning=FALSE, comment=NA}

comb %>% 
  group_by(source, vessel, week) %>% 
  summarise(nhauls = n()) %>% 
  group_by(source, vessel) %>% 
  mutate(mean = mean(nhauls)) %>% 
  
  ggplot(aes(x=week, y=nhauls)) +
  geom_point(aes(colour=source), size=1) +
  geom_hline(aes(colour=source, yintercept=mean)) +
  expand_limits(y=0) +
  facet_wrap(~vessel)

```


\newpage

Effort (hauls) by country

```{r echo=FALSE, fig.asp=1.2, message=FALSE, warning=FALSE, comment=NA}

df %>% 
  # filter(grepl("METANOIA", vesselname)|grepl("MDV", name)) %>% 
  # filter(vessel=="MDV2") %>% 
  # filter(year==2023) %>% View()
  filter(year < 2023) %>% 
  
  ggplot(aes(x=lon, y=lat)) +
  theme_publication() +
  # coord_quickmap(xlim=c(-3,5) , ylim=c(49.5,53)) +
  coord_quickmap(xlim=c(-3,10) , ylim=c(49.5,58)) +
  geom_polygon(data=world_mr_df, aes(x=long, y=lat, group=group), fill = "grey75") +
  geom_point(aes(colour=flag), size=0.5, alpha=0.2) +
  # geom_point(aes(colour=vesselname), size=0.5, alpha=0.5) +
  facet_wrap(~year, ncol=3)
  # facet_wrap(~week, ncol=10)


```

\newpage

Number of hauls per small rectangle (0.05 deg lon x 0.025 deg lat) for selected years and flag. (Note: Germany not shown)

```{r echo=FALSE, fig.asp=1.0, message=FALSE, warning=FALSE, comment=NA}

dx = 0.05; dy = 0.025

t <-
  df %>% 
  filter(year %in% c(2013, 2016, 2019, 2022)) %>% 
  filter(division %in% c("27.4.c", "27.7.d")) %>% 
  filter(flag != "DEU") %>% 
  
  mutate(rect   = encode_zchords(lon, lat, dx = dx, dy = dy)) %>% 
  group_by(flag, year, rect) %>% 
  summarise(
    nhauls        = n()) %>% 
  separate(rect, c("lon", "lat"), sep = ":", convert = TRUE, remove = FALSE) %>% 
  ungroup() %>% 

  mutate(., effort_interval = cut(nhauls, 
                                  scales::trans_breaks("sqrt", function(x) x ^ 2)(c(0, max(nhauls, na.rm=TRUE))), 
                               dig.lab=10 ) ) %>% 
  
  sf::st_as_sf(coords = c("lon", "lat"), crs = 4326, remove = FALSE)  


tmax <- 100*ceiling(max(t$nhauls, na.rm=TRUE)/100)

bb <- sf::st_bbox(t)

t %>% 

  ggplot() +
  theme_publication() +
  theme(plot.margin = margin(1,1,1,1, "mm")) +
  theme(plot.title = element_text(hjust = 0.0)) +
  # theme(legend.position = "none") +
  
  geom_sf(data=world_mr_sf, 
          font = "Arial") +
  # geom_sf(aes(fill=landing_interval), 
  #         font = "Arial") +
  
  geom_rect(aes(xmin=lon, ymin=lat, xmax=lon+dx, ymax=lat+dy, fill=effort_interval), linewidth=0.2) +
  scale_fill_viridis(option = "plasma", direction = -1, discrete=TRUE) +
  
  # coord_sf(xlim=xlim, ylim=ylim) + 
  coord_sf(xlim=c(bb$xmin, bb$xmax), ylim=c(bb$ymin, bb$ymax)) + 
  
  guides(fill=guide_legend(nrow = 1)) + 
  labs(x="",y="", title="Nhauls by rectangle") +
  facet_grid(flag~year)
  

```

\newpage

Number of hauls per small rectangle (0.05 deg lon x 0.025 deg lat) for selected years and gear types 

```{r echo=FALSE, fig.asp=0.8, message=FALSE, warning=FALSE, comment=NA}

dx = 0.05; dy = 0.025

t <-
  df %>% 
  filter(year %in% c(2013, 2016, 2019, 2022)) %>% 
  filter(division %in% c("27.4.c", "27.7.d")) %>% 
  mutate(rect   = encode_zchords(lon, lat, dx = dx, dy = dy)) %>% 
  group_by(gear, year, rect) %>% 
  summarise(
    nhauls        = n()) %>% 
  separate(rect, c("lon", "lat"), sep = ":", convert = TRUE, remove = FALSE) %>% 
  ungroup() %>% 

  mutate(., effort_interval = cut(nhauls, 
                                  scales::trans_breaks("sqrt", function(x) x ^ 2)(c(0, max(nhauls, na.rm=TRUE))), 
                               dig.lab=10 ) ) %>% 
  
  sf::st_as_sf(coords = c("lon", "lat"), crs = 4326, remove = FALSE)  


tmax <- 100*ceiling(max(t$nhauls, na.rm=TRUE)/100)

bb <- sf::st_bbox(t)

t %>% 

  ggplot() +
  theme_publication() +
  theme(plot.margin = margin(1,1,1,1, "mm")) +
  theme(plot.title = element_text(hjust = 0.0)) +
  # theme(legend.position = "none") +
  
  geom_sf(data=world_mr_sf, 
          font = "Arial") +
  # geom_sf(aes(fill=landing_interval), 
  #         font = "Arial") +
  
  geom_rect(aes(xmin=lon, ymin=lat, xmax=lon+dx, ymax=lat+dy, fill=effort_interval), linewidth=0.2) +
  scale_fill_viridis(option = "plasma", direction = -1, discrete=TRUE) +
  
  # coord_sf(xlim=xlim, ylim=ylim) + 
  coord_sf(xlim=c(bb$xmin, bb$xmax), ylim=c(bb$ymin, bb$ymax)) + 
  
  guides(fill=guide_legend(nrow = 1)) + 
  labs(x="",y="", title="Nhauls by rectangle") +
  facet_grid(gear~year)
  

```

