---
title: "GFW Fishing Effort Analysis — ICES Divisions 27.4.C, 27.7.D, 27.7.E"
author: "Martin Pastoors"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  word_document:
    reference_docx: "../report_template.dotx"
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 8,
  fig.height = 5,
  dpi        = 150
)
```

```{r libraries}
library(tidyverse)
library(lubridate)
library(sf)
library(spatstat)
library(rnaturalearth)
library(rnaturalearthdata)
library(ineq)
```

```{r settings}
# ---- Paths ----
spatialdir  <- "C:/Users/MartinPastoors/OneDrive - Martin Pastoors/DATA/RDATA"
flyshootdir <- "C:/Users/MartinPastoors/Martin Pastoors/FLYSHOOT - General/data"

# ---- Settings ----
study_years     <- 2013:2025
study_divisions <- c("27.4.C", "27.7.D", "27.7.E")
threshold_pct   <- 1
xlim <- c(-7, 8)
ylim <- c(48, 54)

flag_colours <- c(
  "BEL"   = "#4E79A7",
  "NLD"   = "#F28E2B",
  "GBR"   = "#59A14F",
  "FRA"   = "#E15759",
  "DEU"   = "#76B7B2",
  "Other" = "#BAB0AC"
)

size_colours <- c(
  "S1 <100 GT"     = "#1D9E75",
  "S2 100-300 GT"  = "#BA7517",
  "S3 300-600 GT"  = "#7F77DD",
  "S4 600-1200 GT" = "#D85A30",
  "S5 >1200 GT"    = "#A32D2D"
)
```

```{r helpers}
load_object <- function(filepath, object_name) {
  local({
    e <- new.env()
    load(filepath, envir = e)
    e[[object_name]]
  })
}

apply_flag_other <- function(df, flag_col = "vessel_flag",
                             days_col = "fishing_days") {
  df %>%
    group_by(across(all_of(flag_col))) %>%
    mutate(total_flag_days = sum(.data[[days_col]])) %>%
    ungroup() %>%
    mutate(
      total_days = sum(.data[[days_col]]),
      flag_pct   = total_flag_days / total_days * 100,
      across(all_of(flag_col),
             ~ if_else(flag_pct < threshold_pct, "Other", .x))
    ) %>%
    rename_with(~ str_replace(., "fishing_days", days_col),
                any_of(c("fishing_days", "total_flag_days", "total_days")))
}
```

```{r load-data}
# NOTE: The following three .RData files must be present in flyshootdir and
# spatialdir before knitting. They are produced by steps 3, 6 and 7 of the
# full GFW extraction pipeline (Extract_GFW_effort_v8.R).
#
# ⚠️  MISSING DATA SOURCES — see section "Data availability" below.

fishing_days_study <- load_object(
  file.path(flyshootdir, "gfw_s7_effort.RData"), "fishing_days_study")

events_classified_fr <- load_object(
  file.path(flyshootdir, "gfw_s6_zones.RData"), "events_classified_fr")

vessel_meta <- load_object(
  file.path(flyshootdir, "gfw_s3_vessel_meta.RData"), "vessel_meta")

# Spatial layers
world <- ne_countries(scale = "medium", returnclass = "sf")

fao_sf <- load_object(file.path(spatialdir, "fao_sf.RData"), "fao_sf")
fao_sf_division <- fao_sf %>%
  filter(F_LEVEL == "DIVISION") %>%
  dplyr::select(division = F_DIVISION) %>%
  mutate(division = toupper(division)) %>%
  filter(division %in% study_divisions) %>%
  group_by(division) %>%
  summarise(geometry = st_union(geometry)) %>%
  st_make_valid()
# rm(fao_sf)
```

---

# Introduction

This report summarises fishing effort in ICES divisions 27.4.C (southern North Sea), 27.7.D (eastern English Channel) and 27.7.E (western English Channel) for the period `r min(study_years)`–`r max(study_years)`, derived from Global Fishing Watch (GFW) AIS-based vessel tracking data.

The analysis covers total fishing days by flag state, ICES division, gear type, vessel size class and spatial distribution. Effort concentration in ICES 27.7.D is additionally quantified using kernel density estimation and grid-cell Gini coefficients.

---

# Data availability

> **⚠️ Missing data source — `fao_sf.RData`**  
> The ICES FAO division shapefile is loaded from a local path (`spatialdir`). This file is not bundled with the code. It must be present at the path defined in `spatialdir` before knitting. The object should contain an `sf` data frame with at minimum the fields `F_LEVEL` and `F_DIVISION`, consistent with the official ICES/FAO spatial layers.

The three RData files produced by the full pipeline (`gfw_s3_vessel_meta.RData`, `gfw_s6_zones.RData`, `gfw_s7_effort.RData`) must likewise be present in `flyshootdir`. These are generated by steps 3, 6 and 7 of `Extract_GFW_effort_v8.R` and are not included in this standalone script.

---

# Methods

Fishing activity data were retrieved via the Global Fishing Watch (GFW) API (`gfwr` v3.0) using the Events endpoint, which returns individual fishing events at the vessel level. Vessels were identified through a discovery step that queried all vessels with recorded fishing activity inside the three study divisions in reference years (2013, 2018, 2023, 2025). Fishing events were subsequently extracted for the full period 2013–2025.

Events were spatially filtered to marine-only areas (internal waters excluded) and joined to the official ICES FAO division polygons. Zone classification distinguished coastal (0–12 NM) from offshore (>12 NM) areas and sub-zones within the French EEZ. Fishing effort was aggregated to the vessel-day level using a majority-rule deduplication approach: where a vessel generated multiple events on the same calendar day, the event type accounting for the majority of time was retained.

Vessel metadata (flag state, gear type, gross tonnage) were obtained from the GFW vessel registry. Gross tonnage was imputed for vessels with missing registry entries using a length-based regression model calibrated on vessels with known GT values. Vessels were assigned to five size classes based on GT: S1 (<100 GT), S2 (100–300 GT), S3 (300–600 GT), S4 (600–1200 GT) and S5 (>1200 GT).

Flag states contributing less than `r threshold_pct`% of total fishing days across the study period are pooled into an "Other" category in all plots.

---

# Results

## 1. Effort by flag state

### 1.1 Total fishing days by year and flag state (all divisions combined)

```{r plot1, fig.cap="Figure 1. Total fishing days at sea per year, stacked by flag state. Flag states contributing less than 1% of total effort are pooled as 'Other'."}
fishing_days_study %>%
  filter(year %in% study_years) %>%
  group_by(year, vessel_flag) %>%
  summarise(fishing_days = sum(fishing_days), .groups = "drop") %>%
  apply_flag_other() %>%
  group_by(year, vessel_flag) %>%
  summarise(fishing_days = sum(fishing_days), .groups = "drop") %>%
  ggplot(aes(x = year, y = fishing_days, fill = vessel_flag)) +
  theme_bw() + geom_col() +
  scale_fill_manual(values = flag_colours, na.value = "grey70") +
  scale_x_continuous(breaks = study_years) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(x = "", y = "Fishing days", fill = "Flag",
       title = "Fishing days at sea by year and country")
```

Fishing effort is expressed as the total number of vessel-days within the three study divisions. The stacked bars show each flag state's contribution, enabling comparison of both the overall trend and the relative importance of individual fleets across the time series.

---

### 1.2 Fishing days by year, flag state and ICES division

```{r plot2, fig.height=6, fig.cap="Figure 2. Fishing days per year by flag state, faceted by ICES division. This allows comparison of effort levels and fleet composition between divisions."}
fishing_days_study %>%
  filter(year %in% study_years, !is.na(division)) %>%
  group_by(year, vessel_flag, division) %>%
  summarise(fishing_days = sum(fishing_days), .groups = "drop") %>%
  apply_flag_other() %>%
  group_by(year, vessel_flag, division) %>%
  summarise(fishing_days = sum(fishing_days), .groups = "drop") %>%
  ggplot(aes(x = year, y = fishing_days, fill = vessel_flag)) +
  theme_bw() + geom_col() +
  scale_fill_manual(values = flag_colours, na.value = "grey70") +
  scale_x_continuous(breaks = study_years) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(x = "", y = "Fishing days", fill = "Flag",
       title = "Fishing days by year, country and ICES division") +
  facet_wrap(~ division)
```

Faceting by division reveals differences in fleet composition and effort trends between the southern North Sea (27.4.C), eastern Channel (27.7.D) and western Channel (27.7.E). Scales are shared across panels to allow direct visual comparison of absolute effort levels.

---

## 2. Effort by gear type

### 2.1 Fishing days by year and gear type

```{r plot5, fig.height=6, fig.cap="Figure 3. Fishing days per year by gear type, faceted by ICES division. Gear types contributing less than 1% of total effort are pooled as 'Other'. Free y-axis scales are used to highlight within-division patterns."}
fishing_days_study %>%
  filter(year %in% study_years, !is.na(division), !is.na(gear)) %>%
  group_by(year, gear, division) %>%
  summarise(fishing_days = sum(fishing_days), .groups = "drop") %>%
  group_by(gear) %>%
  mutate(total_gear_days = sum(fishing_days)) %>%
  ungroup() %>%
  mutate(gear = if_else(total_gear_days / sum(fishing_days) * 100
                        < threshold_pct, "Other", gear)) %>%
  group_by(year, gear, division) %>%
  summarise(fishing_days = sum(fishing_days), .groups = "drop") %>%
  ggplot(aes(x = year, y = fishing_days, fill = gear)) +
  theme_bw() + geom_col() +
  scale_fill_brewer(palette = "Set2") +
  scale_x_continuous(breaks = study_years) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(x = "", y = "Fishing days", fill = "Gear type",
       title = "Fishing days by year and gear type") +
  facet_wrap(~ division, scales = "free_y")
```

Gear type is classified by GFW based on the vessel's AIS behaviour and supplemented by registry information. The free y-axis scales allow gear-type patterns within each division to be seen clearly, while the faceted layout enables cross-division comparison of dominant gear groups.

---

## 3. Effort weighted by vessel size (GT fishing days)

### 3.1 GT fishing days by year and flag state

```{r plot6, fig.cap="Figure 4. GT fishing days (fishing days weighted by vessel gross tonnage) per year by flag state. This metric gives greater weight to larger vessels and is a proxy for capacity-weighted effort."}
fishing_days_study %>%
  filter(year %in% study_years, !is.na(gt_final)) %>%
  group_by(year, vessel_flag) %>%
  summarise(gt_fishing_days = sum(gt_fishing_days, na.rm = TRUE),
            .groups = "drop") %>%
  apply_flag_other(days_col = "gt_fishing_days") %>%
  group_by(year, vessel_flag) %>%
  summarise(gt_fishing_days = sum(gt_fishing_days), .groups = "drop") %>%
  ggplot(aes(x = year, y = gt_fishing_days, fill = vessel_flag)) +
  theme_bw() + geom_col() +
  scale_fill_manual(values = flag_colours, na.value = "grey70") +
  scale_x_continuous(breaks = study_years) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(x = "", y = "GT fishing days", fill = "Flag",
       title = "GT fishing days by year and country")
```

GT fishing days are calculated as the product of a vessel's gross tonnage and its number of fishing days. This capacity-weighted metric reflects the combined influence of effort and fleet size, highlighting differences in fleet structure that may not be apparent from unweighted fishing days alone.

---

## 4. Effort by vessel size class

### 4.1 Fishing days by year and size class

```{r plot7, fig.cap="Figure 5. Fishing days per year by vessel size class. Vessels with unknown GT are excluded."}
fishing_days_study %>%
  filter(year %in% study_years, !is.na(size_class),
         size_class != "Unknown") %>%
  group_by(year, size_class) %>%
  summarise(fishing_days = sum(fishing_days), .groups = "drop") %>%
  ggplot(aes(x = year, y = fishing_days, fill = size_class)) +
  theme_bw() + geom_col() +
  scale_fill_manual(values = size_colours) +
  scale_x_continuous(breaks = study_years) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(x = "", y = "Fishing days", fill = "Size class",
       title = "Fishing days by year and vessel size class")
```

Vessels are grouped into five size classes based on gross tonnage (GT). Size class S1 (<100 GT) corresponds to small coastal vessels; S5 (>1200 GT) to the largest ocean-going fishing vessels. Changes in the size-class composition over time reflect structural shifts in the fishing fleets operating in the study area.

---

### 4.2 Fishing days by year, size class and ICES division

```{r plot8, fig.height=6, fig.cap="Figure 6. Fishing days per year by vessel size class, faceted by ICES division. Vessels with unknown GT are excluded."}
fishing_days_study %>%
  filter(year %in% study_years, !is.na(size_class),
         size_class != "Unknown", !is.na(division)) %>%
  group_by(year, size_class, division) %>%
  summarise(fishing_days = sum(fishing_days), .groups = "drop") %>%
  ggplot(aes(x = year, y = fishing_days, fill = size_class)) +
  theme_bw() + geom_col() +
  scale_fill_manual(values = size_colours) +
  scale_x_continuous(breaks = study_years) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(x = "", y = "Fishing days", fill = "Size class",
       title = "Fishing days by year, size class and ICES division") +
  facet_wrap(~ division)
```

Dividing the size-class breakdown by ICES division shows whether the fleet composition differs systematically between areas — for example, whether the eastern Channel hosts proportionally more large vessels than the western Channel or southern North Sea.

---

## 5. Spatial distribution of fishing events

### 5.1 Fishing event locations — all years combined

```{r plot9, fig.cap="Figure 7. Random sample of 20,000 fishing events (marine only, internal waters excluded) coloured by vessel flag state. Each point represents a single fishing event at its recorded position."}
sample_size <- 20000
events_sample <- events_classified_fr %>%
  dplyr::select(-any_of(c("vessel_flag", "vessel_name"))) %>%
  left_join(vessel_meta %>% dplyr::select(vesselId, vessel_flag),
            by = "vesselId") %>%
  mutate(vessel_flag = case_when(
    is.na(vessel_flag)                   ~ "Other",
    vessel_flag %in% names(flag_colours) ~ vessel_flag,
    TRUE                                 ~ "Other"
  )) %>%
  slice_sample(n = sample_size)

ggplot() +
  theme_bw() +
  geom_sf(data = fao_sf_division, fill = NA, colour = "grey60",
          linewidth = 0.3) +
  geom_sf(data = world, fill = "grey90", colour = "grey70", linewidth = 0.2) +
  geom_point(data = events_sample,
             aes(x = lon, y = lat, colour = vessel_flag),
             size = 0.5, alpha = 0.9) +
  scale_colour_manual(values = flag_colours, na.value = "grey70") +
  coord_sf(xlim = xlim, ylim = ylim, expand = FALSE) +
  guides(colour = guide_legend(override.aes = list(size = 3, alpha = 1))) +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_line(colour = "grey95"),
        legend.position = "bottom") +
  labs(colour = "Flag",
       title  = paste0("Fishing events — marine only (sample of ", sample_size, ")"),
       subtitle = "Internal waters excluded  |  ICES 27.4.C / 27.7.D / 27.7.E")
```

To illustrate the spatial distribution of fishing activity, a random sample of up to `r sample_size` fishing events is plotted across the three study divisions. Each point represents one fishing event at its recorded position. Internal waters are excluded. The plot gives an overview of which flag states fish in which parts of the study area.

---

### 5.2 Fishing event locations by year

```{r plot10, fig.height=10, fig.cap="Figure 8. Random sample of 10,000 fishing events per year, coloured by flag state, faceted by year. This reveals temporal changes in the spatial distribution of fishing effort."}
sample_size <- 10000
events_sample_by_year <- events_classified_fr %>%
  mutate(year = lubridate::year(start)) %>%
  filter(year %in% study_years) %>%
  dplyr::select(-any_of(c("vessel_flag", "vessel_name"))) %>%
  left_join(vessel_meta %>% dplyr::select(vesselId, vessel_flag),
            by = "vesselId") %>%
  mutate(vessel_flag = case_when(
    is.na(vessel_flag)                   ~ "Other",
    vessel_flag %in% names(flag_colours) ~ vessel_flag,
    TRUE                                 ~ "Other"
  )) %>%
  group_by(year) %>%
  slice_sample(n = sample_size)

ggplot() +
  theme_bw() +
  geom_sf(data = fao_sf_division, fill = NA,
          colour = "grey60", linewidth = 0.3) +
  geom_sf(data = world, fill = "grey90",
          colour = "grey70", linewidth = 0.2) +
  geom_point(data = events_sample_by_year,
             aes(x = lon, y = lat, colour = vessel_flag),
             size = 0.5, alpha = 0.9) +
  scale_colour_manual(values = flag_colours, na.value = "grey70") +
  coord_sf(xlim = xlim, ylim = ylim, expand = FALSE) +
  guides(colour = guide_legend(override.aes = list(size = 3, alpha = 1))) +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_line(colour = "grey95"),
        legend.position = "bottom") +
  labs(colour = "Flag",
       title  = paste0("Fishing events — marine only (sample of ", sample_size, ")"),
       subtitle = "Internal waters excluded  |  ICES 27.4.C / 27.7.D / 27.7.E") +
  facet_wrap(~ year)
```

Faceting the event map by year allows shifts in spatial effort patterns to be tracked over the study period. The sample is drawn independently per year to ensure each panel is equally populated regardless of variation in total event counts.

---

## 6. Effort density — kernel density estimation

```{r step11-prep}
# ---- Prepare spatial layers ----
crs_proj <- 32630  # UTM zone 30N

land_study <- world %>%
  st_crop(st_bbox(c(xmin = xlim[1], xmax = xlim[2],
                    ymin = ylim[1], ymax = ylim[2]),
                  crs = 4326)) %>%
  st_union()

study_window_sf <- fao_sf_division %>%
  st_union() %>%
  st_difference(land_study) %>%
  st_make_valid()

study_window_proj <- study_window_sf %>% st_transform(crs_proj)
study_owin        <- as.owin(study_window_proj)

world_proj   <- world        %>% st_transform(crs_proj)
fao_div_proj <- fao_sf_division %>% st_transform(crs_proj)
```

```{r step11-kde}
# ---- KDE — all divisions combined ----
sample_size <- 10000
sigma_m     <- 20000  # 20 km bandwidth

events_for_kde <- events_classified_fr %>%
  mutate(year = lubridate::year(start)) %>%
  filter(year %in% study_years, !is.na(lon), !is.na(lat)) %>%
  dplyr::select(year, lon, lat) %>%
  group_by(year) %>%
  slice_sample(n = sample_size) %>%
  ungroup()

events_for_kde_proj <- events_for_kde %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  st_transform(crs_proj) %>%
  mutate(x = st_coordinates(.)[, 1], y = st_coordinates(.)[, 2]) %>%
  st_drop_geometry()

kde_results <- map_dfr(study_years, function(yr) {
  pts <- events_for_kde_proj %>% filter(year == yr)
  if (nrow(pts) < 50) return(NULL)
  ppp_yr <- tryCatch(
    ppp(x = pts$x, y = pts$y, window = study_owin),
    error = function(e) NULL)
  if (is.null(ppp_yr)) return(NULL)
  kde <- density(ppp_yr, sigma = sigma_m, dimyx = c(200, 200), edge = TRUE)
  as.data.frame(kde) %>%
    rename(x = x, y = y, density = value) %>%
    filter(!is.na(density), density > 0) %>%
    mutate(ndensity = density / max(density), year = yr)
})
```

### 6.1 Fishing effort density — all divisions, all fleets

```{r plot11, fig.height=10, fig.cap="Figure 9. Kernel density estimates of fishing effort for all fleets combined, faceted by year. Events are projected to UTM zone 30N before KDE to ensure distance-based kernel widths are accurate. Land is excluded from the density window. Density is normalised per year (0 = minimum, 1 = maximum within each year)."}
ggplot() +
  theme_bw() +
  geom_tile(data = kde_results,
            aes(x = x, y = y, fill = ndensity)) +
  scale_fill_viridis_c(
    option = "inferno", name = "Relative\ndensity",
    limits = c(0, 1), breaks = c(0, 0.5, 1),
    labels = c("Low", "Mid", "High"), na.value = "transparent") +
  geom_sf(data = world_proj, fill = "grey85", colour = "grey60", linewidth = 0.2) +
  geom_sf(data = fao_div_proj, fill = NA, colour = "grey30", linewidth = 0.4) +
  coord_sf(crs  = crs_proj,
           xlim = st_bbox(study_window_proj)[c("xmin", "xmax")],
           ylim = st_bbox(study_window_proj)[c("ymin", "ymax")],
           expand = TRUE) +
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "grey90"),
    legend.position = "bottom", legend.title = element_text(size = 8),
    strip.background = element_rect(fill = "grey90"),
    strip.text = element_text(face = "bold", colour = "grey20"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "grey50", size = 8)) +
  labs(title    = "Fishing effort density — all fleets combined",
       subtitle = paste0("Masked KDE (sigma = ", sigma_m/1000, " km)  |  ",
                         "Land excluded  |  Normalised per year  |  ",
                         "Sample ", sample_size, " events/year")) +
  facet_wrap(~ year, ncol = 5)
```

Kernel density estimation (KDE) with a Gaussian kernel (bandwidth σ = `r sigma_m/1000` km) is applied to a random sample of fishing events per year. Events are projected to UTM zone 30N (EPSG:32630) so that the bandwidth is specified in metric units. The study window is the union of the three ICES divisions with land removed; densities outside this window are set to zero (edge correction applied). Relative density is normalised within each year to a 0–1 scale, making patterns comparable across years regardless of changes in absolute effort levels.

---

```{r step11-7d-prep}
# ---- 7d-specific spatial objects ----
bbox_7d <- st_bbox(fao_sf_division %>% filter(division == "27.7.D"))

land_7d <- world %>%
  st_crop(bbox_7d) %>%
  st_union()

window_7d_sf <- fao_sf_division %>%
  filter(division == "27.7.D") %>%
  st_difference(land_7d) %>%
  st_make_valid()

window_7d_proj <- window_7d_sf %>% st_transform(crs_proj)
window_7d_owin <- as.owin(window_7d_proj)
bbox_7d_proj   <- st_bbox(window_7d_proj)

events_7d_pts <- events_classified_fr %>%
  mutate(year = lubridate::year(start)) %>%
  filter(year %in% study_years, division == "27.7.D",
         !is.na(lon), !is.na(lat)) %>%
  dplyr::select(year, lon, lat)

events_7d_proj <- events_7d_pts %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  st_transform(crs_proj) %>%
  mutate(x = st_coordinates(.)[,1], y = st_coordinates(.)[,2]) %>%
  st_drop_geometry()
```

```{r step11-7d-kde}
# ---- KDE for ICES 27.7.D ----
kde_7d <- map_dfr(study_years, function(yr) {
  pts <- events_7d_proj %>% filter(year == yr)
  if (nrow(pts) < 50) return(NULL)
  ppp_yr <- tryCatch(
    ppp(x = pts$x, y = pts$y, window = window_7d_owin),
    error = function(e) NULL)
  if (is.null(ppp_yr)) return(NULL)
  kde <- density(ppp_yr, sigma = 10000, dimyx = c(200, 200), edge = TRUE)
  as.data.frame(kde) %>%
    rename(x = x, y = y, density = value) %>%
    filter(!is.na(density), density > 0) %>%
    mutate(ndensity = density / max(density), year = yr)
})
```

### 6.2 Fishing effort density — ICES 27.7.D

```{r plot12c-density, fig.height=10, fig.cap="Figure 10. Kernel density estimates of fishing effort in ICES 27.7.D, faceted by year. Bandwidth σ = 10 km; density normalised per year."}
ggplot() +
  theme_bw() +
  geom_tile(data = kde_7d, aes(x = x, y = y, fill = ndensity)) +
  scale_fill_viridis_c(
    option = "inferno", name = "Relative\ndensity",
    limits = c(0, 1), na.value = "transparent") +
  geom_sf(data = world_proj, fill = "grey40", colour = "grey60", linewidth = 0.2) +
  geom_sf(data = fao_div_proj %>% filter(division == "27.7.D"),
          fill = NA, colour = "white", linewidth = 0.5) +
  coord_sf(crs   = crs_proj,
           xlim  = bbox_7d_proj[c("xmin", "xmax")],
           ylim  = bbox_7d_proj[c("ymin", "ymax")],
           expand = TRUE) +
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "grey90"),
    legend.position = "bottom",
    strip.background = element_rect(fill = "grey90"),
    strip.text = element_text(face = "bold", colour = "grey20"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "grey50", size = 8)) +
  labs(title    = "Fishing effort density — ICES 27.7.D, all fleets combined",
       subtitle = "Masked KDE (sigma = 10 km)  |  Land excluded  |  Normalised per year") +
  facet_wrap(~ year)
```

The same KDE approach is applied at a finer scale to ICES 27.7.D alone, using a narrower bandwidth (σ = 10 km) to resolve within-division spatial structure. The narrower window allows persistent hotspots and seasonal shifts in effort distribution to be identified more precisely than in the whole-study-area maps.

---

## 7. Effort concentration — ICES 27.7.D

```{r step12-grid}
# ---- Assign events to 5 km grid cells ----
cell_size <- 5000  # metres

events_7d_grid <- events_7d_proj %>%
  mutate(
    x_cell  = floor(x / cell_size) * cell_size + cell_size / 2,
    y_cell  = floor(y / cell_size) * cell_size + cell_size / 2,
    cell_id = paste(x_cell, y_cell, sep = "_"))

cell_effort_7d <- events_7d_grid %>%
  group_by(year, cell_id, x_cell, y_cell) %>%
  summarise(n_events = n(), .groups = "drop")

concentration_7d <- cell_effort_7d %>%
  group_by(year) %>%
  summarise(
    n_cells      = n(),
    total_events = sum(n_events),
    gini         = ineq::Gini(n_events),
    top10_pct    = {
      s     <- sort(n_events, decreasing = TRUE)
      n_top <- max(1L, ceiling(0.10 * length(s)))
      sum(s[seq_len(n_top)]) / sum(s) * 100
    },
    n_cells_1pct = sum(n_events / total_events >= 0.01),
    .groups = "drop")
```

### 7.1 Gini coefficient of spatial effort concentration

```{r plot12a, fig.height=4, fig.cap="Figure 11. Gini coefficient of spatial fishing effort concentration in ICES 27.7.D over time. Values close to 1 indicate that effort is concentrated in a small number of grid cells; values close to 0 indicate a uniform spatial distribution."}
ggplot(concentration_7d, aes(x = year, y = gini)) +
  theme_bw() +
  geom_line(colour = "#E15759", linewidth = 1) +
  geom_point(colour = "#E15759", size = 3) +
  geom_text(aes(label = round(gini, 3)), vjust = -0.8, size = 2.8) +
  scale_x_continuous(breaks = study_years) +
  scale_y_continuous(limits = c(0, 1),
                     labels = scales::number_format(accuracy = 0.01)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5),
        plot.title  = element_text(face = "bold")) +
  labs(x = "", y = "Gini coefficient",
       title    = "Spatial concentration of fishing effort — ICES 27.7.D",
       subtitle = paste0("Gini coefficient across ", cell_size/1000,
                         " km grid cells  |  Higher = more concentrated",
                         "  |  All fleets combined"))
```

The Gini coefficient is computed on the distribution of fishing events across `r cell_size/1000` km × `r cell_size/1000` km grid cells. It ranges from 0 (perfectly uniform distribution) to 1 (all effort in a single cell). A trend towards higher Gini values indicates increasing spatial concentration of fishing effort over time.

---

### 7.2 Effort in top-10% most-fished cells

```{r plot12b, fig.height=4, fig.cap="Figure 12. Percentage of total annual fishing effort in ICES 27.7.D contained within the top-10% most-fished grid cells. A higher percentage indicates stronger spatial concentration."}
ggplot(concentration_7d, aes(x = year, y = top10_pct)) +
  theme_bw() +
  geom_line(colour = "#4E79A7", linewidth = 1) +
  geom_point(colour = "#4E79A7", size = 3) +
  geom_text(aes(label = paste0(round(top10_pct, 1), "%")), vjust = -0.8, size = 2.8) +
  scale_x_continuous(breaks = study_years) +
  scale_y_continuous(labels = scales::percent_format(scale = 1),
                     expand = expansion(mult = c(0.05, 0.15))) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5),
        plot.title  = element_text(face = "bold")) +
  labs(x = "", y = "% of total effort",
       title    = paste0("Effort in top-10% most-fished ",
                         cell_size/1000, " km cells — ICES 27.7.D"),
       subtitle = "All fleets combined")
```

This metric captures the share of total annual fishing effort accounted for by the 10% of grid cells with the highest activity. Together with the Gini coefficient, it provides a complementary perspective on spatial concentration: the Gini summarises the full distribution, while the top-10% share focuses on the concentration in the most intensively fished areas.

---

### 7.3 Concentration metrics combined

```{r plot12c-conc, fig.height=5, fig.cap="Figure 13. Gini coefficient and top-10% cell share plotted together on separate panels with free y-axis scales. The two metrics tend to move in parallel but may diverge when effort becomes more concentrated in fewer cells without a proportional increase in the Gini value."}
conc_long <- concentration_7d %>%
  dplyr::select(year, gini, top10_pct) %>%
  pivot_longer(c(gini, top10_pct),
               names_to = "metric", values_to = "value") %>%
  mutate(metric = recode(metric,
                         gini      = "Gini coefficient",
                         top10_pct = "Top-10% cell share (%)"))

ggplot(conc_long, aes(x = year, y = value, colour = metric, group = metric)) +
  theme_bw() +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  scale_colour_manual(values = c("Gini coefficient"       = "#E15759",
                                 "Top-10% cell share (%)" = "#4E79A7")) +
  scale_x_continuous(breaks = study_years) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  theme(axis.text.x     = element_text(angle = 90, vjust = 0.5),
        legend.position = "bottom",
        plot.title      = element_text(face = "bold")) +
  labs(x = "", y = "Value", colour = "",
       title    = paste0("Effort concentration — ICES 27.7.D  |  ",
                         cell_size/1000, " km grid  |  All fleets"),
       subtitle = "Gini coefficient (0–1) and % effort in top-10% most-fished cells") +
  facet_wrap(~ metric, scales = "free_y")
```

---

### 7.4 Mean effort hotspots — ICES 27.7.D (all years)

```{r plot12d, fig.height=5, fig.cap="Figure 14. Mean annual fishing effort per 5 km grid cell in ICES 27.7.D, averaged across all years. Cells are coloured by relative effort (0–1). This identifies persistent spatial hotspots of fishing activity."}
cell_effort_mean <- cell_effort_7d %>%
  group_by(cell_id, x_cell, y_cell) %>%
  summarise(mean_events = mean(n_events), .groups = "drop") %>%
  mutate(ndensity = mean_events / max(mean_events))

ggplot() +
  theme_bw() +
  geom_tile(data = cell_effort_mean,
            aes(x = x_cell, y = y_cell, fill = ndensity)) +
  scale_fill_viridis_c(
    option = "inferno", name = "Relative\neffort",
    limits = c(0, 1), na.value = "transparent") +
  geom_sf(data = world_proj, fill = "grey85", colour = "grey60", linewidth = 0.2) +
  geom_sf(data = fao_div_proj %>% filter(division == "27.7.D"),
          fill = NA, colour = "grey30", linewidth = 0.5) +
  coord_sf(crs   = crs_proj,
           xlim  = bbox_7d_proj[c("xmin", "xmax")],
           ylim  = bbox_7d_proj[c("ymin", "ymax")],
           expand = TRUE) +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "white"),
        panel.grid = element_line(colour = "grey90"),
        legend.position = "bottom",
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(colour = "grey50", size = 8)) +
  labs(title    = paste0("Mean fishing effort hotspots — ICES 27.7.D  |  ",
                         cell_size/1000, " km grid"),
       subtitle = "Mean events per cell averaged across all years  |  All fleets combined")
```

Grid cells of `r cell_size/1000` km × `r cell_size/1000` km are used to summarise fishing intensity. The mean number of events per cell is averaged across all study years and normalised to a 0–1 relative effort scale. This produces a time-averaged spatial fingerprint of fishing activity, showing which locations are persistently most heavily exploited.

---

### 7.5 Annual effort hotspots by year — ICES 27.7.D

```{r step12-full-grid}
# ---- Build complete grid ----
x_range <- seq(
  floor(min(cell_effort_7d$x_cell) / cell_size) * cell_size,
  ceiling(max(cell_effort_7d$x_cell) / cell_size) * cell_size,
  by = cell_size)
y_range <- seq(
  floor(min(cell_effort_7d$y_cell) / cell_size) * cell_size,
  ceiling(max(cell_effort_7d$y_cell) / cell_size) * cell_size,
  by = cell_size)

full_grid_7d <- expand_grid(year = study_years, x_cell = x_range, y_cell = y_range)

full_grid_sf <- full_grid_7d %>%
  st_as_sf(coords = c("x_cell", "y_cell"), crs = crs_proj, remove = FALSE) %>%
  st_filter(window_7d_proj %>% st_buffer(cell_size / 2)) %>%
  st_drop_geometry()

cell_effort_full <- full_grid_sf %>%
  left_join(
    cell_effort_7d %>% dplyr::select(year, x_cell, y_cell, n_events),
    by = c("year", "x_cell", "y_cell")) %>%
  replace_na(list(n_events = 0)) %>%
  group_by(year) %>%
  mutate(ndensity = n_events / max(n_events)) %>%
  ungroup()

scale_cap <- quantile(cell_effort_full$n_events[cell_effort_full$n_events > 0],
                      0.99, na.rm = TRUE)
cell_effort_plot <- cell_effort_full %>%
  mutate(n_events_capped = pmin(n_events, scale_cap))
quants <- quantile(cell_effort_full$n_events[cell_effort_full$n_events > 0],
                   c(0.50, 0.75, 0.90, 0.95, 0.99), na.rm = TRUE)
```

```{r plot12e, fig.height=12, fig.cap="Figure 15. Annual fishing effort per 5 km grid cell in ICES 27.7.D. The colour scale is capped at the 99th percentile of non-zero cells across all years combined to prevent extreme single-cell values from compressing the scale. Cells with zero events are shown in light grey."}
# safe_gradientn: guarantees >=2 unique breakpoints and range > 0,
# preventing the approxfun error when quantiles collapse on sparse data.
safe_gradientn <- function(raw_breaks, palette_colours) {
  upper <- max(c(raw_breaks, 1), na.rm = TRUE)
  pts   <- sort(unique(c(0, raw_breaks[raw_breaks > 0 & raw_breaks <= upper], upper)))
  if (length(pts) < 2) pts <- c(0, upper)
  list(
    colours = colorRampPalette(palette_colours)(length(pts)),
    values  = scales::rescale(pts),
    limits  = c(0, upper)
  )
}

palette_7 <- c("white", "#c7e9b4", "#7fcdbb", "#41b6c4",
               "#1d91c0", "#225ea8", "#0c2c84")
grad <- safe_gradientn(
  c(quants["50%"], quants["75%"], quants["90%"],
    quants["95%"], quants["99%"], scale_cap),
  palette_7)

ggplot() +
  theme_bw() +
  geom_tile(data = cell_effort_plot,
            aes(x = x_cell, y = y_cell, fill = n_events_capped)) +
  scale_fill_gradientn(
    colours = grad$colours,
    values  = grad$values,
    name    = "Events\nper cell",
    limits  = grad$limits, na.value = "grey92") +
  geom_sf(data = world_proj, fill = "grey80", colour = "grey60", linewidth = 0.2) +
  geom_sf(data = fao_div_proj %>% filter(division == "27.7.D"),
          fill = NA, colour = "grey30", linewidth = 0.5) +
  coord_sf(crs  = crs_proj,
           xlim = bbox_7d_proj[c("xmin", "xmax")],
           ylim = bbox_7d_proj[c("ymin", "ymax")],
           expand = TRUE) +
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.background = element_rect(fill = "grey92"),
    panel.grid = element_blank(),
    legend.position = "bottom", legend.key.width = unit(2, "cm"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "grey50", size = 8),
    strip.background = element_rect(fill = "grey90"),
    strip.text = element_text(face = "bold", colour = "grey20")) +
  labs(title    = paste0("Fishing effort hotspots — ICES 27.7.D  |  ",
                         cell_size/1000, " km grid"),
       subtitle = paste0("Events per cell per year  |  All fleets combined  |  ",
                         "Scale capped at p99 (", round(scale_cap), " events)")) +
  facet_wrap(~ year, ncol = 5)
```

Annual hotspot maps use a consistent colour scale across all years, capped at the 99th percentile of non-zero cell counts from the combined dataset. This prevents extreme values in individual years from distorting the colour scale, while retaining the ability to compare effort intensities across years. Cells with no recorded events are shown in light grey.

---

### 7.6 Effort hotspots by flag state — FRA and NLD

```{r step12-flag-prep}
# ---- Build helper function for flag-specific cell effort ----
make_cell_effort <- function(flag_code) {
  events_flag <- events_classified_fr %>%
    mutate(year = lubridate::year(start)) %>%
    filter(year %in% study_years, division == "27.7.D",
           !is.na(lon), !is.na(lat)) %>%
    dplyr::select(year, vesselId, lon, lat) %>%
    left_join(vessel_meta %>% dplyr::select(vesselId, vessel_flag),
              by = "vesselId") %>%
    filter(vessel_flag == flag_code)

  if (nrow(events_flag) == 0) return(NULL)

  events_proj <- events_flag %>%
    st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
    st_transform(crs_proj) %>%
    mutate(x = st_coordinates(.)[,1], y = st_coordinates(.)[,2]) %>%
    st_drop_geometry()

  events_grid <- events_proj %>%
    mutate(
      x_cell  = floor(x / cell_size) * cell_size + cell_size / 2,
      y_cell  = floor(y / cell_size) * cell_size + cell_size / 2,
      cell_id = paste(x_cell, y_cell, sep = "_"))

  cell_effort <- events_grid %>%
    group_by(year, cell_id, x_cell, y_cell) %>%
    summarise(n_events = n(), .groups = "drop")

  full_grid_sf %>%
    mutate(x_cell = round(x_cell), y_cell = round(y_cell)) %>%
    left_join(
      cell_effort %>%
        mutate(x_cell = round(x_cell), y_cell = round(y_cell)) %>%
        dplyr::select(year, x_cell, y_cell, n_events),
      by = c("year", "x_cell", "y_cell")) %>%
    replace_na(list(n_events = 0)) %>%
    mutate(flag = flag_code)
}

make_hotspot_plot <- function(cell_effort_flag, flag_label,
                              scale_cap = NULL, quants = NULL) {
  if (is.null(scale_cap)) {
    scale_cap <- quantile(
      cell_effort_flag$n_events[cell_effort_flag$n_events > 0],
      0.99, na.rm = TRUE)
  }
  if (is.null(quants)) {
    quants <- quantile(
      cell_effort_flag$n_events[cell_effort_flag$n_events > 0],
      c(0.50, 0.75, 0.90, 0.95, 0.99), na.rm = TRUE)
  }

  palette_7 <- c("white", "#c7e9b4", "#7fcdbb", "#41b6c4",
                 "#1d91c0", "#225ea8", "#0c2c84")
  grad <- safe_gradientn(
    c(quants["50%"], quants["75%"], quants["90%"],
      quants["95%"], quants["99%"], scale_cap),
    palette_7)

  cell_effort_flag %>%
    mutate(n_events_capped = pmin(n_events, scale_cap)) %>%
    ggplot() +
    theme_bw() +
    geom_tile(aes(x = x_cell, y = y_cell, fill = n_events_capped)) +
    scale_fill_gradientn(
      colours = grad$colours,
      values  = grad$values,
      name    = "Events\nper cell",
      limits  = grad$limits, na.value = "grey92") +
    geom_sf(data = world_proj, fill = "grey80", colour = "grey60", linewidth = 0.2) +
    geom_sf(data = fao_div_proj %>% filter(division == "27.7.D"),
            fill = NA, colour = "grey30", linewidth = 0.5) +
    coord_sf(crs  = crs_proj,
             xlim = bbox_7d_proj[c("xmin", "xmax")],
             ylim = bbox_7d_proj[c("ymin", "ymax")],
             expand = TRUE) +
    theme(
      axis.text = element_blank(), axis.ticks = element_blank(),
      axis.title = element_blank(),
      panel.background = element_rect(fill = "grey92"),
      panel.grid = element_blank(),
      legend.position = "bottom", legend.key.width = unit(2, "cm"),
      plot.title = element_text(face = "bold"),
      plot.subtitle = element_text(colour = "grey50", size = 8),
      strip.background = element_rect(fill = "grey90"),
      strip.text = element_text(face = "bold", colour = "grey20")) +
    labs(title    = paste0("Fishing effort hotspots — ", flag_label,
                           "  |  ICES 27.7.D  |  ", cell_size/1000, " km grid"),
         subtitle = paste0("Events per cell per year  |  ", flag_label,
                           " fleet only  |  Scale capped at p99 (",
                           round(scale_cap), " events)")) +
    facet_wrap(~ year)
}

cell_effort_fra <- make_cell_effort("FRA")
cell_effort_nld <- make_cell_effort("NLD")

combined_events  <- c(
  cell_effort_fra$n_events[cell_effort_fra$n_events > 0],
  cell_effort_nld$n_events[cell_effort_nld$n_events > 0])
shared_scale_cap <- quantile(combined_events, 0.99, na.rm = TRUE)
shared_quants    <- quantile(combined_events, c(0.50, 0.75, 0.90, 0.95, 0.99),
                             na.rm = TRUE)
```

```{r plot12f, fig.height=12, fig.cap="Figure 16. Annual fishing effort hotspots in ICES 27.7.D — French (FRA) fleet only. Colour scale shared with the Dutch fleet plot to allow direct comparison."}
make_hotspot_plot(cell_effort_fra, "FRA",
                  scale_cap = shared_scale_cap,
                  quants    = shared_quants)
```

```{r plot12g, fig.height=12, fig.cap="Figure 17. Annual fishing effort hotspots in ICES 27.7.D — Dutch (NLD) fleet only. Colour scale shared with the French fleet plot to allow direct comparison."}
make_hotspot_plot(cell_effort_nld, "NLD",
                  scale_cap = shared_scale_cap,
                  quants    = shared_quants)
```

The flag-specific hotspot maps use a shared colour scale derived from the 99th percentile of non-zero cell counts across both fleets combined. This ensures that the spatial patterns of the French and Dutch fleets can be compared directly: identical colours represent identical effort intensities. Cells with no events for that fleet are shown in light grey even where the other fleet is active.

---

# Session information

```{r session-info}
sessionInfo()
```
