---
output:
  word_document:
    reference_docx: ../MPFF report template v1.1.dotx
---


```{r setup, include=FALSE}

# ==========================================================================================================
# Generate FLYSHOOT trips overview
# 
# 8/3/2023 first coding
# 22/4/2026 updated to parquet/Poseidat pipeline (consistent with 01_flyshoot_functions.R)
# ==========================================================================================================

knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 

options(dplyr.summarise.inform = FALSE)

# Reset lists
rm(list=ls())

# Libraries
library(rmarkdown)                   # note: requires knitr 1.21

require(tidyverse, quietly=TRUE)     # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
require(lubridate, quietly=TRUE)     # data handling
require(reshape2, quietly=TRUE)      # reshaping data; e.g. cast
require(RColorBrewer)                # colour schemes
library(viridis)
library(pander)
library(arrow)                       # parquet file support

library(captioner)                   # for captions
tab_nums <- captioner::captioner(prefix = "Table " , levels=1, type=c("n"), infix=".")
fig_nums <- captioner::captioner(prefix = "Figure ", levels=1, type=c("n"), infix=".")

# Source all the utils
source("../r/FLYSHOOT utils.R")

spatialdir = "C:/Users/MartinPastoors/OneDrive - Martin Pastoors/DATA/RDATA"

# spatial data files
load(file.path(spatialdir, "world_mr_sf.RData"))
load(file.path(spatialdir, "world_mr_df.RData"))
load(file.path(spatialdir, "rect_lr_sf.RData"))
load(file.path(spatialdir, "world_lr_df.RData"))
load(file.path(spatialdir, "asfis.RData"))

# set onedrive directory
onedrive <- get_onedrive(team="Martin Pastoors", site="FLYSHOOT - General")

# parquet data directory
parquetdir <- file.path(onedrive, "parquet")

# datasets with price and regulated information derived from read_afslag.r
price  <- 
  readr::read_rds(file.path(onedrive, "rdata/prices.rds")) %>% 
  dplyr::select(year, species, avgprice) %>% 
  mutate(price_cat = cut(avgprice, breaks=c(0,0.5, 1,2,3,4,5,10,15,20), dig.lab=10 ))

price_raw <-
  readr::read_rds(file.path(onedrive, "rdata/prices_raw.rds")) 
  
regulated  <- 
  readr::read_rds(file.path(onedrive, "rdata/prices.rds")) %>% 
  ungroup() %>% 
  dplyr::distinct(species, regulated) 

# ===============================================================================
#  Load data from parquet files (Poseidat naming convention)
# ===============================================================================

haul   <- arrow::read_parquet(file.path(parquetdir, "haul.parquet"))
kisten <- arrow::read_parquet(file.path(parquetdir, "kisten.parquet"))
elog   <- arrow::read_parquet(file.path(parquetdir, "elog.parquet")) %>%
  mutate(species = toupper(species)) %>% 
  mutate(species = ifelse(species == "JAX", "HOM", species)) %>% 
  mutate(species = ifelse(species == "SQC", "SQR", species)) %>%
  left_join(dplyr::select(asfis,
                          species, scientificname, englishname, dutchname),
            by = "species") %>%
  mutate(species = toupper(species)) %>% 
  mutate(englishname = ifelse(englishname == "European pilchard(=Sardine)", 
                              "European pilchard", englishname)) %>% 
  left_join(regulated, by = "species") %>% 
  mutate(regulated = case_when(
    is.na(regulated) & species %in% c("BSE","DGX","ANF","DGH","HER","PLE")             ~ "R",
    is.na(regulated) & species %in% c("BRB","SBX","SQS","WEX","WEG","SQU","SDV",
                                      "OCT","MUL","SCR","GAG","ANE","SQE","SQC","OCZ") ~ "NR",
    TRUE ~ regulated
  )) %>% 
  ungroup()

# Column name mapping (Poseidat → overview script aliases):
#   haul:   vessel, trip_id, haul_id, date, lat (shoot_lat), lon (shoot_lon),
#           zone_fao_division (division), zone_ices_rectangle (rect),
#           water_depth, catchweight
#   kisten: vessel, trip_id, haul_id, weighing_time, species_code (species),
#           weight_kg (gewicht), size_category (maat)
#   elog:   vessel, trip_id, date, species, weight_kg (weight),
#           zone_fao_division (division), zone_ices_rectangle (rect),
#           presentation, preservation, weight_undersized

# Standardise haul column aliases for downstream use
haul <- haul %>%
  rename(
    trip     = trip_id,
    haul     = haul_id,
    division = zone_fao_division,
    rect     = zone_ices_rectangle,
    lat      = shoot_lat,
    lon      = shoot_lon
  ) %>%
  mutate(
    year    = lubridate::year(date),
    month   = lubridate::month(date),
    week    = lubridate::week(date),
    quarter = lubridate::quarter(date)
  )

# Standardise kisten column aliases
kisten <- kisten %>%
  rename(
    trip     = trip_id,
    haul     = haul_id,
    datetime = weighing_time,       # keep 'datetime' name used in downstream code
    species  = species_code,
    gewicht  = weight_kg
  ) %>%
  mutate(
    year    = lubridate::year(datetime),
    month   = lubridate::month(datetime),
    week    = lubridate::week(datetime)
  )

# Standardise elog column aliases
elog <- elog %>%
  rename(
    trip   = trip_id,
    weight = weight_kg
  ) %>%
  mutate(
    year    = lubridate::year(date),
    month   = lubridate::month(date),
    week    = lubridate::week(date),
    quarter = lubridate::quarter(date)
  )

# ===============================================================================
#  Filter parameters
# ===============================================================================

yrs     <- 2022:2026
longyrs <- 2012:2026
mnths   <- NA
vessels <- c("SL9","SCH99", "Z99", "SCH135","SCH65", "SCH144")

myyear <- max(yrs)

# making the data selections
h <- 
  haul %>%
  ungroup() %>% 
  {if(!is.na(all(vessels))) filter(., vessel %in% vessels) else . } %>% 
  {if(!is.na(all(yrs)))     filter(., year %in% yrs) else . } %>% 
  {if(!is.na(all(mnths)))   filter(., month %in% mnths) else .} %>% 
  ungroup()

k <- 
  kisten %>%
  ungroup() %>% 
  {if(!is.na(all(vessels))) filter(., vessel %in% vessels) else . } %>% 
  {if(!is.na(all(yrs)))     filter(., year %in% yrs) else . } %>% 
  {if(!is.na(all(mnths)))   filter(., month %in% mnths) else .} %>% 
  left_join(dplyr::select(h, 
                          vessel, trip, haul, division, lat, lon, rect),
            by = c("vessel", "trip", "haul"))

e <- 
  elog %>%
  ungroup() %>% 
  {if(!any(is.na(vessels))) filter(., vessel %in% vessels) else . } %>% 
  {if(!any(is.na(yrs)))     filter(., year %in% yrs) else . } %>% 
  {if(!any(is.na(mnths)))   filter(., month %in% mnths) else .} 

# ===============================================================================
#  Derived datasets
# ===============================================================================

# Determine number of years
nyears_actual <- length(yrs)

# Determine number of species
nspecies_actual <- 12

effort <-
  e %>% 
  dplyr::group_by(vessel, year) %>% 
  dplyr::summarize(ndays = n_distinct(paste(vessel, date))) 

catch <-
  e %>% 
  dplyr::group_by(vessel, year, species, 
                  scientificname, englishname, dutchname) %>% 
  dplyr::summarize(weight = sum(weight, na.rm=TRUE)) %>% 
  dplyr::group_by(vessel, year) %>% 
  dplyr::mutate(perc = weight/sum(weight, na.rm=TRUE)) %>% 
  left_join(effort, by=c("vessel", "year")) %>% 
  mutate(catch_day = weight / ndays)

revenue_byweek <-
  elog %>% 
  left_join(price, by=c("year","species")) %>% 
  mutate(revenue = weight * avgprice) %>% 
  drop_na(revenue) %>% 
  group_by(vessel, year, quarter, week, rect, division, species, 
           scientificname, englishname, dutchname, regulated) %>% 
  summarise(
    weight  = sum(weight, na.rm=TRUE),
    revenue = sum(revenue, na.rm=TRUE),
    ndays   = n_distinct(date)
  ) %>% 
  ungroup()

# get top species
top      <- 
  readr::read_rds(file=file.path(onedrive,"rdata", "top.rds")) %>%  
  lowcase() %>% 
  dplyr::select(-value, -weight) 

# ===============================================================================
#  Map extent calculations
# ===============================================================================

lon_range <- range(e$lon, na.rm = TRUE)
lat_range <- range(e$lat, na.rm = TRUE)

lon_buffer <- diff(lon_range) * 0.1
lat_buffer <- diff(lat_range) * 0.1

xmin <- floor(lon_range[1] - lon_buffer)
xmax <- ceiling(lon_range[2] + lon_buffer)
ymin <- floor((lat_range[1] - lat_buffer) * 2) / 2
ymax <- ceiling((lat_range[2] + lat_buffer) * 2) / 2

xdistance <- icosa::arcdist(p1 = c(xmin, ymin), p2 = c(xmax, ymin))
ydistance <- icosa::arcdist(p1 = c(xmin, ymin), p2 = c(xmin, ymax))
data_aspect <- ydistance / xdistance

# Choose ncol based on data aspect ratio
if (data_aspect > 1.2) {
  ncol_use_years <- min(4, ceiling(nyears_actual / 2))
} else if (data_aspect < 0.7) {
  ncol_use_years <- min(2, ceiling(nyears_actual / 3))
} else if (nyears_actual < 6) {
  ncol_use_years <- min(2, ceiling(nyears_actual / 3))
} else {
  ncol_use_years <- 3
}

fig_width <- knitr::opts_chunk$get("fig.width")  
nrows_years <- ceiling(nyears_actual / ncol_use_years)

fig_height_years <- (fig_width / ncol_use_years) * data_aspect * nrows_years + 1.5
fig_height_years <- min(fig_height_years, fig_width * 1.2)

if (data_aspect > 1.2) {
  ncol_use_species <- min(4, ceiling(nspecies_actual / 2))
} else if (data_aspect < 0.7) {
  ncol_use_species <- min(2, ceiling(nspecies_actual / 3))
} else {
  ncol_use_species <- 3
}

nrows_species <- ceiling(nspecies_actual / ncol_use_species)

fig_height_species <- (fig_width / ncol_use_species) * data_aspect * nrows_species + 1.5
fig_height_species <- min(fig_height_species, fig_width * 1.2)

# Helper: dynamic year colour scales (works for any length of yrs)
n_yrs       <- length(yrs)
year_cols   <- c(rep("gray80", max(n_yrs - 1, 1)), "red")
year_widths <- c(rep(0.5, max(n_yrs - 1, 1)), 1.5)
year_alphas <- c(rep(1, max(n_yrs - 1, 1)), 0.5)
# trim to actual n_yrs in case n_yrs == 1
year_cols   <- tail(year_cols,   n_yrs)
year_widths <- tail(year_widths, n_yrs)
year_alphas <- tail(year_alphas, n_yrs)

```


**FLYSHOOT trips overview**

For: Jaczon BV

By: Martin Pastoors

Report generated: `r format(Sys.time(), '%d/%m/%Y')`


```{r triptable, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}

tab_nums(name    = "trips", level = 1, display = FALSE, caption = "Self-sampling overview")

t <-
  bind_cols(
    h %>% 
      filter(year %in% yrs) %>% 
      group_by(year, month) %>% 
      summarise(
        nvessels      = n_distinct(vessel),
        ntrips        = n_distinct(paste(vessel, trip)),
        ndays         = n_distinct(paste(vessel, trip, date)),
        nhauls        = n_distinct(paste(vessel, trip, haul)),
        nrects        = n_distinct(rect),
      ),
    k %>% 
      filter(year %in% yrs) %>% 
      group_by(vessel, trip, haul, year, month) %>% 
      summarise(gewicht = sum(gewicht, na.rm=TRUE)) %>% 
      group_by(year, month) %>% 
      summarise(
        landedcatch   = sum(gewicht)
      ) %>% 
      ungroup() %>% 
      dplyr::select(-year, -month)
    
  )  %>% 

  # Discards reported (from elog, presentation == "DIS")
  left_join(
    e %>% 
      filter(year %in% yrs) %>% 
      filter(presentation == "DIS") %>% 
      mutate(gewicht = as.numeric(ifelse(is.na(weight), 0, weight)) + 
               as.numeric(ifelse(is.na(weight_undersized), 0, weight_undersized))) %>% 
      group_by(vessel, trip, haul, year, month) %>% 
      summarise(
        gewicht = sum(gewicht, na.rm=TRUE)      
      ) %>% 
      group_by(year, month) %>% 
      summarise(
        discardedcatch   = sum(gewicht)
      ) %>% 
      ungroup(),
    by = c("year", "month")
  ) %>% 

  # discard ratio (for SCH99/Aravis only, from kisten vs haul catchweight)
  left_join(
    k %>% 
      filter(vessel %in% c("SCH99","Z99"), 
             as.Date(datetime) >= lubridate::dmy("1/11/2023")) %>% 
      group_by(vessel, trip, haul, year, month) %>% 
      summarise(gewicht = sum(gewicht, na.rm=TRUE)) %>% 
      
      left_join(h, by=c("vessel","trip","haul", "year", "month")) %>% 
      filter(!is.na(catchweight)) %>% 
    
      group_by(year, month) %>% 
      summarise(
        landedcatch   = sum(gewicht, na.rm=TRUE),
        totalcatch    = sum(catchweight, na.rm=TRUE)
      ) %>% 
      mutate(discardperc = 1-(landedcatch/totalcatch)) %>% 
      ungroup() %>% 
      dplyr::select(-landedcatch, -totalcatch),
    by = c("year","month")

  ) %>% 
  
  mutate(year=as.character(year), month=as.character(month)) 

t2 <-
  t %>% 
  ungroup() %>% 
  summarise(
    nvessels = mean(nvessels, na.rm=TRUE),
    ntrips = sum(ntrips),
    ndays = sum(ndays),
    nhauls = sum(nhauls),
    nrects = mean(nrects),
    landedcatch = sum(landedcatch),
    discardedcatch = sum(discardedcatch, na.rm=TRUE),
    discardperc = mean(discardperc, na.rm=TRUE)
  ) %>% 
  mutate(
    year = "all",
    month = "all")

bind_rows(t, t2) %>% 
  mutate(discardperc = scales::label_percent(accuracy = 1L)(discardperc)) %>% 
  mutate(discardperc = ifelse(is.na(discardperc), "-", discardperc)) %>% 
  pander::pandoc.table(., 
               style        = "simple",
               split.tables = 200, 
               split.cells  = c(rep(7,10)),
               justify      = "right",
               missing      =" ",
               big.mark     = ',', 
               round        = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0) )

  
cat("Discard percentage based on SCH99 only")

```

_`r tab_nums("trips")`_

\newpage

Totale aanvoer per schip en per week

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.asp=1.0}

fig_nums(
  name    = "d", 
  level = 1, display = FALSE,
  caption = "Totale aanvoer per schip en per week. Zwarte lijn is de gemiddelde aanvoer per schip en per week")

kk <- 
  e %>%
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  group_by(vessel, year, week) %>% 
  summarise(landingweight = sum(weight, na.rm=TRUE)) %>% 
  group_by(vessel, year) %>% 
  mutate(meanweight = mean(landingweight, na.rm=TRUE)) 

se <- function(x) sqrt(var(x)/length(x))

kkk <- 
  kk %>%
  filter(year %in% yrs[-length(yrs)]) %>% 
  group_by(vessel, week) %>% 
  summarise(
    meanweightperweek = mean(landingweight, na.rm=TRUE),
    sd                = sd(landingweight,na.rm=TRUE),
    sterr             = se(landingweight)
  )

kk %>% 
  ggplot(aes(group=vessel)) +
  theme_publication() +
  theme(legend.position = "none") +
  geom_bar(aes(x=week, y=landingweight, fill=vessel), stat="identity") +
  geom_step(data=kkk, aes(x=week, y=meanweightperweek), fill=NA, colour="black") +
  facet_grid(vessel~year)

```

\newpage

Aanvoer per soort en per week

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.asp=1.0}

ee <- 
  e %>%
  filter(species %in% top$species[1:6]) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  group_by(species, year, week) %>% 
  summarise(landingweight = sum(weight, na.rm=TRUE)) %>% 
  group_by(species, year) %>% 
  mutate(meanweight = mean(landingweight, na.rm=TRUE)) %>% 
  mutate(year = as.character(year)) %>% 
  mutate(species = factor(species, levels=top$species[1:6]))

eee <- 
  ee %>%
  filter(year %in% yrs[-length(yrs)]) %>% 
  group_by(species, week) %>% 
  summarise(
    meanweightperweek = mean(landingweight, na.rm=TRUE),
    sd                = sd(landingweight,na.rm=TRUE),
    sterr             = se(landingweight)
  )

ee %>% 
  ggplot() +
  theme_publication() +
  theme(legend.position = "none") +
  geom_bar(aes(x=week, y=landingweight, fill=species), stat="identity") +
  geom_step(data=eee, aes(x=week, y=meanweightperweek), fill=NA, colour="black") +
  facet_grid(species~year, scales="free_y")

```



Aanvoer per soort, per jaar en per week. Rode lijn is het lopende jaar. 

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.height=fig_height_species}

ee <- 
  e %>%
  filter(species %in% top$species) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  group_by(species, year, week) %>% 
  summarise(landingweight = sum(weight, na.rm=TRUE)) %>% 
  group_by(species, year) %>% 
  mutate(meanweight = mean(landingweight, na.rm=TRUE)) %>% 
  mutate(year = as.character(year)) %>% 
  mutate(species = factor(species, levels=top$species))

ee %>% 
  ggplot(aes(group=year)) +
  theme_publication() +
  geom_line(aes(x=week, y=landingweight, colour=year, linewidth=year, alpha=year)) +
  scale_linewidth_manual("year", values = year_widths, guide = "none") +
  scale_colour_manual("year", values = year_cols) +
  scale_alpha_manual("year", values = year_alphas, guide = "none") +
  facet_wrap(~species, scales="free_y", nrow = nrows_species)

```

\newpage

Cumulatieve aanvoer per jaar en per week. Rode lijn is het lopende jaar. 

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.asp=1.0}

ee <- 
  e %>%
  group_by(year, week) %>% 
  summarise(landingweight = sum(weight, na.rm=TRUE)) %>% 
  group_by(year) %>% 
  mutate(cumweight = cumsum(landingweight/1000)) %>% 
  mutate(year = as.character(year)) 

ee %>% 
  ggplot() +
  theme_publication() +
  geom_line(aes(x=week, y=cumweight, colour=year, linewidth=year, alpha=year)) +
  scale_linewidth_manual("year", values = year_widths, guide = "none") +
  scale_colour_manual("year", values = year_cols) +
  scale_alpha_manual("year", values = year_alphas, guide = "none") +
  labs(y="aanvoer (ton)")

```

\newpage

Cumulatieve aanvoer per soort, per jaar en per week. Rode lijn is het lopende jaar. 

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.height=fig_height_species}

ee <- 
  e %>%
  filter(species %in% top$species) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  group_by(species, year, week) %>% 
  summarise(landingweight = sum(weight, na.rm=TRUE)) %>% 
  group_by(species, year) %>% 
  mutate(cumweight = cumsum(landingweight)) %>% 
  mutate(year = as.character(year)) %>% 
  mutate(species = factor(species, levels=top$species))

ee %>% 
  ggplot() +
  theme_publication() +
  geom_line(aes(x=week, y=cumweight, colour=year, linewidth=year, alpha=year)) +
  scale_linewidth_manual("year", values = year_widths, guide = "none") +
  scale_colour_manual("year", values = year_cols) +
  scale_alpha_manual("year", values = year_alphas, guide = "none") +
  facet_wrap(~species, scales="free_y", nrow=nrows_species)

```




```{r aanvoerendiscards, echo=FALSE, fig.align="center", fig.asp=0.9, message=FALSE, warning=FALSE}

print_option <- FALSE

if (nrow(k) > 0 & sum(!is.na(h$catchweight))>0) {

  print_option <- TRUE

  fig_nums(
    name    = "percdiscards", level   = 1, display = FALSE,
    caption = "Berekende percentage discards per dag")

  t1 <-
    haul %>%
    filter(catchweight>0, vessel=="SCH99", date >= dmy("23/10/2023")) %>% 
    filter(date <= max(h$date, na.rm=TRUE)) %>% 
    group_by(vessel, trip, haul) %>%
    summarise(data = sum(catchweight, na.rm=TRUE)) %>%
    mutate(variable = "totaal")

  t2 <-
    kisten %>%
    filter(paste(vessel, trip, haul) %in% paste(t1$vessel, t1$trip, t1$haul)) %>%
    group_by(vessel, trip, haul) %>%
    summarise(data = sum(gewicht, na.rm=TRUE)) %>%
    mutate(variable = "aanvoer")

  tt <-
    bind_rows(t1, t2) %>% 
    arrange(vessel, trip, haul) %>% 
    tidyr::pivot_wider(names_from = variable, values_from = data) %>% 
    drop_na(totaal) %>% 
    mutate(aanvoer = ifelse(is.na(aanvoer),0,aanvoer)) %>% 
    mutate(
      discards = totaal - aanvoer, 
      perc_discards = round((discards) / totaal, digits=2)
    ) %>% 
    left_join(dplyr::select(haul,
                            vessel, trip, haul, year, week, date),
              by = c("vessel","trip","haul"))
  
  tt2 <-
    tt %>% 
    group_by(vessel, year, week, date) %>% 
    summarise(
      aanvoer = sum(aanvoer, na.rm=TRUE),
      discards = sum(discards, na.rm=TRUE),
      totaal  = sum(totaal, na.rm=TRUE)
    ) %>% 
    mutate(aanvoer = ifelse(is.na(aanvoer),0,aanvoer)) %>% 
    mutate(perc_discards = round((totaal - aanvoer) / totaal, digits=2)) %>% 
    mutate(week2 = year + week/52) %>% 
    mutate(yday = lubridate::yday(date))
    
  tt3 <-
    tt2 %>% 
    group_by(vessel, year, week) %>% 
    summarise(
      date = min(date), 
      perc_discards = mean(perc_discards, na.rm=TRUE)
    ) %>% 
    mutate(week2 = year + week/52) %>% 
    mutate(year = as.character(year)) %>% 
    group_by(vessel) %>% 
    mutate(ma5=zoo::rollapply(perc_discards,5,mean,align='right',fill=NA))

  tt2 %>%
    mutate(year = as.character(year)) %>% 
    ggplot(aes(x=week, y=perc_discards)) +
    theme_publication() +
    geom_point(aes(size=totaal, colour=year), alpha=0.4) +
    geom_line(data=tt3, aes(y=ma5, colour=year), size=1) +
    scale_size(
      range=c(0.5,5),
      limits =  c(100,1000*ceiling(max(tt2$totaal)/1000)),
      breaks =  seq(100,1000*ceiling(max(tt2$totaal)/1000),length.out=6)
    ) +
    scale_y_continuous(labels = scales::percent, limits=c(-0.1, 0.5)) +
    guides(size = guide_legend(nrow = 1)) +
    labs(x="week", y="%discards / day-week", size= "Totale vangst (kg)") + 
    labs(title="Percentage discards per dag/week SCH99 Aravis")

}

my_caption   <- ifelse(print_option, paste0("_", fig_nums("aanvoerentotalevangst"),"_"), "")
print_option <- FALSE

```

**Catch rate (landings / week)**

```{r cpue1, echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.height=fig_height_species}

fig_nums(
  name    = "cpue1", 
  level = 1, display = FALSE,
  caption = "CPUE (landings/day, kg) on log-scale. Outliers not shown on the figures. Red dot indicates medians. Blue line is the linear slope through the medians.")

t <-
  elog %>% 
  filter(species %in% top$species) %>%
  filter(year %in% longyrs) %>% 
  mutate(englishspecies = paste(englishname, species)) %>% 
  mutate(englishspecies = factor(englishspecies, levels=top$englishspecies)) %>% 
  drop_na(weight) %>% 
  group_by(vessel, englishspecies, year, date) %>% 
  summarise(catch_day = sum(weight, na.rm=TRUE)) %>% 
  ungroup() 

# remove outliers for display
t2 <-
  t %>% 
  group_by(englishspecies) %>% 
  mutate(catch_day = DescTools::Winsorize(catch_day, val = quantile(catch_day, probs = c(0.00, 0.9))))

# median
m <-
  t %>% 
  group_by(englishspecies, year) %>% 
  summarise(median = median(catch_day, na.rm=TRUE))

# slope
s <-
  m %>%
  group_by(englishspecies) %>%
  summarise(slope = scales::percent(lm(year ~ median)$coefficients[2], accuracy=1)) 

pp <-
  t2 %>%   
  ggplot(aes(x=year, y=catch_day, group=year)) +
  theme_publication() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(plot.margin = margin(1,1,1,1, "mm")) +
  theme(plot.title = element_text(hjust = 0.0)) +
  stat_summary(fun.data = calc_boxplot_stat, geom="boxplot") +  
  geom_point(data=m, 
             aes(x=year, y=median), 
             colour="red", inherit.aes = FALSE) +
  geom_smooth(data=m, 
              aes(x=year, y=median), 
              method="lm", colour="blue", inherit.aes = FALSE, se=FALSE) +
  geom_text(data=s, 
            aes(x=Inf, y=1, label=paste("slope:", slope)), 
            inherit.aes = FALSE, hjust="left", vjust="bottom", colour="blue") +
  expand_limits(y=0) +
  scale_x_continuous(breaks=seq(min(e$year), max(e$year),1)) +
  scale_y_log10() +
  labs(x="", y="landings/day (kg)", title="Average landings per day") +
  facet_wrap(~englishspecies, nrow = nrows_species)

print(pp)

```

_`r fig_nums("cpue1")`_

\newpage

**Average prices per year (euro)**


```{r prices, echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.height=fig_height_species}

fig_nums(
  name    = "prices", 
  level = 1, display = FALSE,
  caption = "Average prices of Jaczon flyshoot fleet (euro)")

t <-
  price_raw %>% 
  filter(species %in% top$species) %>%
  filter(year %in% longyrs) %>% 
  mutate(englishspecies = paste(englishname, species)) %>% 
  mutate(englishspecies = factor(englishspecies, levels=top$englishspecies)) %>% 
  group_by(englishspecies, year) %>% 
  summarise(
    avgprice = weighted.mean(avgprice, weight)
  ) %>% 
  ungroup() 

pp <-
  t %>%   
  ggplot(aes(x=year, y=avgprice)) +
  theme_publication() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(plot.margin = margin(1,1,1,1, "mm")) +
  theme(plot.title = element_text(hjust = 0.0)) +
  geom_point() +
  geom_line() +
  expand_limits(y=0) +
  scale_x_continuous(breaks=seq(min(t$year), max(t$year),1)) +
  labs(x="", y="avg price (euro)", title="Average price per year") +
  facet_wrap(~englishspecies, nrow = nrows_species)

print(pp)

```

_`r fig_nums("prices")`_

\newpage

**Revenue per day (euro)**

```{r cpue2, echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.height=fig_height_species}

fig_nums(
  name    = "cpue2", 
  level = 1, display = FALSE,
  caption = "RPUE (revenue/day, euro) on log-scale. Outliers not shown on the figures. Red dot indicates medians. Blue line is the linear slope through the medians.")

t <-
  revenue_byweek %>% 
  filter(species %in% top$species) %>%
  mutate(englishspecies = paste(englishname, species)) %>% 
  mutate(englishspecies = factor(englishspecies, levels=top$englishspecies)) %>% 
  group_by(vessel, englishspecies, year, week) %>% 
  summarise(
    revenue = sum(revenue, na.rm=TRUE),
    ndays   = sum(ndays, na.rm=TRUE)
  ) %>% 
  mutate(
    revenue_day = revenue / ndays
  ) %>% 
  ungroup() 

# remove outliers for display
t2 <-
  t %>% 
  group_by(englishspecies) %>% 
  mutate(revenue_day = DescTools::Winsorize(revenue_day, val = quantile(revenue_day, probs = c(0.00, 0.9))))

# median
m <-
  t %>% 
  group_by(englishspecies, year) %>% 
  summarise(median = median(revenue_day, na.rm=TRUE))

# slope
s <-
  m %>%
  group_by(englishspecies) %>%
  summarise(slope = scales::percent(lm(year ~ median)$coefficients[2], accuracy=1)) 

pp <-
  t2 %>%   
  ggplot(aes(x=year, y=revenue_day, group=year)) +
  theme_publication() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(plot.margin = margin(1,1,1,1, "mm")) +
  theme(plot.title = element_text(hjust = 0.0)) +
  stat_summary(fun.data = calc_boxplot_stat, geom="boxplot") +  
  geom_point(data=m, 
             aes(x=year, y=median), 
             colour="red", inherit.aes = FALSE) +
  geom_smooth(data=m, 
              aes(x=year, y=median), 
              method="lm", colour="blue", inherit.aes = FALSE, se=FALSE) +
  geom_text(data=s, 
            aes(x=Inf, y=1, label=paste("slope:", slope)), 
            inherit.aes = FALSE, hjust="left", vjust="bottom", colour="blue") +
  expand_limits(y=0) +
  scale_x_continuous(breaks=seq(min(elog$year), max(elog$year),1)) +
  scale_y_log10() +
  labs(x="", y="revenue/day (euro)", title="Average revenue per day") +
  facet_wrap(~englishspecies, nrow = nrows_species)

print(pp)

```

_`r fig_nums("cpue2")`_

\newpage

<!-- ########################################################################################## -->
<!-- Figure: Landings per Species and per Haul (map)                                            -->
<!-- ########################################################################################## -->

```{r generating_maps, echo=FALSE, fig.align="center", fig.asp=1.3, message=FALSE, warning=FALSE}

show_species <- c("MAC","SQR","CTC","MUR","WHG","HOM")

# Prepare data once (k uses gewicht from kisten, joined with haul positions)
t <-
  k %>%
  filter(species %in% show_species) %>% 
  filter(year %in% (yrs[length(yrs)]-3):yrs[length(yrs)]) %>% 
  filter(!(lat==0 & lon == 0)) %>% 
  mutate(date = lubridate::as_date(datetime)) %>% 
  mutate(quarter = lubridate::quarter(date)) %>% 
  group_by(species, vessel, year, quarter, rect, date, haul, lat, lon) %>%
  summarise(gewicht = sum(gewicht, na.rm=TRUE), .groups = 'drop') %>% 
  filter(gewicht >= 20)

world_cropped <-
  world_mr_df %>% 
  filter(
    long >= xmin, 
    long <= xmax,
    lat  >= ymin,
    lat  <= ymax
  )

tt <- scales::trans_breaks("sqrt", function(x) x^2)(c(0, max(t$gewicht)))

h_filtered <- h %>% filter(year %in% (yrs[length(yrs)]-3):yrs[length(yrs)])

t_by_species <- split(t, t$species)

for (myspecies in show_species) {
  
  t_sp <- t_by_species[[myspecies]]
  
  if (is.null(t_sp) || nrow(t_sp) == 0) next
  
  p <-
    ggplot(data=t_sp) +
    theme_publication() +
    theme(
      text             = element_text(size=12),
      plot.margin      = unit(c(0,0,0,0), "cm")
    ) +
    coord_quickmap(xlim=c(xmin, xmax), ylim=c(ymin, ymax)) +
    geom_polygon(data=world_cropped, aes(x=long, y=lat, group=group), fill = "grey75") +
    geom_point(data=h_filtered,
               aes(x=lon, y=lat, colour=as.character(year)),
               shape=3, size=0.5, stroke=0.1, alpha=0.4) +
    geom_point(aes(x=lon, y=lat, size = gewicht, colour = as.character(year)),
                alpha = 0.4, shape=21) +
    scale_size_continuous(breaks=tt, range=c(0.1, 5)) +
    labs(x = NULL, y = NULL, size = "kg aanvoer/trek", colour="",
         title=paste(myspecies,"Landings by haul")) +
    guides(size = guide_legend(nrow = 1)) + 
    facet_grid(quarter~year)

  print(p)
}

```


```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, fig.asp=1.0}

fig_nums(
  name    = "vangstsuccespersoort", level   = 1, display = FALSE,
  caption = "gemiddelde vangst (aanvoer, kg) per dag, per soort en per schip")

ee <- 
  elog %>%
  filter(species %in% top$species) %>% 
  mutate(vessel = ifelse(vessel %in% c("SCH99","Z99"), "SCH99/Z99", vessel)) %>% 
  mutate(species = factor(species, levels=c(top$species))) %>% 
  group_by(vessel, year, date, species) %>% 
  summarise(catchperday = sum(weight, na.rm=TRUE)) %>% 
  group_by(vessel, year, species) %>% 
  summarise(catchperday = mean(catchperday, na.rm=TRUE)) 

ee %>% 
  ggplot(aes(x=year, y=catchperday)) +
  theme_publication() +
  geom_point(data=ee %>% filter(year < max(ee$year, na.rm=TRUE)),
             aes(colour=vessel)) +
  geom_point(data=ee %>% filter(year == max(ee$year, na.rm=TRUE)), colour="red", size=2) +
  geom_line(aes(colour=vessel)) +
  labs(title=paste("Gemiddelde aanvoer per dag"), y="aanvoer per dag (kg)")  +
  expand_limits(y=0) +
  facet_wrap(~species, scales="free_y")

```
